### 2024-03-19 Notebook LaTEX/Machine Learning/Sections/section.tex
>w< mcbr

This is a new section. 
\subsection{Subsection}
This is a subsection.

\subsection{Cute Subsection}
Check out this cute figure. In \hyperref[fig:cute_alto]{fig \ref*{fig:cute_alto}} is cute pink unicorn, and in \hyperref[fig:minty_and_alto]{fig \ref*{fig:minty_and_alto}}, the green earth pony is minty, a cookie pone. Theyare cute together. Read source to understand the use of ``subfloat'' and ``figure'' together with ``hyperref''.

\begin{figure}
    \subfloat[Cute Alto]{
        \includegraphics[width=10em]{Assets/Alto_Legato.png}
        \label{fig:cute_alto}
    }
    \hfil
    \subfloat[Minty and Alto]{
        \includegraphics[width=10em]{Assets/Minty_and_Alto.png}
        \label{fig:minty_and_alto}
    }
\end{figure}