### 2024-04-23 Latex Templates/Generic Report/main.tex
>w< mcbr a lot of changes

\documentclass[]{article}

\input{presets/presets_basics.tex}

\title{Generic Report}
\author{Name of the Author}

\begin{document}
\maketitle

\begin{abstract}
    What is life? Why are we here? Where did we come from?
\end{abstract}

\section{Introduction}
    This is an introduction.

\section{Preliminaries}\label{sec:preliminaries}
    \hyperref[sec:preliminaries]{This} is the preliminary (hyperref without text labeling).

\section{Blah Blah Bleeeh}
    This is the blah blah bleeeh section. 
    \subsection{Blah Blah Blah Bleeh Bleeh Bleeh}
    Check out this cool Bibtext ref \cite[this]{texbook}, woooooooooah, also it's in plain style. For some mind altering psychodelic, read \hyperref[alg:mhc]{algorithm \ref*{alg:mhc}} for the experience. For some brain expanding julia code, read \hyperref[code:brain_expand]{brain expanding julia code}. 
    \begin{algorithm}[H]
        \begin{algorithmic}[t]
            \STATE{\textbf{Input: $X^{(t)}$}}
            \STATE{$Y^{(t)} \sim q (\cdot | X^{(t)})$}
            \STATE{
                $ 
                \rho(x, y) := 
                \min\left\lbrace
                    \frac{f(y)}{f(x)}\frac{q(x|y)}{q(y|x)}, 1
                \right\rbrace
                $ 
            }
            \STATE{
                $
                X^{(t + 1)} := 
                \begin{cases}
                    Y^{(t)} & \text{w.p}:  \rho(X^{(t)}, Y^{(t)})
                    \\
                    X^{(t)} &  \text{otherwise}
                \end{cases}$
            }
        \end{algorithmic}
        \caption{Metropolis Chain}
        \label{alg:mhc}
    \end{algorithm}
    \label{code:brain_expand}
    \lstinputlisting[language=julia, basicstyle=\ttfamily\scriptsize,numbers=left]{Code/juliacode.jl}

\appendix
\section{Bleeh Bleeh Bleeh I am not Listening}
    This is the Bleeh Bleeh Bleeh I am not Listening section.

\section{This section is in another .tex file} 
    \input{Sections/section.tex}

\bibliographystyle{plain}
\bibliography{refs.bib}

\end{document}