### 2024-11-28 Latex Templates/Scientific 2/main.tex
>w<, changes

\documentclass[12pt]{article}
\input{presets/wang/input.tex}
\input{presets/misc.tex}
\input{presets/julia_lstlisting.tex}

% === TEMPLATE HOW TO USE ======
% To begine here is a list of things: 
% [ ]: Change title. 
% [ ]: Fill into the author names, affiliation and contact info. 
% [ ]: Fill in the abstract. 
% [ ]: Fill/change in AMS mathematics subject classification code, and the keywords.

\begin{document}

\title{
    {
        \fontfamily{ptm}\selectfont 
        A big title
    }
    }

\author{
    \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    The University of British Columbia, Kelowna, BC Canada V1V 1V7. 
    E-mail:  \texttt{alto@mail.ubc.ca}.}~Author Name 1
    \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    The University of British Columbia, Kelowna, BC Canada V1V 1V7. 
    E-mail:  \texttt{shawn.wang@ubc.ca}.}~ and~Author Name 2
}

\date{\today}

\maketitle

% \vskip 8mm

\begin{abstract} 
    \noindent
    Abstract abstract abstract. 
\end{abstract}
% \noindent{\bfseries 2010 Mathematics Subject Classification:}
% Primary 65K10, 90c25, 90C30; Secondary 65Y20. 
% \noindent{\bfseries Keywords: } Non-convex Optimizations, Proximal Point.

\tableofcontents

\section{Introduction}

\section{Notations}
    Beck's book \cite{beck_first-order_2017}. 
\subsection{Foundamentals in convex analysis}
\subsection{Canonical forms of Nesterov's accelerated methods}
\section{Unifying variants of Nesterov's accelerated methods}
\section{Generic convergence anlaysis of accelerated methods}
\section{Catalyst accelerations and future works}
\section{Performence estimation problems}
\section{Methods of inexact proximal point}
\section{Nestrov's acceleration in the non-convex case}

\bibliographystyle{siam}
\bibliography{references/refs.bib}
\newpage

\appendix

\end{document}