### 2024-12-20 Notebook LaTEX/Momentum Accelerations/paper3.tex
>w<, changes

\documentclass[12pt]{article}

% \input{presets/wang_full.tex}
\input{presets/wang/input.tex}
\input{presets/misc.tex}
\input{presets/julia_lstlisting.tex}

\begin{document}

\title{{\fontfamily{ptm}\selectfont 
        A Parameter Free Accelerated Proximal Gradient Method Without Restarting
    }}

\author{
    Hongda Li
    \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    The University of British Columbia, Kelowna, BC Canada V1V 1V7. E-mail:  \texttt{alto@mail.ubc.ca}.}~ and ~Xianfu Wang
    \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    The University of British Columbia, Kelowna, BC Canada V1V 1V7. E-mail:  \texttt{shawn.wang@ubc.ca}.}
    % and ~Heinz H. Bauschke~
    % \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    % The University of British Columbia, Kelowna, BC Canada V1V 1V7. E-mail:  \texttt{heinz.bauschke@ubc.ca}.
    % }
}

\date{\today}

\maketitle

% \vskip 8mm

\begin{abstract} 
    \noindent

    \todoinline{Write this abstract when finishing. }
\end{abstract}

\noindent{\bfseries 2010 Mathematics Subject Classification:}
Primary 65K10, 90c25, 90C30; Secondary 65Y20. 
\noindent{\bfseries Keywords: } Nesterov acceleration, Proximal point method. 
\section{Introduction}
    Suppose we are minimizing a differentiable convex function $F: \RR^n \rightarrow \RR$ for simplicity. 
    Let $F$ be Lipschitz smooth meaning that there exists $L \in \RR$ such that for all $x \in \RR^n, y\in \RR^n$ it has $\Vert \nabla F(x) - \nabla F(y)\Vert \le L \Vert x - y \Vert$. 
    \par 
    A good choice of a minimization algorithm is the Nesterov's acceleration schemes that originally proposed in 1983 \cite{nesterov_method_1983}. Initialize $x_1 = y_1$ and $t_0 = 1$, the algorithm finds $(x_k)_{k \ge 1}$ for all $k \ge 1$ by: 
    \begin{align}
        & x_{k + 1} = y_k - L^{-1}\nabla F(y_k), 
        \\
        & t_{k + 1} = 1/2\left(1 + \sqrt{1 + 4t_{k}^2}\right), 
        \\
        & \theta_{k + 1} = (t_{k} - 1)/t_{k + 1}, 
        \\
        & y_{k + 1} = x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k). 
    \end{align}\label{eqn:example_algorithm}
    It's well known that if the minimizer $x^*$ exists, $F(x_k) - F(x^*)$ decreases at a rate of $\mathcal O(1/k^2)$, faster than the gradient descent method which is $\mathcal O(1/k)$, and it's optimal. 
    See Chapter 2 of Nesterov's book \cite{nesterov_lectures_2018}. 
    However, everything is not sunshine and rainbow because the above scheme is associated with an issue. 
    When $F$ is $\mu > 0$ strongly convex, the algorithm oscillates and converges sub-linearly, making it worse than gradient descent. See page 9 of Su et al. \cite{su_differential_2016}. 
    \par
    This issue motivated a vast amount of literatures aims at improving, interpreting, and extending this method. 
    Restarting is a popular solution to address the issue of obtaining faster convergence rate when the objective function is strongly convex. 
    Beck and Toubelle \cite{beck_fast_2009} mitigate the issue by restarting and showed that it $\mathcal O(1/k^2)$ convergence rate is kept and it performs empirically better. 
    See \cite{necoara_linear_2019}\cite{aujol_parameter-free_2024} and references within for recent advancements in restarting accelerated proximal gradient algorithm. 
    \par
    Inspired by prior works and the method of Nestrov's estimating sequence \cite{nesterov_lectures_2018}, we present to you firstly, an alternative to restarting that performs well empirically. 
    Secondly, a unified framework of Accelerated Proximal Gradient (APG) which we call Relaxed Weak Accelerated Proximal Gradient (R-WAPG). 
    The R-WAPG has a convergence claim of $F(x_k)- F(x^*)$ for a much more flexible choice of $(\theta_t)_{t \ge 1}$. 
    It also has descriptive power to describe several variants of FISTA in the literatures.

    \subsection{Contributions}
        Our contributions are two folds, theoretical and practical. 
        Nesterov's acceleration extrapolate $y_{k + 1} = x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k)$ where $\theta_{k + 1} = (t_{k} - 1)/t_{k + 1} \in (0, 1)$ is the ``momentum''. 
        The choices for $\theta_k$ varies. 
        In Chamboolle, Dossal \cite{chambolle_convergence_2015}, it has $t_k = (n + a - 1)/a$ for all $a > 2$ gives strong weak convergence of the iterates $x_k$ in Hilbert space. 
        In Beck \cite{beck_first-order_2017}, it has constant $\theta_t = (\sqrt{\kappa} - 1)/(\sqrt{\kappa} + 1)$ where $\kappa = \mu/L$. 
        \par
        Firstly, we relaxed the traditional choice of the sequence $\theta_k$ in Equation \ref{eqn:example_algorithm} and showed an upper bound of the optimal gap. 
        Let $(\alpha_k)_{k \ge0}, (\rho_k)_{k \ge 0}$ be two sequences that satisfy
        \begin{align*}
            \alpha_0 &\in (0, 1], 
            \\
            \alpha_k &\in (\mu/L, 1) \quad (\forall k \ge 1), 
            \\
            \rho_k &:= \frac{\alpha_{k + 1}^2 - (\mu/L)\alpha_{k + 1}}{(1 - \alpha_{k + 1})\alpha_k^2} \quad \forall (k \ge 0). 
        \end{align*}
        Our first main result shows that if $\theta_{k + 1} = (\rho_k\alpha_k(1 - \alpha_k)/(\rho_k\alpha_k^2 + \alpha_{k + 1}))$, using the R-WAPG we proposed in Definition \ref{def:wapg} with Proposition \ref{prop:wagp-convergence}, \ref{prop:r-wapg-momentum-repr}, we can show that the gap $F(x_k) - F(x^*)$ is bounded by:
        \begin{align*}
            \mathcal O\left(
                \left(
                    \prod_{i = 0}^{k - 1} \max(1, \rho_{k})
                \right)
                \prod_{i = 1}^{k} \left(1  - \alpha_i\right)
            \right). 
        \end{align*}
        Our second main result shows that there exists $\rho_k > 1$ such that our R-WAPG reduces perfectly to a variant of FISTA proposed in Chambolle Dossal \cite{chambolle_convergence_2015} with the same convergence claim. 
        And when $\rho_k = 1$, R-WAPG reduces perfectly to V-FISTA by Beck \cite{beck_first-order_2017} if $\mu > 0$ and FISTA if $\mu = 0$ and the convergence claim is consistent with the literatures. 
        \par
        Our practical contribution is an algorithm inspired by a detail in our convergence proof which we call it ``Parameter Free R-WAPG''. 
        The algorithm is parameter free, meaning that it doesn't require knowing $L, \mu$ in advance, and it determines the value of $\theta_t$ by estimating the local concavity using iterates $y_{k}, y_{k + 1}$ with minimal computational cost. 
        We conducted ample amount of numerical experiments to show that it has a favorable convergence rate in practice and behaves similarly to the FISTA with monotone restart. 
    \subsection{Prior works}

    \subsection{Organizations}

\section{Preliminaries}
    Throughout, we assume the problem of types $F = f + g$ where $f: \RR^n \rightarrow \RR$ is $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex and $g: Q \rightarrow \overline \RR$ is convex. 
    We consider optimization problem: $\min_x \left\lbrace F(x) \defeq f(x) + g(x)\right\rbrace$.
    Fix arbitrary $y \in \RR^n$, define
    \begin{align*}
        \widetilde{\mathcal M}^{L^{-1}}
        (x; y) \defeq
        g(x) + f(y) + \langle \nabla f(y), x - y\rangle 
        + \frac{L}{2}\Vert x - y\Vert^2, 
        \\
        \mathcal M^{L^{-1}}(x; y) := F(x) + \frac{L}{2}\Vert x - y\Vert^2, 
    \end{align*}
    and the proximal gradient operator $T_L = [I + L^{-1}\partial g]^{-1}\circ [I - L^{-1}\nabla f]$. 
    Define the Bregman divergence of $f$: 
    \begin{align*}
        D_f(x, y): \RR^n \times \RR^n \rightarrow \RR 
        \defeq f(x) - f(y) - \langle \nabla f(y), x - y\rangle. 
    \end{align*}
    The lemma follows explains equates between $\mathcal M(\cdot; y), \widetilde{\mathcal M}(\cdot; y)$. 
    \begin{lemma}[Proximal gradient envelope] 
        Fix any $y$, we will have for all $x \in \RR^n$ that: 
        \begin{align*}
            \widetilde{\mathcal M}^{L^{-1}}(x; y)
            &= 
            \mathcal M^{L^{-1}}(x; y)- D_f(x, y) \le \mathcal M^{L^{-1}}(x; y). 
        \end{align*}
    \end{lemma}
    \begin{proof}
        \begin{align*}
            \widetilde{\mathcal M}^{L^{-1}}(x; y) 
            &= 
            g(x) + f(y) + \langle \nabla f(y), x - y\rangle + \frac{L}{2}\Vert x - y\Vert^2
            \\
            &= 
            g(x) + f(x) - f(x) + f(y) 
            + \langle \nabla f(y), x - y\rangle + \frac{L}{2}\Vert x - y\Vert^2
            \\
            &= 
            F(x) - D_f(x, y) + \frac{L}{2}\Vert x - y\Vert^2 
            \\
            &= \mathcal M^{L^{-1}}(x; y) - D_f(x, y). 
        \end{align*}
    \end{proof}
    \subsection{Proximal inequality}
    \begin{theorem}[Proximal inequality]\label{thm:prox-grad-ineq}
        Fix any $y$, we have for all $x$: 
        \begin{align*}
            F(x) - F(T_Ly) - \langle L(y - T_Ly), x - y\rangle
            - L\Vert y - T_L y\Vert^2
            - D_f(x, y)
            &\ge  
            0. 
        \end{align*}
        When $f$ is $\mu \ge 0$ strongly convex, it implies that 
        \begin{align*}
            F(x) - F(T_Ly) - \langle L(y - T_Ly), x - y\rangle
            - \frac{L}{2}\Vert y - T_L y\Vert^2
            - \frac{\mu}{2}\Vert x - y\Vert^2
            &\ge  
            0.
        \end{align*}
    \end{theorem}
    \begin{proof}
        Observe that the function $\widetilde {\mathcal M}(\cdot; x)$ is $L$ strongly convex with minimizer $T_Ly$. 
        For all $x \in \RR^n$, it has quadratic growth condition at its minimizer $T_Ly$ which gives these equivalent inequalities: 
        {\small
        \begin{align*}
            \widetilde{\mathcal M}^{L^{-1}}(x; y) - 
            \widetilde{\mathcal M}^{L^{-1}}(T_Ly; y)
            - 
            \frac{L}{2}\Vert x - T_Ly\Vert^2
            &\ge 
            0
            \\
            \iff
            \left(
                \mathcal M^{L^{-1}}(x; y) - D_f(x, y)
            \right) - 
            \mathcal M^{L^{-1}}(T_Ly; y) 
            - 
            \frac{L}{2}\Vert x - T_Ly\Vert^2
            &\ge 0
            \\
            \iff 
            \left(
                \mathcal M^{L^{-1}}(x; y)
                - 
                \mathcal M^{L^{-1}}(T_Ly; y)
            \right)
            - 
            D_f(x, y) 
            - \frac{L}{2}\Vert x - T_Ly\Vert^2
            &\ge 0
            \\
            \iff 
            \left(
                F(x) - F(T_Ly) 
                + 
                \frac{L}{2}\Vert x - y\Vert^2 - 
                \frac{L}{2}\Vert T_Ly - y\Vert^2
            \right)
            - 
            D_f(x, y) 
            - \frac{L}{2}\Vert x - T_Ly\Vert^2
            &\ge 0
            \\
            \iff 
            \left(
                F(x) - F(T_Ly) 
                + 
                \frac{L}{2}
                \left(
                    \Vert x - T_Ly + T_Ly - y\Vert^2
                    - 
                    \Vert y - T_Ly\Vert^2
                \right)
            \right)
            - 
            D_f(x, y) 
            - \frac{L}{2}\Vert x - T_Ly\Vert^2
            &\ge 0
            \\
            \iff 
            \left(
                F(x) - F(T_Ly) 
                + 
                \frac{L}{2}
                \left(
                    \Vert x - T_Ly\Vert^2 + 
                    2\langle x - T_Ly, T_Ly - y\rangle
                \right)
            \right)
            - 
            D_f(x, y) 
            - \frac{L}{2}\Vert x - T_Ly\Vert^2
            &\ge 0
            \\
            \iff
            \left(
                F(x) - F(T_Ly) + \frac{L}{2}\Vert x - T_Ly\Vert^2 
                - L\langle  x - T_Ly, y - T_Ly\rangle
            \right)
            - 
            D_f(x, y) 
            - \frac{L}{2}\Vert x - T_Ly\Vert^2
            &\ge 0
            \\
            \iff 
            F(x) - F(T_Ly)
            - \langle L(y - T_Ly), x - T_Ly\rangle
            - D_f(x, y) 
            &\ge 0. 
        \end{align*}
        }
        Continuing, it has: 
        \begin{align*}
            F(x) - F(T_Ly) - \langle L(y - T_Ly), x - T_Ly\rangle - D_f(x, y) &\ge 0
            \\
            \iff
            F(x) - F(T_Ly)
            - \langle L(y - T_Ly), x - y + y - T_Ly\rangle - D_f(x, y) 
            &\ge 0
            \\
            \iff
            F(x) - F(T_Ly)
            - \langle L(y - T_Ly),x - y \rangle
            - L\Vert y - T_Ly\Vert^2
            - D_f(x, y) 
            &\ge 0
            \\
            \implies 
            F(x) - F(T_Ly)
            - \langle L(y - T_Ly),x - y \rangle
            - \frac{L}{2}\Vert y - T_Ly\Vert^2
            - D_f(x, y) 
            &\ge 0
            \\
            \implies
            F(x) - F(T_Ly)
            - \langle L(y - T_Ly),x - y \rangle
            - \frac{L}{2}\Vert y - T_Ly\Vert^2
            - \frac{\mu}{2}\Vert x - y\Vert^2
            &\ge 0. 
        \end{align*}
        Going from the second last inequality to the last inequality, we used the fact that $f$ is strongly convex hence for all $x, y$, it's true that $D_f(x, y) \ge \mu/2 \Vert x - y\Vert^2$. 
    \end{proof}

\section{Stepwise formulation of weak accelerated proximal gradient}
    This entire section is building up the R-WAPG algorithm which is described in the Definition \ref{def:wapg} of the next section. 
    \par 
    Definition \ref{def:stepwise-wapg} which describes what happens at a single iteration of the R-WAPG algorithm. 
    It defines a procedure of generating $x_{k + 1}, v_{k + 1}$ given any $x_k, v_k$. 
    Proposition \ref{prop:stepwise-lyapunov} states the inequality that describes a decreasing quantity that involves $F(x_k), F(x_{k + 1})$ at each single iteration. 
    \begin{assumption}
        Choose any integer $k\ge 0$. 
        Given $x_k, y_k, v_k$, we define the following quantities
        \begin{align}
            g_k &\defeq L(y_k - T_L y_k), 
            \label{eqn:grad-map}
            \\
            l_F(x; y_k) &\defeq F(T_Ly_k) + \langle g_k, x - y_k\rangle + \frac{1}{2L}\Vert g_k\Vert^2, 
            \label{eqn:lower-linearization}
            \\
            \epsilon_{k} &\defeq F(x_k) - l_F(x_k; y_k), 
            \label{eqn:regret}
        \end{align}
        Observe that by convexity of $F$, $\epsilon_k \ge 0$ for all $x_k, L > 0$. 
        To see, use Theorem \ref{thm:prox-grad-ineq} and let $y = y_k, x = x_k$ which gives: 
        \begin{align*}
            F(x_k) - F(T_Ly_k)
            - \langle L(y_k - T_Ly_k),x_k - y_k \rangle
            - \frac{L}{2}\Vert y_k - T_Ly_k\Vert^2
            - \frac{\mu}{2}\Vert x_k - y_k\Vert^2
            &\ge 0
            \\
            \iff 
            F(x_k) - F(T_Ly_k)
            - \langle g_k,x_k - y_k \rangle
            - \frac{1}{2L}\Vert g_k\Vert^2
            &\ge 0. 
        \end{align*}
    \end{assumption}

    \begin{definition}[Stepwise weak accelerated proximal gradient]\label{def:stepwise-wapg}\;\\
        Assume $0 \le \mu < L$.
        Fix any $k \in \mathbb Z$. 
        For any $(v_k, x_k), \alpha_k \in (0, 1), \gamma_k > 0$, let $\hat \gamma_{k + 1}$, and vectors $y_k, v_{k + 1}, x_{k + 1}$ be given by: 
        \begin{align*}
            \hat \gamma_{k + 1} &= (1 - \alpha_k)\gamma_k + \mu \alpha_k, 
            \\
            y_k &= 
            (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k), 
            \\
            g_k &= \mathcal G_L y_k, 
            \\
            v_{k + 1} &= \hat\gamma^{-1}_{k + 1}
            (\gamma_k(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k), 
            \\
            x_{k + 1} &= T_L y_k. 
        \end{align*}
    \end{definition}
    \begin{observation}\label{obs:stepwise-wapg}
        Let's observe Definition \ref{def:stepwise-wapg} closely. 
        We make two crucial observations here. 
        We have $x_k, y_k, v_k$ lie on the same line because and they have the following equivalent equalities: 
        \begin{align*}
            (Q1): 
            y_k - v_k &= 
            \frac{\hat \gamma_{k + 1}}{\alpha_k \gamma_k}(x_k - y_k),
            \\
            (Q2): 
            y_k - x_k &= 
            \frac{\alpha_k \gamma_k}{\gamma_k + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
    \end{observation}
        To see (Q1), observe 
        \begin{align*}
            y_k - v_k &= 
            \frac{\hat \gamma_{k + 1}}{\alpha_k \gamma_k}(x_k - y_k)
            \\
            \iff 
            -(\alpha_k \gamma_k \hat \gamma^{-1}_{k + 1} + 1)y_k
            &= 
            - \alpha_k \gamma_k \hat \gamma^{-1}_{k + 1}v_k - x_k
            \\
            \iff 
            y_k &= 
            \frac{
                \alpha_k \gamma_k \hat \gamma_{k + 1}^{-1}v_k + x_k
            }{1 + \alpha_k \gamma_k \hat \gamma_{k + 1}^{-1}}
            \\
            &=  
            \frac{\alpha_k \gamma_k v_k + \hat \gamma_{k + 1} x_k}{\gamma_k + \alpha_k \mu}.
        \end{align*}
        On the first equality (Q1), we multiplied both side of the equation by $\hat\gamma_{k + 1}/\alpha_k \gamma_k$ and then group $y_k$ on the LHS. 
        The last equality comes by multiplying both the numerator and denominator by $\hat \gamma_{k + 1}$, leaving the numerator $\hat \gamma_{k + 1} + \alpha_k \gamma_k = \gamma_k + \alpha_k \mu$. 
        We used the definition of $\hat \gamma_{k + 1}$ here. 
        To see the second equality (Q2), consider: 
        \begin{align*}
            y_k &= (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k)
            \\
            \iff
            y_k - x_k &= 
            (\gamma_k + \alpha_k \mu)^{-1}
            (\alpha_k \gamma_k v_k - (\gamma_k + \alpha_k \mu)x_k + \hat \gamma_{k + 1} x_k)
            \\
            \iff 
            (\gamma_k + \alpha_k \mu)(y_k - x_k)
            &= 
            \alpha_k\gamma _kv_k + 
            (\hat \gamma_k - \gamma_k - \alpha_k \mu) x_k
            \\
            &= \alpha_k \gamma_k v_k - \alpha_k \gamma_k x_k 
            \\
            &= \alpha_k \gamma_k(v_k - x_k)
            \\
            \iff 
            y_k - x_k &= 
            \frac{\alpha_k \gamma_k}{\gamma_k + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
        On the second equality that follows the second $\iff$ we just substituted $\hat\gamma_{k + 1} = (1 - \alpha_k)\gamma_k + \alpha_k \mu$. 
        Therefore, we discover that $x_k, v_k, y_k$ lies on the same line because (Q1) indicates $y_k - v_k$ parallels to $-(y_k - x_k)$ and both vector shares the same head which anchors at $y_k$.

    \begin{proposition}[Stepwise Lyapunov]\label{prop:stepwise-lyapunov}\;\\
        Fix any integer $k \in \mathbb Z$.
        Given any $v_k, x_k$ and $\gamma_k > 0$, invoke Definition \ref{def:stepwise-wapg} to obtain $v_{k + 1}, x_{k + 1}, y_k, \hat \gamma_{k + 1}$. 
        Fix any arbitrary $R_k \in \RR$.
        Define: 
        \begin{align*}
            R_{k + 1}
            \defeq
            \frac{1}{2}\left(
                L^{-1} - \frac{\alpha_k^2}{\hat \gamma_{k + 1}}
            \right)\Vert g_k\Vert^2
            + 
            (1 - \alpha_k)
            \left(
                \epsilon_k + R_k + 
                \frac{\mu\alpha_k\gamma_k}{2\hat \gamma_{k + 1}}
                \Vert v_k - y_k\Vert^2
            \right). 
        \end{align*}
        Then it has for all $x^* \in \RR^n$ where $F^* = F(x^*)$, the inequality: 
        \begin{align*}
            F(x_{k + 1}) - F^* + R_{k + 1} + \frac{\hat \gamma_{k + 1}}{2}\Vert v_{k + 1} - x^*\Vert^2
            &\le 
            (1 - \alpha_k)
            \left(
                F(x_k) - F^* + R_k + \frac{\gamma_{k}}{2}\Vert v_k - x^*\Vert^2
            \right). 
        \end{align*}
    \end{proposition}
    \begin{proof}
        The proof is a direct proof. 
        Here we drop the subscript $k$, $k + 1$ on $\gamma_{k}, \hat \gamma_{k + 1}$ because they are fixed throughout the proof, and it makes for better readability! 
        \par 
        We started with $F(x_{k + 1}) + R_{k + 1}$ as Equation (1*), and $\hat \gamma/2\Vert v_{k + 1} - x^*\Vert^2$ as (2*). 
        In the proof that follows, we show (2*) equals to (2.2*). 
        We added (2*), (2.2*) to produce (3*). 
        Then we show (3*) equals to (3.1*). 
        The content of the entire LHS of the conclusion of the claim is conclused in (3.2*). 
        Going from (3.2*) to (4*) is an inequality which we used Theorem \ref{thm:prox-grad-ineq}. 
        \par
        Start by considering the first and the third term of the LHS of the inequality that we want to prove. 
        \begin{align*}
            F(x_{k + 1}) &= 
            F(x_k) - \epsilon_k - \langle  g_k, x_k - y_k\rangle - \frac{1}{2L}\Vert g_k\Vert^2, 
            \\
            R_{k + 1}
            &= 
            \frac{1}{2}\left(
                L^{-1} - \frac{\alpha_k^2}{\hat \gamma}
            \right)\Vert g_k\Vert^2
            + 
            (1 - \alpha_k)
            \left(
                \epsilon_k + R_k + 
                \frac{\mu\alpha_k\gamma}{2\hat \gamma}
                \Vert v_k - y_k\Vert^2
            \right), 
            \\
            \implies 
            F(x_{k + 1}) + R_{k + 1}
            &= 
            F(x_k) - \epsilon_k - \langle  g_k, x_k - y_k\rangle
            - \frac{\alpha_k^2}{2\hat \gamma}\Vert g_k\Vert^2
            \\
            &\quad 
                + 
                (1 - \alpha_k)
                \left(
                    \epsilon_k + R_k + 
                    \frac{\mu\alpha_k\gamma}{2\hat \gamma}
                    \Vert v_k - y_k\Vert^2
                \right). 
        \tag{1*}
        \end{align*}
        Coefficient of $\epsilon_k$ and $\Vert g_k\Vert^2$ are grouped. 
        Next, we have: 
        \begin{align*}
            \frac{\hat \gamma}{2}\Vert v_{k + 1} - x^* \Vert^2
            &= 
            \frac{\hat \gamma}{2}\Vert 
                \hat \gamma^{-1}
                (
                    \gamma(1 - \alpha_k)v_k - 
                    \alpha_k g_k + \mu \alpha_k y_k
                )
                - x^* 
            \Vert^2
            \\
            &=  
            \frac{\hat \gamma}{2}
            \Vert 
                \hat \gamma^{-1}
                (
                \hat \gamma v_k + \mu \alpha_k(y_k - v_k)
                    - \alpha_k g_k
                )
                - x^* 
            \Vert^2
            \\
            &= 
            \frac{\hat \gamma}{2}
            \Vert 
                v_k + \hat \gamma^{-1} \mu \alpha_k (y_k - v_k)
                - \hat \gamma^{-1}\alpha_k g_k
                - x^* 
            \Vert^2
            \\
            &= 
            \frac{\hat \gamma}{2}
            \Vert v_k - x^*\Vert^2 
            + 
            \frac{\alpha_k^2}{2\hat \gamma}\Vert \mu(y_k - v_k) - g_k\Vert^2 
            \\ &\quad 
                + 
                \langle v_k - x^*, \mu \alpha_k(y_k - v_k) - \alpha_k g_k\rangle
            \\
            &= 
            \left(
            \frac{(1 - \alpha_k)\gamma + \mu \alpha_k}{2} 
            \right)\Vert v_k - x^*\Vert^2
            \\ &\quad
                + 
                \frac{\alpha_k^2}{2\hat \gamma}
                \Vert \mu(y_k - v_k) - g_k\Vert^2 
                + 
                \langle v_k - x^*, \mu \alpha_k(y_k - v_k) - \alpha_k g_k\rangle. 
        \tag{2*}
        \end{align*}
        On the above derivation, the first equality is by definition of $v_{k + 1}$; the second equality is by $\hat \gamma = (1 - \alpha_k)\gamma + \mu \alpha_k$ the second equality comes by considering: 
        \begin{align*}
            \gamma(1 - \alpha_k) v_k &= 
            (\hat \gamma  - \mu \alpha_k)v_k
            = \hat \gamma v_k - \mu\alpha_k v_k
            \\
            \iff 
            \gamma(1 - \alpha_k) v_k + \mu \alpha_k y_k
            &= 
            \hat \gamma v_k + \mu \alpha_k(y_k - v_k). 
        \end{align*}
        Focusing on the last two terms by the end of expression (2*), we have  
        \begin{align*}
            \frac{\alpha^2_k}{2\hat \gamma} 
            \Vert \mu(y_k - v_k) - g_k\Vert^2
            & = 
            \frac{\alpha_k^2\mu}{\hat \gamma}
            \left(
                \frac{\mu}{2}\Vert y_k - v_k\Vert^2 
                - \langle y_k - v_k, g_k\rangle
            \right)
            + \frac{\alpha_k^2}{2\hat \gamma}\Vert g_k\Vert^2, 
            \\
            \langle v_k - x^*, \mu \alpha_k(y_k - v_k) - \alpha_k g_k\rangle
            &= 
            \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle 
            - \alpha_k \langle v_k - x^*, g_k\rangle. 
        \tag{2.1*}
        \end{align*}
        Adding the LHS of both equations above together gives: 
        {\small
        \begin{align*}
            & \quad 
            \frac{\alpha^2_k}{2\hat \gamma} 
            \Vert \mu(y_k - v_k) - g_k\Vert^2
            + 
            \langle v_k - x^*, \mu \alpha_k(y_k - v_k) - \alpha_k g_k\rangle
            \\
            &= 
            \left\langle g_k, 
                - \alpha_k(v_k - x^*) 
                - \frac{\alpha_k^2\mu}{\hat \gamma}(y_k - v_k)
            \right\rangle
            + \frac{\alpha_k^2}{2\hat \gamma}\Vert g_k\Vert^2
            + \frac{\alpha_k^2 \mu^2}{2\hat \gamma}\Vert y_k - v_k\Vert^2
            + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle. 
        \end{align*}
        }
        With the above we can conclude that (2*) simplifies to 
        {\small
        \begin{align*} 
            & 
            \left(
            \frac{(1 - \alpha_k)\gamma + \mu \alpha_k}{2} 
            \right)\Vert v_k - x^*\Vert^2
            + 
            \left\langle g_k, 
                - \alpha_k(v_k - x^*) 
                - \frac{\alpha_k^2\mu}{\hat \gamma}(y_k - v_k)
            \right\rangle
            \\
            & \quad 
                + \frac{\alpha_k^2}{2\hat \gamma}\Vert g_k\Vert^2
                + \frac{\alpha_k^2 \mu^2}{2\hat \gamma}\Vert y_k - v_k\Vert^2
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle. 
        \tag{2.2*}
        \end{align*}
        }
        Recall from 
        \hyperref[obs:stepwise-wapg]{Observations \ref*{obs:stepwise-wapg}} that: 
        \begin{align*}
            (Q1): 
            y_k - v_k &= 
            \frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k),
            \\
            (Q2): 
            y_k - x_k &= 
            \frac{\alpha_k \gamma}{\gamma + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
        Ok that is a lot, we list the following equations to assist things: 
        \begin{align*}
            &  
            - \alpha_k(v_k - x^*) - \frac{\alpha_k^2 \mu}{\hat \gamma}(y_k - v_k) - (x_k - y_k)
            \\
            & =
            -\alpha_k(v_k - x^*) -
            \frac{\alpha_k^2\mu}{\hat \gamma}\frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k)
            - (x_k - y_k) 
            &\text{by Q1} 
            \\
            &= 
            -\alpha_k(v_k - x^*) -
            \frac{\alpha_k \mu}{\gamma}(x_k - y_k)
            - (x_k - y_k) 
            \\
            &= 
            -\alpha_k(v_k - x^*) -
            \left(
                1 + \frac{\alpha_k \mu}{\gamma}
            \right)(x_k - y_k)
            \\
            &= 
            -\alpha_k(v_k - x^*) - 
            \frac{\alpha_k \mu + \gamma}{\gamma}
            \frac{\alpha_k \gamma}{\gamma + \alpha_k \mu}(x_k - v_k)
            &\text{by Q2}
            \\
            &= 
            -\alpha_k(v_k - x^*)
            - \alpha_k(x_k - v_k)
            \\
            &= \alpha_k(x^* - x_k). 
        \tag{Q3}
        \end{align*}
        Adding (2.2*) to (1*) gives: 
        \begin{align*}
            &
            F(x_k) - \epsilon_k - \langle  g_k, x_k - y_k\rangle
            - \frac{\alpha_k^2}{2\hat \gamma}\Vert g_k\Vert^2
            + (1 - \alpha_k)
            \left(
                \epsilon_k + R_k + 
                \frac{\mu\alpha_k\gamma}{2\hat \gamma}
                \Vert v_k - y_k\Vert^2
            \right)
            \\
            &\quad 
                + 
                \left(
                \frac{(1 - \alpha_k)\gamma + \mu \alpha_k}{2} 
                \right)\Vert v_k - x^*\Vert^2
                + 
                \left\langle g_k, 
                    - \alpha_k(v_k - x^*) 
                    - \frac{\alpha_k^2\mu}{\hat \gamma}(y_k - v_k)
                \right\rangle
            \\
            & \quad 
                + \frac{\alpha_k^2}{2\hat \gamma}\Vert g_k\Vert^2
                + \frac{\alpha_k^2 \mu^2}{2\hat \gamma}\Vert y_k - v_k\Vert^2
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle
            \\
            &= 
            F(x_k) - \epsilon_k 
            + \left\langle 
                g_k, 
                - \alpha_k(v_k - x^*) 
                - \frac{\alpha_k^2\mu}{\hat \gamma}(y_k - v_k)
                - (x_k - y_k)
            \right\rangle
            \\
            &\quad 
                + (1 - \alpha_k)
                \left(
                    \epsilon_k + R_k + 
                    \frac{\mu\alpha_k\gamma}{2\hat \gamma}
                    \Vert v_k - y_k\Vert^2
                \right)
                + 
                \left(
                \frac{(1 - \alpha_k)\gamma + \mu \alpha_k}{2} 
                \right)\Vert v_k - x^*\Vert^2
            \\
            & \quad 
                + \frac{\alpha_k^2 \mu^2}{2\hat \gamma}\Vert y_k - v_k\Vert^2
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle
            \\
            \text{Use Q3}&= 
            F(x_k) - \epsilon_k 
            + \alpha_k\left\langle 
                g_k, 
                x^* - x_k
            \right\rangle
            \\
            &\quad 
                + (1 - \alpha_k)
                \left(
                    \epsilon_k + R_k + 
                    \frac{\mu\alpha_k\gamma}{2\hat \gamma}
                    \Vert v_k - y_k\Vert^2
                \right)
                + 
                \left(
                \frac{(1 - \alpha_k)\gamma + \mu \alpha_k}{2} 
                \right)\Vert v_k - x^*\Vert^2
            \\
            & \quad 
                + \frac{\alpha_k^2 \mu^2}{2\hat \gamma}\Vert y_k - v_k\Vert^2
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle
            \\
            &= 
            F(x_k) - \alpha_k\epsilon_k + \alpha_k\langle g_k, x^* - x_k\rangle
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\&\quad 
                + \frac{(1 - \alpha_k)\mu\alpha_k\gamma}{2\hat \gamma}\Vert v_k - y_k\Vert^2
                + \frac{\mu \alpha_k}{2}\Vert v_k - x^*\Vert^2
            \\&\quad 
                + \frac{\alpha_k^2 \mu^2}{2\hat\gamma}\Vert y_k - v_k\Vert^2
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle. 
        \tag{3*}
        \end{align*}
        From the left to the right on the first equality, coefficients of $\Vert g_k\Vert^2$ cancels out to zero and the inner product containing $g_k$ are grouped. 
        Going from the left to the right of the third equality, we applied (Q3) derived earlier to simplify the inner product term. 
        The last equalities re-arranged terms and grouped the coefficients of $\epsilon_k$ together. 
        Taking a page break, continuing on (3*) we have
        \begin{align*}
            &
            F(x_k) - \alpha_k(\epsilon_k + \langle g_k, x_k - x^*\rangle)
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\&\quad 
                + \frac{(1 - \alpha_k)\mu\alpha_k\gamma}{2\hat \gamma}\Vert v_k - y_k\Vert^2
                + \frac{\mu \alpha_k}{2}\Vert v_k - x^*\Vert^2
            \\&\quad 
                + \frac{\alpha_k^2 \mu^2}{2\hat \gamma}\Vert y_k - v_k\Vert^2
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle
            \\
            &= 
            F(x_k) - \alpha_k(\epsilon_k + \langle g_k, x_k - x^*\rangle)
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\ &\quad 
                + 
                \left(
                    \frac{(1 - \alpha_k)\mu\alpha_k\gamma}{2\hat \gamma}
                    + 
                    \frac{\alpha_k^2 \mu^2}{2\hat \gamma}
                \right)\Vert y_k - v_k\Vert^2
                + \frac{\mu \alpha_k}{2}\Vert v_k - x^*\Vert^2 
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle
            \\
            & =
            F(x_k) - \alpha_k(\epsilon_k + \langle g_k, x_k - x^*\rangle)
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\ &\quad 
                + 
                \frac{\mu \alpha_k}{2}\Vert y_k - v_k\Vert^2
                + \frac{\mu \alpha_k}{2}\Vert v_k - x^*\Vert^2 
                + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle
            \\ &=
            F(x_k) - \alpha_k(\epsilon_k + \langle g_k, x_k - x^*\rangle)
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\ &\quad 
                +
                \frac{\mu\alpha_k}{2} \Vert y_k - x^*\Vert^2
            \\&= 
            F(x_k) - \alpha_k\left(
                \epsilon_k + \langle g_k, x_k - x^*\rangle
                - \frac{\mu}{2}\Vert y_k - x^*\Vert^2
            \right)
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right). 
            \tag{3.1*}
        \end{align*}
        On the first equality, coefficients of $\Vert y_k - v_k\Vert^2$ are grouped together. 
        On the second to the third equality, the terms had been simplified by 
        \begin{align*}
            \frac{(1 - \alpha_k)\mu\alpha_k\gamma}{2\hat \gamma} + 
            \frac{\alpha_k^2 \mu^2}{2\hat \gamma}
            &= 
            \frac{\mu\alpha_k}{2}\left(
                \frac{(1 - \alpha_k)\gamma_k + \alpha_k \mu}{\hat \gamma}
            \right)
            \\
            &= \frac{\mu\alpha_k}{2}\left(
                \frac{\hat \gamma}{\hat \gamma}
            \right) = \frac{\mu\alpha_k}{2}. 
        \end{align*}
        (3.1*) was adding (1*) and (2.2*) together, which is the same as adding (1*) and (2*) together. 
        So that was all equal to (1*) + (2*) and it says: 
        \begin{align*}
            & F(x_{k + 1}) + R_{k + 1} + 
            \frac{\hat \gamma}{2}\Vert v_{k + 1} - x^*\Vert^2
            \\
            &= 
            F(x_k) - \alpha_k\left(
                \epsilon_k + \langle g_k, x_k - x^*\rangle
                - \frac{\mu}{2}\Vert y_k - x^*\Vert^2
            \right)
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\
            & \iff 
            \\
            & F(x_{k + 1}) - F(x^*) + R_{k + 1} + 
            \frac{\hat \gamma}{2}\Vert v_{k + 1} - x^*\Vert^2
            \\
            &= 
            F(x_k) - F(x^*) - \alpha_k\left(
                \epsilon_k + \langle g_k, x_k - x^*\rangle
                - \frac{\mu}{2}\Vert y_k - x^*\Vert^2
            \right)
            + 
            (1 - \alpha_k)\left(
                R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\
            &= (1 - \alpha_k)(F(x_k) - F(x^*))
            + \alpha_k\left(
                F(x_k) - F(x^*) - \epsilon_k - \langle g_k, x_k - x^*\rangle + \frac{\mu}{2}\Vert y_k - x^*\Vert^2
            \right)
            \\ &\quad 
                + 
                (1 - \alpha_k)\left(
                    R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
                \right). 
        \tag{3.2*}
        \end{align*}
        Focusing on the second term, we simplify the multiplier inside: 
        {\small
        \begin{align*}
            & F(x_k) - F(x^*) - \epsilon_k - \langle g_k, x_k - x^*\rangle + \frac{\mu}{2}\Vert y_k - x^*\Vert^2
            \\
            &= 
            F(x_k) - F(x^*) - \left(
                F(x_k) - F(T_L y_k) - \langle g_k, x_k - y_k\rangle - \frac{1}{2L}\Vert g_k\Vert^2
            \right)- \langle g_k, x_k - x^*\rangle + \frac{\mu}{2}\Vert y_k - x^*\Vert^2
            \\
            &= F(T_L y_k) - F(x^*) + \langle g_k, x^* - y_k\rangle + \frac{\mu}{2}\Vert y_k - x^*\Vert^2
            + \frac{1}{2L}\Vert g_k\Vert^2 \le 0. 
        \tag{4*}    
        \end{align*}
        }
        On the last line, we expand the definition of $g_k$ and then used the
        \hyperref[thm:prox-grad-ineq]{Theorem \ref*{thm:prox-grad-ineq}}. 
        Therefore, we conclude that: 
        {\small
        \begin{align*}
            F(x_{k + 1}) - F(x^*) + R_{k + 1} + 
            \frac{\hat \gamma}{2}\Vert v_{k + 1} - x^*\Vert^2
            &\le 
            (1 - \alpha_k)\left(
                F(x_k) - F(x^*) + R_k + \frac{\gamma}{2}\Vert v_k - x^*\Vert^2
            \right). 
        \end{align*}
        }
        Adding back the $k$, $k + 1$ subscript for $\gamma$, $\hat \gamma$, it's proven. 
    \end{proof}
    \begin{remark}
        \textcolor{red}
        {
        Given $y_k$, we may choose to increase $\mu_\kappa = 2D_f(x^*, y_k)/\Vert y_k - x^*\Vert^2 \ge \mu$ which also works on (4*), $\mu$ is a pessimistic choice for the inequality above. 
        }
        But in general the choice of $\mu$ remains the strong convexity modulus or equivalently, any value that is smaller than the true strong convexity constant for claiming the convergence rate for all initial guesses. 
    \end{remark}

\section{Formulation of R-WAPG and its convergence rates}
    The sole focus of this section is Definition \ref{def:wapg}.
    It describes the R-WAPG algorithm which generates iterates $(x_k, y_k, v_k)$ and admits potential convergence claim using Proposition \ref{prop:stepwise-lyapunov}. 
    Definition \ref{def:rwapg-seq} introduces the concept of an R-WAPG sequences which is crucially important. 
    The sequence formulates the R-WAPG algorithm stated in Definition \ref{def:wapg}, it connects the step-wise formulation of R-WAPG (Definition \ref{def:stepwise-wapg}) and describes the convergence claim of R-WAPG in Proposition \ref{prop:wagp-convergence}. 
    In the next section, it continues to play a crucial role in describing several equivalent forms of the R-WAPG algorithm, and their convergence claim. 
    \begin{definition}[R-WAPG sequences]\label{def:rwapg-seq}\;\\
        Assume $0 \le \mu < L$. 
        The sequences $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge1}$ are sequences parameterized by $\mu, L$. 
        They are valid for R-WAPG if all the following holds: 
        \begin{align*}
            \alpha_0 &\in (0, 1], 
            \\
            \alpha_k &\in (\mu/L, 1) \quad (\forall k \ge 1), 
            \\
            \rho_k &:= \frac{\alpha_{k + 1}^2 - (\mu/L)\alpha_{k + 1}}{(1 - \alpha_{k + 1})\alpha_k^2} \quad \forall (k \ge 0). 
        \end{align*}
        We call $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ the \textbf{R-WAPG Sequences}. 
    \end{definition}
    \begin{observation}\label{obs:r-wapg-observation-1}
        The following is true: 
        \begin{enumerate}
            \item $\rho_k > 0$ for all $k \ge 0$. 
            \item $\forall k \ge 1: L\alpha_k^2 = (1 - \alpha_k)L\rho_{k - 1}\alpha_{k - 1}^2 +\mu\alpha_k$. 
        \end{enumerate}

        % If $\alpha_1 \in (0, 1)$ which is true by $\gamma_1 \in (0, L]$, then the entire sequence $\alpha_k \in (0, 1)$. 
        % Suppose inductively that $\alpha_{k - 1}, \rho_{k - 1}$ are given such that they satisfy $\alpha_{k -1} \in (0, 1)$ and $0 < \rho_{k - 1} \alpha_{k - 1}^2 < 1$. 
        % Solving the quadratic $L\alpha_k^2=(1 - \alpha_k)\rho_{k - 1}L\alpha_{k - 1}^2 + \mu \alpha_k$ for $\alpha_k$ yields candidates. 
        % \begin{align*}
        %     \alpha_k &= 
        %     \frac{1}{2}\left(
        %         \frac{\mu}{L} - \rho_{k - 1}\alpha_{k - 1}^2 
        %         \pm
        %         \sqrt{(\rho_{k - 1}\alpha_{k - 1}^2 - \mu/L)^2 + 4\rho_{k - 1}\alpha_{k - 1}^2}
        %     \right). 
        % \end{align*}
        % Choosing the positive sign which is the larger one of the roots of the quadratic. 
        % We have $\alpha_k > 0$ because: 
        % \begin{align*}
        %     \alpha_k &=
        %     \frac{1}{2}\left(
        %         \frac{\mu}{L} - \rho_{k - 1}\alpha_{k - 1}^2 
        %         +
        %         \sqrt{(\rho_{k - 1}\alpha_{k - 1}^2 - \mu/L)^2 + 4\rho_{k - 1}\alpha_{k - 1}^2}
        %     \right)
        %     \\
        %     &\ge 
        %     \frac{1}{2}\left(
        %         \frac{\mu}{L} - \rho_{k - 1}\alpha_{k - 1}^2 
        %         +
        %         \left|\rho_{k - 1}\alpha_{k - 1}^2 - \mu/L\right|
        %     \right) + \alpha_{k - 1}\sqrt{\rho_{k - 1}}
        %     \\
        %     & > 0. 
        % \end{align*}
        % On the last strict inequality we used the fact that $\rho_{k - 1}> 0, \alpha_{k - 1} > 0$. 
        % An upper bound can be identified by using inductive hypothesis and considering: 
        % \begin{align*}
        %     \alpha_k &= 
        %     \frac{1}{2}\left(
        %         \frac{\mu}{L} - \rho_{k - 1}\alpha_{k - 1}^2 
        %         +
        %         \sqrt{(\rho_{k - 1}\alpha_{k - 1}^2 - \mu/L)^2 + 4\rho_{k - 1}\alpha_{k - 1}^2}
        %     \right)
        %     \\
        %     &<
        %     \frac{1}{2}\left(
        %         \frac{\mu}{L} + 
        %         \sup_{x \in (0, 1)}
        %         \left\lbrace
        %             -x + \sqrt{(x - \mu/L)^2 + 4x}
        %         \right\rbrace
        %     \right)
        %     \\
        %     & \le \frac{1}{2}\left(
        %         \mu/L + \max\left(\mu/L, -1 + \sqrt{(1 - \mu/L)^2 + 4}\right)
        %     \right) \le 1. 
        % \end{align*}
        % Going to the first inequality, we used $\rho_{k - 1} \alpha_{k - 1}^2 < 1$ to get the strict inequality. 
        % Going from the second to the third inequality, we maximized $\mu/L$ by monotone increasing of linear function and the square root function. 
        % Therefore, inductively, $\alpha_k \in (0, 1)$ holds. 
        % However, the limit of $\alpha_k$ could be 1. 
    \end{observation}
    \begin{definition}[Relaxed weak accelerated proximal gradient (R-WAPG)]\label{def:wapg}
        \; \\
        Choose any $x_1 \in \RR^n, v_1 \in \RR^n$. 
        Let $(\alpha_k)_{k \ge0}, (\rho_k)_{k \ge 0}$ be given by Definition \ref{def:rwapg-seq}. 
        The algorithm generates a sequence of vector $(y_k, x_{k + 1}, v_{k + 1})_{k \ge 1}$ for $k\ge 1$ by the procedures:  
        \begin{tcolorbox}
            For $k=1, 2, 3, \cdots$
            \begin{align*}
                \gamma_k &:= \rho_{k -1}L\alpha_{k - 1}^2, 
                \\
                \hat \gamma_{k + 1} & := (1 - \alpha_k)\gamma_k + \mu \alpha_k = L\alpha_k^2, 
                \\
                y_k &= 
                (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k), 
                \\
                g_k &= \mathcal G_L y_k, 
                \\
                v_{k + 1} &= 
                \hat\gamma^{-1}_{k + 1}
                (\gamma_k(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k), 
                \\
                x_{k + 1} &= T_L y_k. 
            \end{align*}    
        \end{tcolorbox}
    \end{definition}
    \begin{observation}\label{obs:r-wapg-observation-2}
        Observe that if $\rho_k = 1$ for all $k\ge 0$, then the above algorithm is similar to (2.2.7) in Nesterov's book \cite{nesterov_lectures_2018} because $L\alpha_k^{2} = (1 - \alpha_k)L\alpha_{k - 1}^2 + \mu \alpha_k$. 
        \par
        For all $k \ge 1$, the algorithm chains together the sequence of $(\hat \gamma_{k+1})_{k \ge 1}$ and $(\gamma_k)_{k \ge 1}$ in Definition \ref{prop:stepwise-lyapunov} through the equalities: 
        \begin{align*}
            \hat \gamma_{k + 1} 
            &= L\alpha_k^2 
            \\
            &= (1 - \alpha_k)\gamma_k + \alpha_k \mu
            \\
            &= (1 - \alpha_k)L\rho_{k - 1}\alpha_{k - 1}^2 + \alpha_k \mu
            \\
            &= (1 - \alpha_k)\rho_{k - 1}\hat \gamma_{k} + \alpha_k \mu. 
        \end{align*}
        The third equality from Observation \ref{obs:r-wapg-observation-1}. 
    \end{observation}

    \begin{proposition}[R-WAPG convergence claim]\label{prop:wagp-convergence}\; \\
        Fix any arbitrary $x^* \in \RR^n, N \in \mathbb N$. 
        Let vector sequence $(y_k, v_{k}, x_{k})_{k \ge 1}$ and scalar sequences $\alpha_k, \rho_k$ be generated from Definition \ref{def:wapg}. 
        Define $R_1 = 0$ and suppose that for $k = 1, 2, \cdots, N$, we have $R_k$ recursively given by: 
        \begin{align*}
            R_{k + 1}
            := 
            \frac{1}{2}\left(
                L^{-1} - \frac{\alpha_k^2}{\hat \gamma_{k + 1}}
            \right)\Vert g_k\Vert^2
            + 
            (1 - \alpha_k)
            \left(
                \epsilon_k + R_k + 
                \frac{\mu\alpha_k\gamma_k}{2\hat \gamma_{k + 1}}
                \Vert v_k - y_k\Vert^2
            \right). 
        \end{align*} 
        Then for all $k = 1, 2, \cdots, N$: 
        \begin{align*}
            & F(x_{k + 1}) - F(x^*) + \frac{L \alpha_k^2}{2}\Vert v_{k + 1} - x^*\Vert^2
            \\
            &\le 
            \left(
                \prod_{i = 0}^{k - 1} \max(1, \rho_{k})
            \right)
            \left(
                \prod_{i = 1}^{k} \left(1  - \alpha_i\right)
            \right)
            \left(
                F(x_1) - F(x^*) + \frac{\gamma_1}{2}\Vert v_1 - x^*\Vert^2
            \right). 
        \end{align*}
    \end{proposition}
    \begin{proof}  
        $\hat \gamma_2 = (1 - \alpha_1)\gamma_1 + \mu \alpha_1$ is true by Observation \ref{obs:r-wapg-observation-1}. 
        Observe that $R_2$ is defined in the same manner as in Proposition \ref{prop:stepwise-lyapunov} with k = 1 so we can invoke Proposition \ref{prop:stepwise-lyapunov}, and it gives: 
        \begin{align*}
            & F(x_{2}) - F(x^*) + R_2 + \frac{L \alpha_1^2}{2}\Vert v_{2} - x^*\Vert^2
            \\
            & = F(x_{2}) - F(x^*) + R_2 + \frac{\hat \gamma_{2}}{2}\Vert v_{2} - x^*\Vert^2
            & \hat \gamma_2 = L \alpha_1^2 , \text{ Definition \ref{def:wapg}}
            \\
            &\le 
            (1 - \alpha_1)\left(
                F(x_1) - F(x^*) + R_1 + \frac{\gamma_1}{2}\Vert v_1 - x^*\Vert^2
            \right) 
            & \quad \text{By Proposition \ref{prop:stepwise-lyapunov}}
            \\
            &= 
            (1 - \alpha_1)\left(
                F(x_1) - F(x^*) + R_1 + \frac{L\rho_{0}\alpha_{0}^2}{2}\Vert v_1 - x^*\Vert^2
            \right) 
            \\
            &= \max\left(\rho_0, 1\right)
            (1 - \alpha_1)\left(
                F(x_1) - F(x^*) + R_1 + \frac{L\alpha_0^2}{2}\Vert v_1 - x^*\Vert^2
            \right). 
        \end{align*}
        We had demonstrated the base case. 
        \par 
        Next, for all $k = 2, 3, \cdots, N$, from Observation \ref{obs:r-wapg-observation-2}: 
        \begin{align*}
            \hat \gamma_{k + 1} = L\alpha_{k}^2 
            &=(1 - \alpha_k)\rho_{k - 1}L\alpha_{k - 1}^2 + \mu\alpha_k
            \\
            &= (1 - \alpha_k)\gamma_k + \mu\alpha_k. 
        \end{align*}
        So $\gamma_k, \hat \gamma_{k + 1}, \alpha_k$ fits Definition \ref{def:stepwise-wapg}. 
        The definition of $R_k$ given is the same as in Proposition \ref{prop:stepwise-lyapunov} hence we can use Proposition \ref{prop:stepwise-lyapunov} to unroll recursively for all $k \ge 1$: 
        {\small
        \begin{align*}
            &
            F(x_{k + 1}) - F^* + R_{k + 1} + \frac{\hat\gamma_{k + 1}}{2}\Vert v_{k + 1} - x^*\Vert^2 
            \\
            &=
            F(x_{k + 1}) - F^* + R_{k + 1} + \frac{L \alpha_k^2}{2}\Vert v_{k + 1} - x^*\Vert^2
            \\
            &\le 
            (1 - \alpha_k)
            \left(
                F(x_k) - F^* + R_k + \frac{\rho_{k - 1}L \alpha_{k - 1}^2}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\
            &\le 
            (1 - \alpha_k)
            \left(
                F(x_k) - F^* + R_k + \max(1, \rho_{k - 1})\frac{L \alpha_{k - 1}^2}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\
            &\le 
            \max(1, \rho_{k - 1})(1 - \alpha_k)
            \left(
                F(x_k) - F^* + R_k + \frac{L \alpha_{k - 1}^2}{2}\Vert v_k - x^*\Vert^2
            \right)
            \\
            &\le 
            \left(
                \prod_{i = 0}^{k - 1} \max(1, \rho_{k})
            \right)
            \left(
                \prod_{i = 1}^{k} \left(1  - \alpha_i\right)
            \right)
            \left(
                F(x_1) - F^* + R_1 + \frac{L\alpha_0^2}{2}\Vert v_1 - x^*\Vert^2
            \right). 
        \end{align*}
        }
        We had demonstrated in the inductive case for $k=1, 2, \cdots, N$. 
        Additionally, for all $k = 1, 2, \cdots, N$ it has $R_{k + 1} \ge 0$ because: 
        \begin{align*}
            R_{k + 1}
            &= 
            \frac{1}{2}\left(
                L^{-1} - \frac{\alpha_k^2}{\hat \gamma_{k + 1}}
            \right)\Vert g_k\Vert^2
            + 
            (1 - \alpha_k)
            \left(
                \epsilon_k + R_k + 
                \frac{\mu\alpha_k\gamma_k}{2\hat \gamma_{k + 1}}
                \Vert v_k - y_k\Vert^2
            \right)
            \\
            &= (1 - \alpha_k)
            \left(
                \epsilon_k + R_k 
                + \frac{\mu\alpha_k\gamma_k}{2\hat \gamma_{k + 1}}
                \Vert v_k - y_k\Vert^2
            \right)
            \\
            &\ge 
            (1 - \alpha_k) R_k
            \\
            &\ge R_1 \prod_{i = 1}^{k} \left(1 - \alpha_i\right) = 0. 
        \end{align*}
        Going from the left to the right on the first equality, we used the fact that $\hat \gamma_{k + 1} = L \alpha_{k}^2$.
        This makes coefficient of $\Vert g_k\Vert^2$ zero. 
        The first inequality is by $\epsilon_k \ge 0$ and the non-negativity of the remaining terms. 
        The last equality is by the assumption that $R_1 = 0$. 
        Therefore: 
        {\small
        \begin{align*}
            & 
            F(x_{k + 1}) - F^* +
            \frac{L\alpha_k^2}{2}\Vert v_{k + 1} - x^*\Vert^2
            \\
            &\le 
            \left(
                \prod_{i = 0}^{k - 1} \max(1, \rho_{k})
            \right)
            \left(
                \prod_{i = 1}^{k} \left(1  - \alpha_i\right)
            \right)
            \left(
                F(x_1) - F^* + \frac{L\alpha_0^2}{2}\Vert v_1 - x^*\Vert^2
            \right). 
        \end{align*}
        }
    \end{proof}
    \begin{remark}        
        \textcolor{red}
        {
            The choice of $\rho_k$ allows for nontraditional choices of step sizes in the literatures. 
            As we can see from the above convergence claim, there exists choices of $\rho_k\ge 1$ where convergence remains possible. 
        }
    \end{remark}

\section{Equivalent representations of R-WAPG}
    This section reduces Definition \ref{def:wapg} into simpler forms that are comparable to what commonly appears in the literatures. 
    In the literatures, variants of Accelerated Proximal Gradient algorithm such as FISTA, V-FISTA has different representations. 
    This section carries out the tedious but necessary works for demonstrating the descriptive power of R-WAPG. 
    We will formulate it into three generic forms that are commonly seemed in the literatures. 
    These equivalent representations are listed in Definition \ref{def:r-wapg-intermediate}, \ref{def:r-wapg-st-form} and \ref{def:r-wapg-momentum-form}. 
    \par 
    Proposition \ref{prop:wapg-first-equivalent-repr} simplifies Definition \ref{def:wapg} and finds a representation without using auxiliary sequence $\gamma_k, \hat \gamma_k$. 
    Definition \ref{def:r-wapg-intermediate} states the first simplified form of the R-WAPG algorithm which we call: ``R-WAPG intermediate form''. 
    Following a similar pattern, Proposition \ref{prop:wagp-st-form}, \ref{prop:r-wapg-momentum-repr} demonstrates two more equivalent representations of the R-WAPG intermediate form (Definition \ref{def:r-wapg-intermediate}) which are formulated into Definition \ref{def:r-wapg-st-form}, \ref{def:r-wapg-momentum-form}. 
    Convergence results from Proposition \ref{prop:wagp-convergence} applies to all these equivalent forms of R-WAPG.

    To start, the following proposition on ``abstract similar triangle form'' were made to simplify arguments and notations to make nicer proofs.

    \begin{proposition}[Abstract similar triangle form]\label{prop:abs-st-form}\;\\
        Given any initial $(x_1, v_1)$, and a sequence $(\tau_k)_{k \ge 1}, (\xi_k)_{k \ge 1}$ such that for all $k \ge 1$ $\tau_k \in (0, 1), \xi_k \in (0, 1)$. 
        For all $k\ge 1$, iterates $(y_k, z_{k + 1}, x_{k + 1})_{k \ge 1}$ satisfies recursively that: 
        \begin{align*}
            y_k &= (1 + \tau_k)^{-1}(v_k + \tau_k x_k),
            \\
            v_{k + 1} &= (1 + \xi_k)^{-1}(v_k + \xi_k y_k) - (1 + \xi_k)^{-1}\delta_k g_k,
            \\
            x_{k + 1} &= y_k - L^{-1} g_k. 
        \end{align*}
        If $1 + \xi_k + \tau_k = L\delta_k, \tau_k \neq 0\; \forall k \ge 1$. 
        Then for all $k \ge 1$: 
        $$
            v_{k + 1} - x_{k + 1} = (1 + \xi_k)^{-1}\tau_k(x_{k + 1} - x_k). 
        $$
        Which makes the algorithm a similar triangle form. 
    \end{proposition}
    \begin{proof}
        We are interested in identifying the conditions required for the sequence of $\xi_k, \tau_k, \delta_k$ such that there exists $\theta_k$ satisfying: 
        \begin{align*}
            v_{k + 1} - x_{k + 1} 
            &= \theta_k(x_{k + 1} - x_k).
        \end{align*}
        To verify, do 
        \begin{align*}
            v_{k + 1} &= 
            (1 + \xi_k)^{-1}(v_k + \xi_t y_k - \delta_k \mathcal G_L(y_k))
            \\
            &= 
            (1 + \xi_k)^{-1}((1 + \tau_k)y_k - \tau_t x_k + \xi_k y_k - \delta_k \mathcal G_L(y_k))
            \\
            &= 
            (1 + \xi_k)^{-1}((1 + \tau_k + \xi_k)y_k - \tau_k x_k - \delta_k \mathcal G_L(y_k))
            \\
            v_{k + 1} - x_{k + 1}
            &= 
            (1 + \xi_k)^{-1}((1 + \tau_k + \xi_k)y_t - \tau_k x_k - \delta_t \mathcal G_L(y_k))
            - (y_k - L^{-1}\mathcal G_Ly_k)
            \\
            &= 
            (1 + \xi_k)^{-1}(\tau_ky_k - \tau_k x_k - \delta_k \mathcal G_L(y_k))
            + L^{-1}\mathcal G_Ly_k
            \\
            &= 
            (1 + \xi_k)^{-1}
            \left(
                \tau_ty_k - \tau_t x_k + (L^{-1}(1 + \xi_k) - \delta_k) \mathcal G_L(y_k)
            \right)
            \\
            &= 
            (1 + \xi_k)^{-1}\tau_k
            \left(
                y_k - x_k + 
                \tau_k^{-1}(L^{-1}(1 + \xi_k) - \delta_k) \mathcal G_L(y_k)
            \right).
        \end{align*}
        Going between the first and second inequality we used $v_k = (1 + \tau_k)y_k - \tau_k x_k$ which is rearranged $y_k = (1 + \tau_k)^{-1}(v_k + \tau_k x_k)$. 
        The RHS is can be verified through 
        \begin{align*}
            x_{k + 1} - x_k &= 
            y_t - L^{-1}\mathcal G_L(y_k) - x_k
            \\
            &= (y_k - x_k) - L^{-1}\mathcal G_L(y_k). 
        \end{align*}
        It necessitates the condition: 
        \begin{align*}
            \tau_k^{-1}(L^{-1}(1 + \xi_k) - \delta_k) 
            &= - L^{-1}
            \\
            (1 + \xi_k) - L\delta_k
            &= 
            - \tau_k
            \\
            1 + \xi_k + \tau_k
            &=
            L\delta_k. 
        \end{align*}
        Which allows for: 
        \begin{align*}
            v_{k + 1} - x_{k + 1} &= 
            (1 + \xi_k)^{-1}\tau_t
            \left(y_k - x_k - L^{-1}\mathcal G_L(y_k)\right) 
            = 
            (1 + \xi_k)^{-1}\tau_k(x_{k + 1} - x_k). 
        \end{align*}
    \end{proof}
    \subsection{Listing the euivalent representations of R-WAPG}
        This section lists three equivalent representations of the R-WAPG algorithm. 
        They are comparable to existing APG algorithms in the literatures. 
        \begin{definition}[R-WAPG intermediate form]\label{def:r-wapg-intermediate}\;\\
            Assume $\mu < L$ and let $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ given by Definition \ref{def:rwapg-seq}. 
            Initialize any $x_1, v_1$ in $\RR^n$. 
            For $k \ge 1$, the algorithm generates sequence of vector iterates $(y_{k}, v_{k + 1}, x_{k + 1})_{k \ge 1}$ by the procedures: 
            \begin{tcolorbox}
                For $k = 1, 2, \cdots$
                \begin{align*} 
                    & y_{k} = 
                    \left(
                        1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                    \right)^{-1}
                    \left(
                        v_{k + 1} + 
                        \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                    \right), 
                    \\
                    & x_{k + 1} = 
                    y_k - L^{-1} \mathcal G_L y_k, 
                    \\
                    & v_{k + 1} = 
                    \left(
                        1 + \frac{\mu}{L \alpha_k - \mu}
                    \right)^{-1}
                    \left(
                        v_k + 
                        \left(\frac{\mu}{L \alpha_k - \mu}\right) y_k
                    \right) - \frac{1}{L\alpha_{k}}\mathcal G_L y_k. 
                \end{align*}
            \end{tcolorbox}
        \end{definition}
        \begin{definition}[R-WAPG similar triangle form]\label{def:r-wapg-st-form} \; \\
            Given any $(x_1, v_1)$ in $\RR^n$. 
            Assume $\mu < L$.
            Let the sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k\ge 0}$ be given by Definition \ref{def:rwapg-seq}. 
            For $k \ge 1$, the algorithm generates sequences of vector iterates $(y_k, v_{k + 1}, x_{k + 1})_{k \ge 1}$ by the procedures: 
            \begin{tcolorbox}
                For $k=1, 2, \cdots $
                \begin{align*}
                    & y_k = 
                    \left(
                        1 + \frac{L - L\alpha_k}{L\alpha_k - \mu}
                    \right)^{-1}
                    \left(
                        v_k + 
                        \left(\frac{L - L\alpha_k}{L\alpha_k - \mu} \right) x_k
                    \right), 
                    \\
                    & x_{k + 1} = 
                    y_k - L^{-1} \mathcal G_L y_k, 
                    \\
                    & v_{k + 1} = 
                    x_{k + 1} + (\alpha_k^{-1} -1)(x_{k + 1} - x_k). 
                \end{align*}    
            \end{tcolorbox}
        \end{definition}
        \begin{definition}[R-WAPG momentum form]\label{def:r-wapg-momentum-form}
            For any $y_1 = x_1 \in \RR^n$, $(\rho_k)_{k \ge 0}, (\alpha_k)_{k\ge 0}$ given by Definition \ref{def:rwapg-seq}. 
            The algorithm generates iterates $x_{k + 1}, y_{k + 1}$ For $k = 1, 2, \cdots $ by the procedures: 
            \begin{tcolorbox}
                For $k=1, 2,\cdots $
                \begin{align*}
                    & x_{k + 1} = y_k - L^{-1}\mathcal G_Ly_k, 
                    \\
                    & 
                    y_{k + 1} = 
                    x_{k + 1} + 
                    \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}(x_{k + 1} - x_k). 
                \end{align*}    
            \end{tcolorbox}

            In the special case where $\mu = 0$, the momentum term can be represented without relaxation parameter $\rho_k$: 
            $$
                (\forall k \ge 1)\quad \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}} 
                = \alpha_{k + 1}(\alpha_k^{-1} - 1).  
            $$
        \end{definition}
        \begin{remark}
           \todo[inline]{Consider adding a remark here on how this is similar in format to the Nesterov's accelerated gradient (2.2.19) in his book. }
        \end{remark}

    \subsection{Proving the equivalent representations of R-WAPG}
        \begin{proposition}[First equivalent representation of R-WAPG]\label{prop:wapg-first-equivalent-repr}\;\\
            If the sequence $(y_k, v_k, x_k)_{k \ge 1}$ is produced by R-WAPG (Definition \ref{def:wapg}). 
            Then the iterates can be expressed without $(\gamma_k)_{k \ge1},(\hat \gamma_k)_{k \ge 2}$, and for all $k\ge 1$ they are algebraically equivalent to
            \begin{align*}
                & 
                y_{k} = 
                \left(
                    1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                \right)^{-1}
                \left(
                    v_{k} + 
                    \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                \right), 
                \\
                & x_{k + 1} = 
                y_k - L^{-1} \mathcal G_L y_k, 
                \\
                & v_{k + 1} = 
                \left(
                    1 + \frac{\mu}{L \alpha_k - \mu}
                \right)^{-1}
                \left(
                    v_k + 
                    \left(\frac{\mu}{L \alpha_k - \mu}\right) y_k
                \right) - \frac{1}{L\alpha_{k}}\mathcal G_L y_k. 
            \end{align*}
        \end{proposition}
        \begin{proof}
            For all $k \ge 1$, by R-WAPG (Definition \ref{def:wapg}), it has: 
            \begin{align*}
                y_{k} &= 
                (\gamma_k + \alpha_k \mu)^{-1}
                (\alpha_k \gamma_k v_k + \hat \gamma_{k + 1}x_k)
                \\
                &= 
                (\hat \gamma_{k + 1} + \alpha_k \gamma_k)^{-1}
                (\alpha_k \gamma_k v_k + \hat \gamma_{k + 1}x_k)
                \\
                &= 
                \left(
                    \frac{\hat \gamma_{k + 1}}{\alpha_k\gamma_k} + 1
                \right)^{-1}
                \left(
                    v_k + \frac{\hat \gamma_{k + 1}}{\alpha_k \gamma_k} x_k
                \right)
                \\
                &= 
                \left(
                    \frac{L\alpha_k^2}{\alpha_k\gamma_k} + 1
                \right)^{-1}
                \left(
                    v_k + \frac{L\alpha_k^2}{\alpha_k \gamma_k} x_k
                \right)
                \\
                &= 
                \left(
                    \frac{L\alpha_k}{\gamma_k} + 1
                \right)^{-1}
                \left(
                    v_k + \frac{L\alpha_k}{ \gamma_k} x_k
                \right)
                \\
                &= 
                \left(
                    1 + \frac{L - L \alpha_k}{L \alpha_k - \mu}
                \right)^{-1}
                \left(
                    v_k + 
                    \frac{L - L \alpha_k}{L \alpha_k - \mu} x_k
                \right). 
            \end{align*}
            From the left to right of the second inequality, we used the fact that $\hat \gamma_{k + 1} = (1 - \alpha_k)\gamma_k + \alpha_k\mu$. 
            Going from the left to the right of the second last inequality, we did the following: 
            \begin{align*}
                L\alpha_k^2 &= 
                (1 - \alpha_k)\gamma_k + \alpha_k \mu 
                \\
                \iff 
                L \alpha_k^2 - \alpha_k\mu &= 
                (1 - \alpha_k)\gamma_k
                \\
                \iff 
                \gamma_k/L
                &= 
                \frac{L \alpha_k^2 - \alpha_k\mu}{L (1 - \alpha_k)}
                \\
                \iff 
                L/\gamma_k
                &= 
                \frac{L (1 - \alpha_k)}{L \alpha_k^2 - \alpha_k\mu}
                \\
                \iff 
                L\alpha_k/\gamma_k
                &= 
                \frac{L - L\alpha_k}{L\alpha_k - \mu}. 
            \end{align*}
            On the third $\iff$, we can assume $\alpha_k \neq \mu/L\;  \forall k \ge 1$ because from Definition \ref{def:rwapg-seq}: $\alpha_k \in (\mu/L, 1)$ for all $k \ge 1$. 
            % On the third $\iff$, we can assume $\alpha_k \neq \mu/L\; \forall k \ge 2$ because if $\alpha_k = \mu/L$ were impossible, it then has
            % \begin{align*}
            %     L \alpha_k^2 - \alpha_k\mu &= 
            %     (1 - \alpha_k)L \rho_{k - 1}\alpha_{k - 1}^2 = 0 \implies \alpha_{k - 1} = 0. 
            % \end{align*}
            % Which contradict $\alpha_{k - 1} \in (0, 1)\; \forall k \ge 2$ because $\alpha_1 \in (0, 1)$ is given to satisfy the condition in the base case, and $\alpha_k \in (0, 1)\; \forall k \ge 2$ by justifications in Observation \ref{obs:r-wapg-observation-1}. 
            \\
            For all $k \ge 1$, $v_{k + 1}$ has: 
            \begin{align*}
                v_{k + 1} &= 
                \hat \gamma_{k + 1}^{-1}
                ((1 - \alpha_k)\gamma_k v_k + \mu\alpha_k y_k)
                - \alpha_k\hat \gamma_{k + 1}^{-1}\mathcal G_L y_k
                \\
                &= 
                ((1 - \alpha_k)\gamma_k + \alpha_k \mu)^{-1}
                \left(
                    (1 - \alpha_k)\gamma_k v_k + \mu\alpha_k y_k
                \right)
                - \alpha_k\hat \gamma_{k + 1}^{-1}\mathcal G_L y_k
                \\
                &= 
                \left(
                    1 + \frac{\alpha_k\mu}{(1 - \alpha_k)\gamma_k}
                \right)^{-1}
                \left(
                    v_k + 
                    \frac{\alpha_k\mu}{(1 - \alpha_k)\gamma_k} y_k
                \right)
                - \alpha_k\hat \gamma_{k + 1}^{-1}\mathcal G_L y_k
                % \\
                % &= 
                % \left(
                %     1 + \frac{\alpha_k\mu}{(1 -\alpha_k)L\rho_{k - 1}\alpha_{k - 1}^2}
                % \right)^{-1}
                % \left(
                %     v_k + 
                %     \frac{\alpha_k\mu}{(1 - \alpha_k)\rho_{k - 1}L\alpha_{k - 1}^2} y_k
                % \right)
                % - \frac{1}{L\alpha_{k}}\mathcal G_L y_k
                \\
                &= 
                \left(
                    1 + \frac{\alpha_k \mu}{L \alpha_k^2 - \alpha_k \mu}
                \right)^{-1}
                \left(
                    v_k + 
                    \frac{\alpha_k \mu}{L \alpha_k^2 - \alpha_k \mu} y_k
                \right)
                - \frac{1}{L\alpha_{k}}\mathcal G_L y_k. 
            \end{align*}
            Going from the left to the right of the second equality, we substitute $\hat \gamma_{k + 1} = (1 - \alpha_k)\gamma_k + \mu\alpha_k$. 
            At the end, recall that for all $k \ge 1$, it has $\hat \gamma_{k + 1} = L \alpha_k^2 = (1 - \alpha_k)\gamma_k + \alpha_k \mu$, so: 
            \begin{align*}
                (1 - \alpha_k)\gamma_k
                &= 
                \hat \gamma_{k + 1} - \mu \alpha_k
                = 
                L\alpha_{k}^2 - \alpha_k\mu. 
            \end{align*}
            The is now complete. 
            This form doesn't have $\rho_k, \gamma_k, \hat \gamma_k$ in it. 
        \end{proof}

        \begin{proposition}[Second equivalent representation of R-WAPG]\label{prop:wagp-st-form}\;\\
            Let iterates $(y_k, x_{k}, v_{k})_{k \ge 1}$ and sequence $(\alpha_k, \rho_k)_{k \ge 0}$ be given by Definition \ref{def:r-wapg-intermediate}. 
            Then for all $k \ge 1$, iterate $y_k, x_{k + 1}, v_{k + 1}$
            satisfy: 
            \begin{align*}
                y_{k} &= 
                \left(
                    1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                \right)^{-1}
                \left(
                    v_{k} + 
                    \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                \right), 
                \\
                x_{k + 1} &= 
                y_k - L^{-1} \mathcal G_L y_k, 
                \\
                v_{k + 1} &= x_{k + 1} + (\alpha_k^{-1} - 1)(x_{k + 1} - x_k). 
            \end{align*}
        \end{proposition}
        \begin{proof}
            From Definition \ref{def:r-wapg-intermediate}, define $(\tau_k, \xi_k, \delta_k)_{k \ge 1}$ in Proposition \ref{prop:abs-st-form} to be
            \begin{align*}
                (\forall k \ge 1) \quad \tau_k &= \frac{L(1 - \alpha_k)}{L\alpha_k - \mu},
                \\
                (\forall k \ge 1)\quad 
                \xi_k &= \frac{\mu}{L \alpha_k - \mu},
                \\
                (\forall k \ge 1)\quad 
                \delta_k &\defeq \frac{1 + \xi_k}{L\alpha_k}. 
            \end{align*}
            We claim that these parameters satisfy $L\delta_k = 1 + \tau_k + \xi_k$. 
            To see, we have forall $k\ge 1$: 
            \begin{align*}
                1 + \tau_k + \xi_k &= 
                1 + \frac{L(1 - \alpha_k)}{L \alpha_k - \mu} 
                + \frac{\mu}{L \alpha_k - \mu}
                \\
                &= 
                1 + \frac{L - L \alpha_k + \mu}{L\alpha_k - \mu}
                \\
                &= 
                \frac{L - L \alpha_k + \mu + L \alpha_k - \mu}{L\alpha_k - \mu}
                \\
                &= \frac{L}{L\alpha_k - \mu}. 
            \end{align*}
            Next, it also has for all $k \ge 1$: 
            \begin{align*}
                \frac{1 + \xi_k}{\alpha_k}
                &= 
                \frac{1 + \frac{\mu}{L\alpha_k - \mu}}{\alpha_k}
                = 
                \frac{\frac{L\alpha_k - \mu + \mu}{L \alpha_k - \mu}}{\alpha_k}
                = 
                \frac{L}{L\alpha_k - \mu}.
            \end{align*}
            Hence,  $v_{k + 1} - x_{k + 1} = (1 + \xi_k)^{-1}(x_{k + 1} - x_k)\; \forall k \ge 1$ by Proposition \ref{prop:abs-st-form}.
            Therefore, substituting it gives: 
            \begin{align*}
                v_{k + 1} &= 
                x_{k + 1} + \left(
                    1 + \frac{\mu}{L\alpha_k - \mu}
                \right)^{-1}\left(
                    \frac{L(1 - \alpha_k)}{L\alpha_k - \mu}
                \right)(x_{k + 1} - x_k)
                \\
                &= 
                x_{k + 1} + \left(
                    \frac{L\alpha_k}{L\alpha_k - \mu}
                \right)^{-1}\left(
                    \frac{L(1 - \alpha_k)}{L\alpha_k - \mu}
                \right)(x_{k + 1} - x_k)
                \\
                &= 
                x_{k + 1} + \left(
                    \frac{L\alpha_k - \mu}{L\alpha_k}
                \right)\left(
                    \frac{L - L\alpha_k}{L\alpha_k - \mu}
                \right)(x_{k + 1} - x_k)
                \\
                &= x_{k + 1} + (\alpha_k^{-1} - 1)(x_{k + 1} - x_k). 
            \end{align*}
        \end{proof}

    \begin{proposition}[Third equivalent representation of R-WAPG]\label{prop:r-wapg-momentum-repr}
        \;\\
        Let sequence $(\alpha_k, \rho_k)_{k \ge 0}$ and iterates $(x_k, v_k, y_k)_{k\ge 1}$ given by R-WAPG intermediate form (Definition \ref{def:r-wapg-st-form}). 
        Then for all $k \ge 1$, the iterates $(x_{k + 1}, y_{k + 1})_{k \ge 1}$ are algebraically equivalent to: 
        \begin{align*}
            x_{k + 1} &= y_k - L^{-1}\mathcal G_Ly_k, 
            \\
            y_{k + 1} &= 
            x_{k + 1} + 
            \frac{\rho_k\alpha_k(1 - \alpha_k)}
            {\rho_k\alpha_k^2 + \alpha_{k + 1}}(x_{k + 1} - x_k). 
        \end{align*}
        If in addition, $v_1 = x_1$ then 
        \begin{align*}
            y_1 = \left(
                1 + \frac{L - L \alpha_1}{L\alpha_1 - \mu}
            \right)^{-1}\left(
                v_1 + \left(
                    \frac{L - L \alpha_1}{L \alpha_1 - \mu}
                \right)x_1
            \right) = x_1. 
        \end{align*}
        In the special case when $\mu = 0$, the momentum term admits simpler representation 
        \begin{align*}
            (\forall k \ge 1) \quad 
            \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}
            & = \alpha_{k + 1}(\alpha_k^{-1} - 1). 
        \end{align*}
    \end{proposition}
    \begin{proof}
        Start by considering the update rule for $v_k$ from the Definition \ref{def:r-wapg-st-form} which has for all $k \ge 1$: 
        \begin{align*}
            v_{k + 1} &= 
            x_{k + 1} + (\alpha_k^{-1} - 1)(x_{k + 1} - x_k)
            \\
            \iff 
            (L \alpha_{k + 1} - \mu)v_{k + 1} 
            &= 
            (L \alpha_{k + 1} - \mu)x_{k + 1} + (L\alpha_{k + 1} - \mu)(\alpha_k^{-1} - 1)(x_{k + 1} - x_k). 
        \end{align*}
        Next, we simplify the update $y_{k}$ by Definition \ref{def:r-wapg-st-form} for all $k \ge 1$: 
        \begin{align*}
            y_k &= 
            \left(
                1 + \frac{L - L\alpha_k}{L\alpha_k - \mu}
            \right)^{-1}
            \left(
                v_k + 
                \left(\frac{L - L\alpha_k}{L\alpha_k - \mu} \right) x_k
            \right)
            \\
            &= 
            \left(
            \frac{L - \mu}{L\alpha_k - \mu} 
            \right)^{-1}
            \left(
                v_k + 
                \left(\frac{L - L\alpha_k}{L\alpha_k - \mu} \right) x_k
            \right)
            \\
            &= 
            \frac{L\alpha_k - \mu}{L - \mu} v_k
            + 
            \frac{L - L \alpha_k}{L - \mu} x_k
            \\
            &= (L - \mu)^{-1}((L \alpha_k - \mu)v_k + (L - L \alpha_k)x_k). 
        \end{align*}
        We have $y_1 = x_1$ when $v_1 = x_1$ because substituting $v_1 = x_1$ for how $y_1$ is defined it has: 
        \begin{align*}
            y_1 &= (L - \mu)^{-1}(L(\alpha_1 - \mu)x_1 + (L - L\alpha_1)x_1)
            \\
            &= (L - \mu)^{-1}(L(\alpha_1 - \mu)x_1 + (L - L \alpha_1)x_1)
            \\
            &= (L - \mu)^{-1}((L - \mu)x_1) = x_1. 
        \end{align*}
        Now substituting $v_{k + 1}$ into $y_{k + 1}$ so it has for all $k\ge 1$ : 
        {\small
        \begin{align*}
            y_{k + 1} &= 
            (L - \mu)^{-1}((L\alpha_{k + 1} - \mu)v_{k + 1} + (L - L \alpha_{k + 1})x_{k + 1})
            \\
            &= (L - \mu)^{-1}
            \left(
                (L\alpha_{k + 1} - \mu)x_{k + 1} + 
                (L\alpha_{k + 1} - \mu)(\alpha_k^{-1} - 1)(x_{k + 1} - x_k)
                + (L - L \alpha_{k + 1})x_{k + 1}
            \right)
            \\
            &= 
            (L - \mu)^{-1}
            \left(
                (L\alpha_{k + 1} - \mu)x_{k + 1} + (L\alpha_{k + 1} - \mu)(\alpha_k^{-1} - 1)(x_{k + 1} - x_k)
            \right)
            \\
            &= x_{k + 1} + \frac{(L\alpha_{k + 1} - \mu)(\alpha_k^{-1} - 1)}{L - \mu}(x_{k + 1} - x_k). 
        \end{align*}
        }
        The coefficient for $(x_{k + 1} - x_k)$ needs some works to formulate it without the parameter $\mu, L$. 
        To do that we have: 
        \begin{align*}
            \frac{(L\alpha_{k + 1} - \mu)(\alpha_k^{-1} - 1)}{L - \mu}
            &= \frac{(L\alpha_{k + 1} - \mu)\alpha_k(1 - \alpha_k)}{\alpha_k^2(L - \mu)}
            \\
            &= 
            \alpha_k(1 - \alpha_k)
            \left(
                \frac{\alpha_k^2(L - \mu)}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= \alpha_k(1 - \alpha_k)
            \left(
                \frac{L\alpha_k^2 - \mu\alpha_k^2}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= 
            \alpha_k(1 - \alpha_k)
            \rho_k\left(
                \frac{L\rho_k\alpha_k^2 - \mu\rho_k\alpha_k^2}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= 
            \rho_k\alpha_k(1 - \alpha_k)
            \left(
                \frac{(L\alpha_{k + 1} - \mu)(\rho_k\alpha_k^2 + \alpha_{k + 1})}
                {L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}.
        \end{align*}
        Going from the left to right on the fourth equality, we used $L\alpha_{k + 1}^2 = (1 - \alpha_{k + 1})\rho_kL\alpha_k^2 + \mu \alpha_{k + 1}$ with 
        \begin{align*}
            L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2 
            &= 
            (1 - \alpha_{k + 1})L \rho_k \alpha_k^2 + \alpha_{k + 1} L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= 
            ((1 - \alpha_{k + 1})L \rho_k \alpha_k^2 + \mu \alpha_{k + 1}) - \mu\alpha_{k + 1} + \alpha_{k + 1} L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= L \alpha_{k + 1}^2 - \mu\alpha_{k + 1} + \alpha_{k + 1}L\rho_k\alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= 
            L\alpha_{k + 1}(\alpha_{k + 1} + \rho_k \alpha_k^2) - \alpha_{k + 1}\mu - \mu \rho_k \alpha_k^2
            \\
            &= (L \alpha_{k + 1} - \mu)(\alpha_{k + 1} + \rho_k \alpha_k^2). 
        \end{align*}
        When $\mu = 0$, things simplify. 
        Consider that $\forall k \ge 1: \alpha_{k + 1}^2 = (1 - \alpha_{k + 1})\rho_k\alpha_k^2$. 
        \begin{align*}
            \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}
            &= 
            \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}
            {\rho_k\alpha_{k + 1}\alpha_k^2 + \alpha_{k + 1}^2}
            \\
            &= 
            \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}
            {\rho_k\alpha_{k + 1}\alpha_k^2 + \rho_k(1 - \alpha_{k + 1})\alpha_k^2}
            \\
            &= \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2}
            \\
            &= \alpha_{k + 1}(\alpha_k^{-1} - 1). 
        \end{align*}
    \end{proof}

\section{R-WAPG describes existing accelerations scheme in the literatures}
    In addition to various equivalent forms of the R-WAPG algorithm, the R-WAPG sequences are also descriptive and they generalizes many existing sequences used in accelerated proximal gradient schemes. 
    \par
    This section demonstrates that several variants of FISTA in the literatures reduces to the R-WAPG method by setting up the R-WAPG sequences with additional assumptions. 
    This section will also demonstrates that the convergence claim (Proposition \ref{prop:wagp-convergence}) holds and it derives convergence rates consistent with results in the literatures.  
    \begin{lemma}[R-WAPG sequences as Inverted FISTA sequence]
        Let R-WAPG sequence $(\rho_k)_{k \ge 0}, (\alpha_k)_{k \ge 0}$ given by Definition \ref{def:rwapg-seq}. 
        If $\mu = 0, \rho_k \ge 1\; \forall k \ge 0$, and $\alpha_0 = 1$, then: 
        \begin{enumerate}
            \item $\alpha_k^{-2} \ge \alpha_{k + 1}^{-2} - \alpha_{k + 1}^{-1}\; \forall k \ge 0$
            \item $\alpha_k \ge \alpha_{k + 1}; \forall k \ge 0$
            \item Let $t_k := \alpha_k^{-1}$, then $0 < t_{k + 1} \le (1/2)\left(1 + \sqrt{1 + 4t_k^2}\right)\;\forall k\ge 0$, hence the name: ``Inverted FISTA sequence''. 
            \item $\prod_{i = 1}^k\max(1, \rho_{k - 1})(1 - \alpha_k) = \alpha_k^2 \;(\forall k \ge 0)$. 
        \end{enumerate}
    \end{lemma}
    \begin{proof}

    \end{proof}

    \begin{lemma}[Chambolle, Dossal 2015 sequences as R-WAPG sequences]
        Fix rabitrary $a \ge 2$.
        Define $\forall k \ge 1$ the sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ by 
        \begin{align*}
            \alpha_k &= a/(k + a), 
            \\
            \rho_k &= \frac{(k + a)^2}{(k + 1)(k + a + 1)}. 
        \end{align*}
        If $\mu = 0$, $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ is a valid R-WAPG sequence that satisfies Definition \ref{def:rwapg-seq}. 
        Applying this sequence to Definition \ref{def:r-wapg-momentum-form}, which is the the momentum form of R-WAPG, then R-WAPG is reduced to the following scheme: 
        \begin{align*}
        \end{align*}
    \end{lemma}
    \begin{proof}

    \end{proof}

\subsection{The method of Inexact Momentum}
    This section introduces an algorithm of our creation through the remark of Proposition \ref*{prop:stepwise-lyapunov}. 
    The algorithm estimates the $\mu$ constant as the algorithm executes and pools the information using the Bregman Divergence of the smooth part function $f$.

\subsection{Numerical experiments}

\bibliographystyle{siam}
\bibliography{references/Inexact_Momentum.bib}

\appendix

\end{document}