### 2025-01-23 Latex Templates/Scientific/main.tex
>w<, latex changes on formatting

\documentclass[12pt]{article}

% \input{presets/wang_full.tex}
\usepackage{ifthen}\newboolean{draftmode}\setboolean{draftmode}{true}
\input{presets/wang/input.tex}
\input{presets/misc.tex}
\input{presets/julia_lstlisting.tex}

\begin{document}

\title{{\fontfamily{ptm}\selectfont Lorum ipsum}}

\author{
    Author 1 Name, Author 2 Name
    \thanks{
        Subject type, Some Department of Some University, Location of the University,
        Country. E-mail: \texttt{author.name@university.edu}.
    }
}

\date{March 2, 2020}

\maketitle
\todoinline{This paper is currently in draft mode. Check source to change options. }
% \vskip 8mm

\begin{abstract} 
    \noindent
    Lorem ipsum dolor sit amet, dicta iudicabit consequat ex vix, veniam legimus appetere has id, an pri graece epicuri detraxit. Ea aliquam expetendis posidonium eos, nam invenire corrumpit imperdiet ei. Et constituto dissentias usu, mel solum erant et. Mel dolorem menandri in.
    \todo{This is just psuedo text. }
\end{abstract}

\noindent{\bfseries 2010 Mathematics Subject Classification:}
Primary 47H05, 52A41, 90C25; Secondary 15A09, 26A51, 26B25, 26E60, 47H09, 47A63.
\noindent{\bfseries Keywords:}

% ==============================================================================

\section{Schur's Cone and Its Polar}
    \begin{example}
        Define
        \begin{equation}\label{eqn:cone-1}
            Q = \left\lbrace
                x\in \RR^n \left|
                    \sum_{i = 1}^{k} x_i
                    \ge 0, \;
                    k = 1 ,\cdots, n-1, \;
                    x_1 + \cdots + x_n = 0
                \right.
            \right\rbrace.
        \end{equation}
        Then
        \begin{align}
            Q^\circ &= \left\lbrace
               y\in\RR^n |
               \langle y, x\rangle \ge 0 \; \forall x \in Q
            \right\rbrace
            \\
            &=
            \left\lbrace
                y\in\RR^n | y_{1} \ge y_2\ge \cdots  \ge y_n
            \right\rbrace.
        \end{align}
    \end{example}
    \begin{proof}
        We have $Q = H(\RR^{n - 1}_+)$, where
        \begin{align}
            H =
            \begin{bmatrix}
                1 &  & &
                \\
                -1 & 1 & &
                \\
                 & -1 & &
                \\
                &  & \ddots&
                \\[-1em]
                &  & & 1
                \\
                &  && -1
            \end{bmatrix},
        \end{align}
        $H$ is a bi-diagonal matrix with zeros on the diagonal and -1 on its sub-diagonals. Indeed, for $x = (x_1, \cdots, x_{n - 1})\in \RR^{n - 1}_+$, we have
        \begin{align}
            y &= H x
            \\
            &=
            \begin{bmatrix}
                1 \\ -1 \\ 0 \\ \vdots \\ 0
            \end{bmatrix}x_1 +
            \begin{bmatrix}
                0 \\ 1 \\ -1 \\ \vdots \\ 0
            \end{bmatrix}x_2 +
            \cdots
            +
            \begin{bmatrix}
                0 \\ 0 \\ \vdots \\ 1 \\ -1
            \end{bmatrix} x_{n - 1}
            \\
            &=
            \begin{bmatrix}
                x_1 \\ x_2 - x_1 \\ x_3 - x_2
                \\ \vdots
                \\-x_{n - 2} + x_{n - 1}\\
                 - x_{n - 1}
            \end{bmatrix}
            = \begin{bmatrix}
                y_1 \\ y_2 \\ y_3 \\ \vdots \\ y_{n - 1} \\ y_n
            \end{bmatrix},
        \end{align}
        where $y_1 \ge 0, y_2 + y_1 \ge 0, \cdots, y_{n - 1}+ \cdots + y_1 \ge 0$, and $y_1 + \cdots + y_n = 0$. So $y\in Q$. Hence $H(\RR^{n - 1}_+)\subseteq Q$. Similarly, one can write for $y\in Q$, $\exists x \in \RR_+^{n- 1}$ with $y = Hx$. In fact, $x = y_1, x_2 = y_1 + y_2, \cdots, x_{n - 1} = y_1 + \cdots + y_{n - 1}, -x_{n - 1} = y_n$ because $y_1 + \cdots + y_n = 0$. In particular, $x\in \RR^{n - 1}_+$.
        \par
        Note that $y\in Q^\circ\iff y^{\intercal}a_i \ge 0$ for $i = 1,\cdots, n -1$. Where
        \begin{align}
            a_1 = \begin{bmatrix}
                1 \\ -1 \\ 0 \\ \vdots \\ 0,
            \end{bmatrix},
            a_2 = \begin{bmatrix}
                0 \\ 1 \\ -1 \\ \vdots \\ 0,
            \end{bmatrix},
            & \cdots,
            a_n =
            \begin{bmatrix}
                0 \\ \vdots \\ 0  \\ 1 \\ -1
            \end{bmatrix}.
        \end{align}
        Now $y^{\intercal} a_i \ge 0 \iff y_i - y_{i + 1}\ge 0$, for $i = 1,\cdots, n-1$. Hence $Q^\circ = \{ y\in \RR^n | y_1 \ge y_2 \ge \cdots \ge y_n\}$.

    \end{proof}

\section{Lorentz Cone and its Polar}
    \begin{example}
        \begin{equation}
            L = \left\lbrace
            (\xi, t) \in \RR^n \left| \;
                \Vert \xi\Vert \le t, \xi \in \RR^{n - 1}
            \right.
            \right\rbrace.
        \end{equation}
        Then the polar cone of $L$ is $L^\circ = \{(x, s) \;|\; \langle (x, s), (\xi, t)\rangle \ge 0\}\} = L$.
    \end{example}
    \begin{proof}
        \par
        Let $(x, s) \in L^\circ$. Then $\langle (x, s), (\xi, t)\rangle \ge 0$, $\forall (\xi, t) \in L$, i. e. $\langle x, \xi\rangle + st \ge 0$.
        \par
        For $t > 0$, we have $\langle x, \xi/t\rangle \ge -s$, i.e. $\langle x, - \xi/t\rangle \le s$. Using $\Vert \xi\Vert \le t$ and by taking supremum over $\Vert \xi/t\Vert\le 1$, we obtain  $\Vert x\Vert \le s$.
        \par
        Hence $L^\circ = L$.
    \end{proof}
    \begin{corollary}
        Let $A \in \RR^{n\times n}$ be positive definite. Define
        $$
            L_A =
            \left\lbrace
                (\xi, t)\in \RR^n \left|
                    \sqrt{\langle \xi, A\xi\rangle} \le t
                \right.
            \right\rbrace.
        $$
        Then $(L_A)^\circ = \{ (\xi, t) \;|\; \sqrt{\langle \xi, A^{-1}\xi\rangle}\le t\}$.
    \end{corollary}

\section{Testing the references}
    This book by Nesterov \cite{nesterov_lectures_2018}. 
    This book by Heinz \cite{bauschke_convex_2017}.

\section{Testing figures and images}
    In Figure \ref{fig:two-test-figures}, it has Figure \ref{fig:two-test-fig-1} on the left and Figure \ref{fig:two-test-fig-2} on the right. 
    \begin{figure}[H]
        \begin{subfigure}[b]{0.47\textwidth}
            \centering
            \includegraphics[width=\textwidth]{assets/test-image-1.png}
            \caption{Alto the unicorn. }
            \label{fig:two-test-fig-1}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.47\textwidth}
            \centering
            \includegraphics[width=\textwidth]{assets/test-image-2.png}
            \caption{Alto and Minty. }
            \label{fig:two-test-fig-2}
        \end{subfigure}
        \caption{The above 2 images are test images for the figure, subfigure latex commands. }
    \label{fig:two-test-figures}
    \end{figure}

\section*{Acknowledgments}
    XW were supported by NSERC Discovery grants.

\bibliographystyle{siam}

\bibliography{references/Books.bib}

\end{document}