### 2025-01-24 Notebook LaTEX/Proposal/presentation.tex
>w<, changes

\documentclass[11pt]{beamer}

\usetheme{Madrid}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{wrapfig}
\usepackage{subcaption}
\graphicspath{{.}}
\input{presets/wang/custom_commands.tex}

\author{Hongda Li}
\title[Thesis Proposal Talk]{First Order Nonsmooth Optimization: Catalyst Acceleration and Unifying Nesterov's Acceleration}
\newcommand{\email}{lalala@lala.la}
\setbeamercovered{transparent}
\setbeamertemplate{navigation symbols}{}
%\logo{}
\institute[UBCO]{
    University of British Columbia Okanagan
}
\date{\today}
\subject{Nesterov's acceleration and its applications}

% ---------------------------------------------------------
% Selecione um estilo de referência
\bibliographystyle{IEEEtran}

%\bibliographystyle{abbrv}
%\setbeamertemplate{bibliography item}{\insertbiblabel}
% ---------------------------------------------------------

% ---------------------------------------------------------
\newtheorem{remark}{Remark}
\newtheorem{assumption}{Assumption}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}
\begin{frame}{Overview}
    This talks will be based on the content of our draft paper and selected content of the Catalyst Meta Acceleration Framework. 
    Our preprint: 
    \begin{enumerate}
        \item X. Wang and H. Li, \textit{A Parameter Free Accelerated Proximal Gradient Method Without Restarting}, preprint, (2025).
    \end{enumerate}
    Catalyst Meta Acceleration:  
    \begin{enumerate}
        \item H. Lin, J. Mairal and Z. Harchaoui, \textit{A universal catalyst for first-order optimization}, in NISP, vol. 28, (2015). 
        \item \underline{\hspace{4em}}, \textit{Catalyst acceleration for first-order convex optimization: from theory to practice}, JMLR, 18 (2018), pp. 1–54.
    \end{enumerate}
\end{frame}
\begin{frame}{ToC}
    \tableofcontents
\end{frame}

\section{Introduction}
    \subsection{Notations and preliminaries}
        \begin{frame}{Notations and preliminaries}
            Throughout this talk, let $\RR^n$ be the ambient space equiped with Euclidean inner product and norm. 
            We consider 
            \begin{align}\label{eqn:additive-comp-obj}
                \min_{x \in \RR^n} \left\lbrace
                    F(x):= f(x) + g(x)
                \right\rbrace.
            \end{align}
            Unless specified, assume: 
            \begin{enumerate}
                \item $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex, 
                \item $g:\RR^n \rightarrow \overline \RR$ is closed convex proper. 
            \end{enumerate}
        \end{frame}
        \begin{frame}{Our works on R-WAPG}

        \end{frame}
        \begin{frame}{Introduction to Catalyst Acceleration}

        \end{frame}
\section{Content of the draft paper}

    \subsection{Direction of future works}

\section{Selected contents from Catalyst Meta Accelerations}
    \subsection{Direction of future works}

\section{References}
    \begin{frame}{Citation examples}
        Citation examples \cite{chambolle_convergence_2015}
    \end{frame}
    \begin{frame}[allowframebreaks]{References}
        % \bibliographystyle{apalike}
        \bibliography{references/proposal.bib}
    \end{frame}

\end{document}