### 2025-01-24 Notebook LaTEX/Proposal/presentation.tex
>w<, changes

\documentclass[11pt]{beamer}

\usetheme{Madrid}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{wrapfig}
\usepackage{subcaption}
\graphicspath{{.}}
\input{presets/wang/custom_commands.tex}

\author{Hongda Li}
\title[Thesis Proposal Talk]{First Order Nonsmooth Optimization: Catalyst Acceleration and Unifying Nesterov's Acceleration}
\newcommand{\email}{lalala@lala.la}
\setbeamercovered{transparent}
\setbeamertemplate{navigation symbols}{}
%\logo{}
\institute[UBCO]{
    University of British Columbia Okanagan
}
\date{\today}
\subject{Nesterov's acceleration and its applications}

% ---------------------------------------------------------
% Selecione um estilo de referência
\bibliographystyle{IEEEtran}

%\bibliographystyle{abbrv}
%\setbeamertemplate{bibliography item}{\insertbiblabel}
% ---------------------------------------------------------

% ---------------------------------------------------------
\newtheorem{remark}{Remark}
\newtheorem{assumption}{Assumption}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}
\begin{frame}{Overview}
    This talk will be based on the content of our draft paper and selected content of the Catalyst Meta Acceleration Framework. 
    Our preprint: 
    \begin{enumerate}
        \item X. Wang and H. Li, \textit{A Parameter Free Accelerated Proximal Gradient Method Without Restarting}, preprint, (2025).
    \end{enumerate}
    Catalyst Meta Acceleration:  
    \begin{enumerate}
        \item H. Lin, J. Mairal and Z. Harchaoui, \textit{A universal catalyst for first-order optimization}, in NISP, vol. 28, (2015). 
        \item \underline{\hspace{4em}}, \textit{Catalyst acceleration for first-order convex optimization: from theory to practice}, JMLR, 18 (2018), pp. 1–54.
    \end{enumerate}
\end{frame}
\begin{frame}{ToC}
    \tableofcontents
\end{frame}

\section{Introduction}
    \subsection{Notations and preliminaries}
        \begin{frame}{Notations and preliminaries}
            Throughout this talk, let $\RR^n$ be the ambient space equiped with Euclidean inner product and norm. 
            We consider 
            \begin{align}\label{eqn:additive-comp-obj}
                \min_{x \in \RR^n} \left\lbrace
                    F(x):= f(x) + g(x)
                \right\rbrace.
            \end{align}
            Unless specified, assume: 
            \begin{enumerate}
                \item $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex, 
                \item $g:\RR^n \rightarrow \overline \RR$ is closed convex proper. 
            \end{enumerate}
        \end{frame}
        \begin{frame}{Notations and preliminaries}
            \begin{definition}[Proximal gradient operator]\label{def:proximal-gradient-operator}
                Define the proximal gradient operator $T_L$ on all $y \in \RR^n$: 
                \begin{align*}
                    T_L y := \argmin_{x \in \RR^n} \left\lbrace
                        g(x) + f(y) + \langle \nabla f(y), x - y\rangle 
                        + \frac{L}{2}\Vert x- y \Vert^2
                    \right\rbrace. 
                \end{align*}
            \end{definition}
            \begin{definition}[Gradient mapping operator]\label{def:gradient-mapping-operator}
                % Take $F := f + g$ as defined in this section. 
                Define the gradient mapping operator $\mathcal G_L$ on all $y \in \RR^n$: 
                \begin{align*}
                    \mathcal G_L (y):= L(y - T_L y). 
                \end{align*}
            \end{definition}
        \end{frame}
        \begin{frame}{Proximal gradient inequality}
            \begin{lemma}[The proximal gradient inequality]\label{thm:prox-grad-ineq}
                % Take $F:= f + g$ as defined in this section. 
                For all $y \in \RR^n$, $x \in \RR^n$, it has: 
                {\scriptsize
                \begin{align*}
                    (\forall x \in \RR^n)\quad 
                    F(x)  - F(T_Ly) - \langle L(y - T_Ly), x - y\rangle
                    - \frac{\mu}{2}\Vert x - y\Vert^2 - \frac{L}{2}\Vert y - T_Ly\Vert^2 
                    &\ge 0. 
                \end{align*}
                }
            \end{lemma}
            This lemma is crucial to developing results in our current draft paper. 
        \end{frame}
        \begin{frame}{Nesterov's estimating sequence example}
            \begin{definition}[Nesterov's estimating sequence]\label{def:nes-est-seq}
                For all $k \ge 0$, let $\phi_k : \RR^n \rightarrow\RR$ be a sequence of functions. 
                We call this sequence of functions a Nesterov's estimating sequence when it satisfies conditions: 
                \begin{enumerate}
                    \item There exists another sequence $(x_k)_{k \ge 0}$ such that for all $k \ge 0$ it has $F(x_k) \le \phi_k^*: =\min_{x}\phi_k(x)$. 
                    \item There exists a sequence of $(\alpha_k)_{k \ge 0}$ where $\alpha_k \in (0, 1)\; \forall k \ge0 $ such that for all $x \in \RR^n$ it has $\phi_{k + 1}(x) - \phi_k(x) \le - \alpha_k(\phi_k(x) - F(x))$. 
                \end{enumerate}
            \end{definition}
            The technique is widespread in the literatures and it's used to derive the convergence rate of acceleration on first order method, and the numerical algorithm itself. It is a two birds one stone technique. 
        \end{frame}
        \begin{frame}{Our works on R-WAPG}
            Here are contributions of our draft paper. 
            Recall the Nesterov's acceleration has momentum extrapolation updates on $y_{k + 1} = x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k)$. 
            We proposed the idea of R-WAPG, a generic method that: 
            \begin{enumerate}
                \item Describe for momentum sequences that doesn't follow Nesterov's rules.
                \item Unifies the convergence rate analysis for several Euclidean variants of the FISTA method. 
                \item A parameter free numerical algorithm: ``Free R-WAPG'' method that has competitive numerical performance in practical settings without restarting. 
            \end{enumerate}
            Our work is inspired by considering Nesterov's estimating sequence where $F(x_k) + R_k = \phi_k^*$. 
        \end{frame}
        \begin{frame}{Introduction to Catalyst Acceleration}

        \end{frame}
\section{Content of the draft paper}

    \subsection{Direction of future works}

\section{Selected contents from Catalyst Meta Accelerations}
    \subsection{Direction of future works}

\section{References}
    \begin{frame}{Citation examples}
        Citation examples \cite{chambolle_convergence_2015}
    \end{frame}
    \begin{frame}[allowframebreaks]{References}
        % \bibliographystyle{apalike}
        \bibliography{references/proposal.bib}
    \end{frame}

\end{document}