### 2025-05-02 Notebook LaTEX/R-WAPG/presentation.tex
>w<, changes

\documentclass[11pt]{beamer}
\usetheme{Madrid}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
% \DeclareMathOperator{\argmin}{argmin}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage{mathtools}

% THEMES AND BEAMER SETTINGS ===================================================
% \usetheme{Madrid}
\graphicspath{{.}}
\input{presets/wang/custom_commands.tex} % IMPORT WANG'S LATEX CUSTOM COMMANDS. 
\setbeamertemplate{theorems}[numbered] % ADD NUMBERING TO ALL AMS THEOREMS. 
\setbeamertemplate{footline}[frame number] % ADD PAGE NUMBERS ON BOTTOM. 
\setbeamertemplate{blocks}[rounded][shadow=true]
\setbeamertemplate{navigation symbols}{} 
% \setbeamercolor{block title}{bg=cyan,fg=black}  % CHANGE THE BLOCK STYLE IN BEAMER. 
% \setbeamercolor{block body}{bg=lime,fg=black} % CHANGE THE BLOCK STYLE IN BEAMER.

% BIB STYLES SETTINGS ----------------------------------------------------------
\setbeamertemplate{bibliography item}{\insertbiblabel}
\setbeamerfont{bibliography item}{size=\footnotesize}
\setbeamerfont{bibliography entry author}{size=\footnotesize}
\setbeamerfont{bibliography entry title}{size=\footnotesize}
\setbeamerfont{bibliography entry location}{size=\footnotesize}
\setbeamerfont{bibliography entry note}{size=\footnotesize}
% \setbeamercovered{transparent}  % GREY OUT PAUSED FUTURE ITEMS IN SLID. 
\bibliographystyle{siam}

% SLIDE INFORMATION ============================================================
\author{Hongda Li, Wang Xianfu}
\title[Thesis Proposal Talk]{
    Relaxed Weak Accelerated Proximal Gradient Method: a Unified Framework for Nesterov's Accelerations
}
% \newcommand{\email}{lalala@lala.la}
\institute[UBCO]{
    University of British Columbia Okanagan
}
\date{\today}
\subject{Nesterov's acceleration and its applications}

% SLIDES ELEMENTS CUSTOMIZATIONS ===============================================
\theoremstyle{definition}
\newtheorem{remark}{Remark}[section]
\newtheorem{assumption}{Assumption}[section]
\newtheorem{proposition}{Proposition}[section]

% DOCUMENT STARTS ==============================================================
\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\begin{frame}{ToC}
    \tableofcontents
\end{frame}

\section{Introduction and background}
    \subsection{Motivations}
        \begin{frame}{Nesterov's accelerated gradient in 1983}
            Let's recall a classic. 
            \begin{block}{Nesterov's accelerated gradient in 1983}
                Initialized $x_0 = y_0 \in \RR^n$. 
                The original formulation by Nesterov in 1983 \cite{nesterov_method_1983} has
                {\small
                \begin{align*}
                    & x_{k + 1} := y_k - L^{-1}\nabla F(y_k),
                    \\
                    & t_{k + 1} := 1/2\left(1 + \sqrt{1 + 4t_{k}^2}\right),
                    \\
                    & \theta_{k + 1} := (t_{k} - 1)/t_{k + 1}, \label{eqn:example-algorithm}
                    \\
                    & y_{k + 1} := x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k).
                \end{align*}    
                }
            \end{block}
            If $F$ is convex and has $L$ Lipschitz smooth gradient and minimizer $x^+$ exists. 
            Then it has
            \begin{itemize}
                \item a convergence rate of $\mathcal O(1/k^2)$ of $(F(x_k) - F(x^+))_{k \ge 1}$;
                \item this convergence rate is optimal in the sense as proposed by Nesterov \cite{nesterov_lectures_2018}.
            \end{itemize}
        \end{frame}
        \begin{frame}{Chambolle, Dossal}
            Generalized by Beck \cite{beck_fast_2009-1} in 2009 is the FISTA algorithm for additive smooth and non-smooth objective. 
            It's available in their proof that the sequence $(t_k)_{k}$ can be relaxed: 
            \begin{tcolorbox}\noindent\vspace{-1em}
                \begin{align*}
                    (\forall k \ge 1)\quad t_k (t_k - 1) &\le t_{k - 1}^2. 
                \end{align*}    
            \end{tcolorbox}
            \begin{enumerate}
                \item When it's equal, as in the case of Nesterov's 1983, the convergence rate is the best. 
                \item It's showed by Chambolle and Dossal in 2015 \cite{chambolle_convergence_2015} that the choice $(k + a -1)/a$ has weak convergence of iterates $(x_k)_{k \ge 0}$ for all $a > 2$, under the same assumption as before. 
            \end{enumerate}
        \end{frame}
        \begin{frame}{V-FISTA and strongly convex variants}
            Fixing the momentum sequence gives an optimal convergence rate of the algorithm for the class of $F:\RR^n \rightarrow \RR$ that has 
            \begin{enumerate}
                \item $L$ smooth gradient, 
                \item and $\mu \ge 0$ strongly convexity. 
            \end{enumerate}
            Available in Nesterov's book \cite{nesterov_lectures_2018} and Beck's book \cite{beck_first-order_2017} showed that if the sequence $(\theta_k)_{k\ge 0}$ has: 
            \begin{tcolorbox}
                \begin{align*}
                   \theta_k = \frac{\sqrt{L/\mu} - 1}{\sqrt{L/\mu} + 1}. 
                \end{align*}
            \end{tcolorbox}
            Disregarding the definition of $(t_k)_{k \ge 0}$ for now. 
            Then the algorithm has
            \begin{enumerate}
                \item A linear convergence rate of $\mathcal O((1 - \sqrt{\mu/L})^k)$ for $(F(x_k) - F(x^+))_{k \ge 1}$. 
            \end{enumerate}
        \end{frame}
    \subsection{Our discoveries and inspirations}
        \begin{frame}{The assumptions we make throughout}
            \begin{assumption}[Convex smooth plus nonsmooth]
                Let the ambient space be $\RR^n$ equipped with Euclidean inner product and norm. 
                Define $F := f + g$.
                \begin{enumerate}
                    \item $f: \RR^n \rightarrow \RR$ is $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex.
                    \item $g: \RR^n \rightarrow \overline \RR$ is proper, closed and convex. The extended real is defined as $\overline \RR := \RR \cup \{\infty\}$.
                    \item A minimizer exists for the optimization problem: $F^+ = \min_x \left\lbrace f(x) + g(x)\right\rbrace$.
                \end{enumerate}
            \end{assumption}
            \textbf{This assumption the full scope of the theoretical and practical discussion, assume it throughout.}
        \end{frame}
        \begin{frame}{Our theoretical contributions, statement of results}
            Here are the promises: 
            \begin{enumerate}
                \item We relaxed the traditional choices of the sequence $(\theta_k)_{k \ge 1}$ in the weakest possible way to show that an upper bound for the sequence $(F(x_k) - F^+)_{k \ge 1}$ exists. 
                \item We proposed a unified framework which we call ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)" that unifies the convergence theories for functions that are strongly convex or not necessarily strongly convex. 
                \item We demonstrate that the R-WAPG framework can be represented is equivalent to various representations of the Nesterov's type accelerated proximal gradient methods with Euclidean geometry in the literature. 
            \end{enumerate}

        \end{frame}
        \begin{frame}{Summarized results of our theoretical contributions}
            Let $(\alpha_k)_{k \ge0}, (\rho_k)_{k \ge 0}$ be two sequences. 
            Suppose that $\alpha_0 \in (0, 1]$ and for all $k \ge 1$, $\alpha_k \in (\mu/L, 1)$ and define $(\rho_k)_{k\ge0 }$ by: 
            \begin{tcolorbox}\noindent\vspace{-1em}
                \begin{align*}
                    \rho_k &:= \frac{\alpha_{k + 1}^2 - (\mu/L)\alpha_{k + 1}}{(1 - \alpha_{k + 1})\alpha_k^2} \quad \forall (k \ge 0).
                \end{align*}
            \end{tcolorbox}
            We can show Nesterov's type accelerated proximal gradient algorithm (similar to what's show in prior but with $g \equiv 0$) which generates $(F(x_k) - F^+)_{k\ge 1}$ has an upper bound of: 
            \begin{tcolorbox}\noindent\vspace{-1em} 
                \begin{align*}
                    \mathcal O\left(
                        \left(
                            \prod_{i = 0}^{k - 1} \max(1, \rho_{i})
                        \right)
                        \prod_{i = 1}^{k} \left(1  - \alpha_i\right)
                    \right).
                \end{align*}
            \end{tcolorbox}
        \end{frame}
        \begin{frame}{Our convergence theory unifies variants in the literatures}
            Let's label 
            \begin{enumerate}
                \item[(i)] R-WAPG with $\mu \ge 0$, %\ref{def:wapg} 
                \item[(ii)] Chambolle, Dossal 2015 \cite{chambolle_convergence_2015} with $\mu \ge 0$,
                \item[(iii)] V-FISTA Beck (10.7.7) \cite{beck_first-order_2017}, with $\mu > 0$,
                \item[(iv)] R-WAPG with $\mu > 0$. %\ref{def:wapg}
            \end{enumerate}
            \begin{table}[H]
                \centering
                {\scriptsize
                \begin{tabular}{|l|l|l|l|l|}
                \hline
                    Algorithm 
                    & 
                    $\alpha_k$ 
                    & 
                    $\rho_k$ 
                    & 
                    $F(x_k) - F^+ \le \mathcal O(\cdot)$ 
                \\ \hline
                    (i) &
                    $\alpha_k \in(\mu/L, 1)$ &
                    $\rho_k > 0$ &
                    \begin{tabular}{l}
                        $\prod_{i = 0}^{k - 1}\max(1, \rho_i)(1 - \alpha_{i + 1})$
                        \\
                        % (Proposition \ref{prop:wapg-convergence})
                    \end{tabular}
                \\ \hline
                    (ii) &
                    $ 0 < \alpha_k^{-2} \le \alpha_{k + 1}^{-1} - \alpha_{k + 1}^{-2}$ &
                    $\rho_k \ge 1$ &
                    \begin{tabular}{l}
                        $\alpha_k^{2}$ \\ 
                        % (Theorem \ref{thm:r-wapg-on-cham-doss})
                    \end{tabular}
                \\ \hline
                    (iii) &
                    $\alpha_k = \sqrt{\mu/L}$ &
                    $\rho_k = 1$ &
                    \begin{tabular}{l}
                        $(1 - \sqrt{\mu/L})^k$,
                        \\
                        % (Theorem \ref{thm:fixed-momentum-fista}, remark)
                    \end{tabular}
                \\ \hline
                    (iv) &
                    $\alpha_k = \alpha \in (\mu/L, 1)$ &
                    $\rho_k = \rho > 0$ &
                    \begin{tabular}{l}
                        $\left(1 - \min\left(\mu/(\alpha L), \alpha\right)\right)^{k}$
                        \\
                        % (Theorem \ref{thm:fixed-momentum-fista})
                    \end{tabular}
                \\ \hline
                \end{tabular}
                }
            \end{table}
            Note that in Chambolle and Dossal \cite{chambolle_convergence_2015} and their sequence $t_k = \alpha_k^{-1}$ and $\alpha_k = a/(k + 1)$ satisfies $\alpha_k^{-1} \le \alpha_{k + 1}^{-1} - \alpha_{k + 1}^{-2}$. 
        \end{frame}
        \begin{frame}{A practical realization inspired by the theories}
            Our theories stated that, the convergence of R-WAPG algorithm is ultimately described by the sequence $(\alpha_k)_{k \ge 0}$.
            \textbf{Taking this idea further, this is what we did in practice}: 
            \begin{tcolorbox}
                We made the Free R-WAPG algorithm estimates the strong convexity $\mu$ constant through the Bregman Divergence at two successive iterates while the algorithm is running and put it into the sequence to run the algorithm. 
                \par
                It's completely parameter free by equipping it with Armijo line search. 
            \end{tcolorbox}
        \end{frame}

\section{The R-WAPG framework and convergence rate}
    \subsection{Introducing the R-WAPG framework}
        \begin{frame}{The R-WAPG sequences}
            \begin{definition}{R-WAPG sequences}\label{def:rwapg-seq}
                Assume $0 \le \mu < L$. 
                Let $(\alpha_k)_{k \ge 0}$ be a sequence such that it satisfies: 
                \begin{align*}
                    & \alpha_0 \in (0, 1], \\
                    & \alpha_k \in (\mu, 1] \quad (\forall k \ge 1). 
                \end{align*}    
                Define $(\rho_k)_{k \ge 0}$ by 
                \begin{align*}
                    \rho_k := \frac{\alpha_{k + 1}^2 - (\mu/L) \alpha_{k + 1}}{
                        (1 - \alpha_{k + 1}) \alpha_k^2.
                    }
                \end{align*}
                We all the sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge0}$ the R-WAPG sequences. 
            \end{definition}
        \end{frame}
        \begin{frame}{R-WAPG}
            \begin{definition}[relaxed weak accelerated proximal gradient (R-WAPG)]\label{def:wapg}
                Choose any $x_1 \in \RR^n, v_1 \in \RR^n$.
                Let $(\alpha_k)_{k \ge0}, (\rho_k)_{k \ge 0}$ be given by Definition \ref{def:rwapg-seq}.
                The algorithm generates a sequence of vector $(y_k, x_{k + 1}, v_{k + 1})_{k \ge 1}$ for $k\ge 1$ by the procedures:
                \begin{tcolorbox}
                    For $k=1, 2, 3, \ldots$
                    \begin{align*}
                        \gamma_k &\defeq \rho_{k -1}L\alpha_{k - 1}^2,
                        \\
                        \hat \gamma_{k + 1} & \defeq (1 - \alpha_k)\gamma_k + \mu \alpha_k = L\alpha_k^2,
                        \\
                        y_k &\defeq
                        (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k),
                        \\
                        g_k &\defeq \mathcal G_L (y_k),
                        \\
                        v_{k + 1} &\defeq
                        \hat\gamma^{-1}_{k + 1}
                        (\gamma_k(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k),
                        \\
                        x_{k + 1} &\defeq T_L y_k.
                    \end{align*}
                \end{tcolorbox}
            \end{definition}
        \end{frame}
        \begin{frame}{Convergence of R-WAPG}
            After six pages of dense proof in the paper, we can deduce the following theorem: 
            \begin{proposition}[R-WAPG convergence]\label{prop:wapg-convergence}
                Fix any arbitrary $x^* \in \RR^n, N \in \mathbb N$.
                Let vector sequence $(y_k, v_{k}, x_{k})_{k \ge 1}$ and R-WAPG sequences $\alpha_k, \rho_k$ be given by Definition \ref{def:wapg}.
                % Let $\epsilon_k, y_k$ be given by Assumption \ref{ass:regret-gap}.
                % Define $R_1 = 0$ and suppose that for $k = 1, 2, \ldots, N$, we have $R_k$ recursively given by:
                % \begin{align*}
                %     R_{k + 1}
                %     :=
                %     \frac{1}{2}\left(
                %         L^{-1} - \frac{\alpha_k^2}{\hat \gamma_{k + 1}}
                %     \right)\Vert g_k\Vert^2
                %     +
                %     (1 - \alpha_k)
                %     \left(
                %         \epsilon_k + R_k +
                %         \frac{\mu\alpha_k\gamma_k}{2\hat \gamma_{k + 1}}
                %         \Vert v_k - y_k\Vert^2
                %     \right).
                % \end{align*}
                Then for all $k = 1, 2, \ldots, N$:
                {\small
                \begin{align*}
                    & F(x_{k + 1}) - F(x^*) + \frac{L \alpha_k^2}{2}\Vert v_{k + 1} - x^*\Vert^2
                    \\
                    &\le
                    \left(
                        \prod_{i = 0}^{k - 1} \max(1, \rho_{i})
                    \right)
                    \left(
                        \prod_{i = 1}^{k} \left(1  - \alpha_i\right)
                    \right)
                    \left(
                        F(x_1) - F(x^*) + \frac{L\alpha_0^2}{2}\Vert v_1 - x^*\Vert^2
                    \right).
                \end{align*}
                }
            \end{proposition}
            To use this theorem for the convergence of existing variants of Accelerated Proximal Gradient in the literature, we need alternative representations of R-WAPG that make it similar to what appears in the literature. 
        \end{frame}
    \subsection{Different representation of accelerated proximal gradient}
        \begin{frame}{The intermediate form}
            \begin{definition}[R-WAPG intermediate form]\label{def:r-wapg-intermediate}
                {\footnotesize
                Assume $\mu < L$ and let $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ given by Definition \ref{def:rwapg-seq}.
                Initialize any $x_1, v_1$ in $\RR^n$.
                For $k \ge 1$, the algorithm generates sequence of vector iterates $(y_{k}, v_{k + 1}, x_{k + 1})_{k \ge 1}$ by the procedures:
                \begin{tcolorbox}
                    For $k = 1, 2, \ldots$
                    \begin{align*}
                        & y_{k} \defeq
                        \left(
                            1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                        \right)^{-1}
                        \left(
                            v_{k} +
                            \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                        \right),
                        \\
                        & x_{k + 1} \defeq
                        y_k - L^{-1} \mathcal G_L (y_k),
                        \\
                        & v_{k + 1} \defeq
                        \left(
                            1 + \frac{\mu}{L \alpha_k - \mu}
                        \right)^{-1}
                        \left(
                            v_k +
                            \left(\frac{\mu}{L \alpha_k - \mu}\right) y_k
                        \right) - \frac{1}{L\alpha_{k}}\mathcal G_L (y_k).
                    \end{align*}
                \end{tcolorbox}
                }
            \end{definition}
            The closest algorithm that fits the form but with $\mu = 0$ is Chapter 12 of in Ryu and Yin's Book \cite{ryu_large-scale_2022}, right after Theorem 17. 
        \end{frame}
        \begin{frame}{Intermediate form equivalency}
            \begin{proposition}[R-WAPG is equivalent to intermediate form]\label{prop:wapg-first-equivalent-repr}
                If the sequence $(y_k, v_k, x_k)_{k \ge 1}$ is produced by R-WAPG (Definition \ref{def:wapg}),
                then the iterates can be expressed without $(\gamma_k)_{k \ge1},(\hat \gamma_k)_{k \ge 2}$,  namely, for all $k\ge 1$,
                \begin{align}
                    & y_{k} =
                    \left(
                        1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                    \right)^{-1}
                    \left(
                        v_{k} +
                        \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                    \right), 
                    % \label{eqn:rwapg-first-equiv-form-eqn-1}
                    \\
                    & x_{k + 1} =
                    y_k - L^{-1} \mathcal G_L (y_k),
                    \\
                    & v_{k + 1} =
                    \left(
                        1 + \frac{\mu}{L \alpha_k - \mu}
                    \right)^{-1}
                    \left(
                        v_k +
                        \left(\frac{\mu}{L \alpha_k - \mu}\right) y_k
                    \right) - \frac{1}{L\alpha_{k}}\mathcal G_L (y_k).
                    % \label{eqn:rwapg-first-equiv-form-eqn-2}
                \end{align}
            \end{proposition}
        \end{frame}
        \begin{frame}{Similar triangle form}
            \begin{proposition}[R-WAPG is equivalent to similar triangle form]\label{prop:wagp-st-form}
                Let iterates $(y_k, x_{k}, v_{k})_{k \ge 1}$ and sequence $(\alpha_k, \rho_k)_{k \ge 0}$ be given by Definition \ref{def:r-wapg-intermediate}.
                Then for all $k \ge 1$, iterates $y_k, x_{k + 1}, v_{k + 1}$
                satisfy:
                \begin{align}
                    y_{k} &=
                    \left(
                        1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                    \right)^{-1}
                    \left(
                        v_{k} +
                        \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                    \right),
                    \label{eqn:rwapg-st-form-eqn-1}
                    \\
                    x_{k + 1} &=
                    y_k - L^{-1} \mathcal G_L (y_k),
                    \\
                    v_{k + 1} &= x_{k + 1} + (\alpha_k^{-1} - 1)(x_{k + 1} - x_k).
                    \label{eqn:rwapg-st-form-eqn-3}
                \end{align}
            \end{proposition}
            Combining both of these representations, we can draw a similar triangle. 
        \end{frame}
        \begin{frame}{Similar triangle form illustration Part I}
            To understand the name ``similar triangle form" for it in the literature, we draw it out. 
            Write $\tau_k = \frac{L - L\alpha_k}{L\alpha_k - \mu}$ so $y_k = (1 + \tau_k)^{-1}(v_k + \tau_k x_k)$ from the similar triangle form. 
            The R-WAPG sequence has $\alpha_k \in (\mu/L, 1)$ for all $k \ge 1$, so it has: 
            \begin{tcolorbox}
                \begin{figure}
                    \centering
                    \includegraphics[width=15em]{assets/st_part1.png}
                \end{figure}
            \end{tcolorbox}
        \end{frame}
        \begin{frame}{Similar triangle form illustration part II}
            Then $x_{k + 1} = y_k - L^{-1}\mathcal G_L(y_k)$ hence it has 
            \begin{tcolorbox}
                \begin{figure}
                    \centering
                    \includegraphics[width=15em]{assets/st_part2.png}
                \end{figure}
            \end{tcolorbox}
        \end{frame}
        \begin{frame}{Similar triangle form illustration part III}
            Recall that $v_{k + 1} = x_{k + 1} + (\alpha_k^{-1} - 1)(x_{k + 1} - x_k)$, so $v_{k + 1}$ is on the line defined by $x_{k + 1} - x_k$, and it looks like: 
            \begin{tcolorbox}
                \begin{figure}
                    \includegraphics[width=15em]{assets/st_part3.png}
                \end{figure}
            \end{tcolorbox}
            Let $\xi_k = \mu(L\alpha_k - \mu)^{-1}$, the intermediate form it has $v_{k+1} = (1 + \xi_k)^{-1}(v_k + \xi_k y_k) - L^{-1}\alpha_k^{-1} \mathcal G_L(y_k)$, hence the gray line. 
        \end{frame}
        \begin{frame}{Similar triangle form is equivalent to the momentum form}
            \begin{proposition}[Momentum form]\label{prop:r-wapg-momentum-repr}
                Let sequence $(\alpha_k, \rho_k)_{k \ge 0}$ and iterates $(x_k, v_k, y_k)_{k\ge 1}$ be given by the R-WAPG intermediate form.
                Then it has:
                \begin{enumerate}
                    \item[(i)] For all $k \ge 1$, the iterates $x_{k + 1}, y_{k + 1}$ satisfy:
                    \begin{align*}
                        x_{k + 1} &= y_k - L^{-1}\mathcal G_L (y_k),
                        \\
                        y_{k + 1} &=
                        x_{k + 1} +
                        \frac{\rho_k\alpha_k(1 - \alpha_k)}
                        {\rho_k\alpha_k^2 + \alpha_{k + 1}}(x_{k + 1} - x_k).
                    \end{align*}
                    \item[(ii)] $v_1 = x_1$ if and only if $y_1 = x_1$.
                    \item[(iii)] If in addition $\mu = 0$, then the momentum term in (i) admits a simpler representation:
                    \begin{align*}
                        (\forall k \ge 1) \quad
                        \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}
                        & = \alpha_{k + 1}(\alpha_k^{-1} - 1).
                    \end{align*}
                \end{enumerate}
            \end{proposition}
        \end{frame}
    \subsection{Unifying different variants of accelerated proximal gradient}
        \begin{frame}{Convergence of Chambolle and Dossal variant of FISTA}
            FISTA algorithm proposed in Chambolle and Dossal \cite{chambolle_convergence_2015} has a function value convergence fully described by Proposition \ref{prop:wapg-convergence}. 
            \begin{lemma}[R-WAPG sequence as inverted FISTA sequence]\label{lemma:inverted-fista-seq}
                Let R-WAPG sequence $(\rho_k)_{k \ge 0}, (\alpha_k)_{k \ge 0}$ be given by Definition \ref{def:rwapg-seq}.
                If $\mu = 0, \rho_k \ge 1\; \forall k \ge 0$, and $\alpha_0 = 1$, then:
                \begin{enumerate}
                    \item $\alpha_k^{-2} \ge \alpha_{k + 1}^{-2} - \alpha_{k + 1}^{-1}\; \forall k \ge 0$.
                    \item Let $t_k := \alpha_k^{-1}$, then $0 < t_{k + 1} \le (1/2)\left(1 + \sqrt{1 + 4t_k^2}\right)\;\forall k\ge 0$, hence the name: ``inverted FISTA sequence''.
                    \item $\prod_{i = 1}^k\max(1, \rho_{k - 1})(1 - \alpha_k) = \alpha_k^2 \quad (\forall k \ge 1)$.
                \end{enumerate}
            \end{lemma}
            Let $\alpha_k = a/(k + 1)$, $a \ge 2$, Applying Proposition \ref{prop:wapg-convergence} yields $\mathcal O(1/k^2)$ convergence rate of $(F(x_k) - F^+)_{k \ge 1}$ after some algebra routine. 
        \end{frame}
        \begin{frame}{In addition, our theory made a new prediction}
            \begin{theorem}[fixed momentum APG]\label{thm:fixed-momentum-fista}
                {\small
                Assume $L > \mu > 0$, let a pair of constant R-WAPG sequence: $(\alpha_k)_{k \ge0}$ is a constant and $\alpha_k \in (\mu/L, 1)$ for all $k \ge 0$.
                Define $q := \mu/L$ and for any fixed $r \in \left(\sqrt{q}, \sqrt{q^{-1}}\right)$, represent $\alpha_k = \alpha = r \sqrt{q}$. 
                Consider the algorithm with a constant momentum specified by the following:
                \begin{tcolorbox}
                    Define $\theta = \left(1 - r^{-1}\sqrt{q}\right)(1 - r\sqrt{q})(1 - q)^{-1}$.
                    \\
                    Initialize $y_1 = x_1$; for $k = 1, 2, \ldots, N$, update:
                    \begin{align*}
                        &x_{k + 1} = y_k + L^{-1}\mathcal G_L (y_k)
                        ,
                        \\
                        & y_{k + 1} = x_{k + 1} + \theta(x_{k + 1} - x_k).
                    \end{align*}
                \end{tcolorbox}
                Then the algorithm generates $(x_k)_{k \ge 1}$ such that $(F(x_{k}) - F(x^*))_{k\geq 1}$ converges at a rate of $\mathcal O\left(\left(1 - \min\left(\mu/(\alpha L), \alpha\right)\right)^k\right)$.
                }
            \end{theorem}
        \end{frame}
\section{Adaptive momentum sequence and numerical experiments}
    \subsection{Free R-WAPG}
        \begin{frame}{Free R-WAPG}
            \begin{algorithm}[H]
                \begin{algorithmic}
                {\footnotesize
                \STATE{\textbf{Input: } $f, g, L > \mu \ge 0, x_0 \in \RR^n, N \in \N$}
                \STATE{\textbf{Initialize: }$y_0 := x_0;L_0 := 1; \mu_0 := 1/2; \alpha_0 = 1$;}
                \STATE{\textbf{Compute: } $f(y_k)$; }
                \FOR{$k = 0, 1, 2, \cdots, N$}
                    \STATE{\textbf{Compute: }$\nabla f(y_k); x^+:= [I + L_k^{-1}\partial g]^{-1}(y_k - L_k^{-1}\nabla f(y_k))$;}
                    \WHILE{$L_k/2\Vert x^+ - y_k\Vert^2 < D_f(x^+, y_k)$}
                        \STATE{$L_k:= 2L_k$;}
                        \STATE{$x^+ = [I + L_k^{-1}\partial g]^{-1}(y_k - L_k^{-1}\nabla f(y_k))$; }
                    \ENDWHILE
                    \STATE{$x_{k + 1} := x^+$;}
                    \STATE{
                        \textcolor{red}{
                            $\alpha_{k + 1} := (1/2)\left(\mu_k/L_k - \alpha_{k}^2 + \sqrt{(\mu_k/L_k - \alpha_{k}^2)^2 + 4\alpha_{k}^2}\right)$;
                        }
                    }
                    \STATE{$\theta_{k + 1} := \alpha_k(1 - \alpha_k)/(\alpha_k^2 + \alpha_{k + 1})$;}
                    \STATE{$y_{k + 1}:= x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k)$; }
                    \STATE{\textbf{Compute: } $f(y_{k + 1})$}
                    \STATE{
                        \textcolor{red}{
                            $\mu_{k + 1} := D_f(y_{k + 1}, y_{k})/\Vert y_{k + 1} - y_k\Vert^2 + (1/2)\mu_k$;
                        }
                    }
                \ENDFOR
                }
                \end{algorithmic}
                \caption{Free R-WAPG}
                \label{alg:free-rwapg}
            \end{algorithm}
        \end{frame}
    \subsection{Numerical experiment, simple quadratic}
        \begin{frame}{A basic experiment on convex quadratic functions}
            Consider $\min_{x}\{F(x) := f(x) + 0\}$ with $f(x) = (1/2)\langle x, Ax\rangle$. 
            We are measuring: 
            \begin{align*}
                \delta_k := \log_2\left(
                    \frac{F(x) - F^+}{F(x_0) - F^+}\right).  
            \end{align*}
            Here are the parameters. 
            \begin{enumerate}
                \item $A \in \RR^{N \times N}$ square diagonal defined by 
                \begin{align*}
                    (\forall i = 1, \ldots, N)\; A_{i, i} = \begin{cases}
                        0 & i = 1
                        \\
                        \mu + \frac{(i - 1)(L - \mu)}{N - 1} & i \ge 2
                    \end{cases}
                \end{align*}
                \item $L = 1, \mu = 10^{-5}$ are known in prior. 
                \item All algorithm terminates after $\Vert \mathcal G_L(y_k) \Vert \le 10^{-10}$. 
                \item Initial conditions $x_0 \sim \mathcal N(I, \mathbf 0)$, i.i.d. 
            \end{enumerate}
        \end{frame}
        \begin{frame}{Experiment results 1}
            A realization of $x_0 \sim \mathcal N(I, \mathbf 0)$ is used for each trial for all algorithms. 
            Each trial results of V-FISTA, M-FISTA (Monotone restarted FISTA) and Free R-WAPG were collected. 
            We plot the medium, minimum, and maximum of $\delta_k$: 
            \begin{figure}[H]
                \begin{subfigure}[b]{0.47\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{assets/simple_regression_batched-256.png}
                    \caption{$N = 256$, simple convex quadratic.}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.47\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{assets/simple_regression_batched-1024.png}
                    \caption{$N = 1024$, simple convex quadratic. }
                \end{subfigure}
                \caption{
                    Statistics for experiments with simple convex quadratic for V-FISTA, M-FISTA, and R-WAPG.
                }
                \label{fig:simple-quadratic-NOG}
            \end{figure}
        \end{frame}
        \begin{frame}{$\mu$ estimations}
            Free R-WAPG estimates $\mu_k$ each iteration. 
            The value of $\mu_k$ had been recorded for one trial and this is a plot of the estimations: 
            \begin{figure}[H]
                \centering
                \begin{subfigure}[b]{0.47\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{assets/simple_regression_loss_sc_estimates_1024.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.47\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{assets/simple_regression_loss_1024.png}
                \end{subfigure}
                \caption{
                    $N = 1024$, the $\mu$ estimates produced by Algorithm \ref{alg:free-rwapg} (R-WAPG) is recorded.
                }
                \label{fig:simple-quadratic-r-wapg-mu-estimates}
            \end{figure}
        \end{frame}
    \subsection{Numerical experiment, LASSO}
        \begin{frame}{The LASSO problem}
            The problem of LASSO from Tibshirani \cite{tibshirani_regression_1996} is the optimization problem $\min_{x \in \RR^n}\{(1/2)\Vert Ax - b\Vert^2 - \lambda\Vert x\Vert_1\}$. 
            \begin{enumerate}
                \item $A \in \RR^{N \times N}$, full of i.i.d random variable from a standard normal distribution. 
                \item Computed in prior, $L, \mu$ are parameters estimated by $\mu = 1/\Vert (A^TA)^{-1}\Vert$ and $L = \Vert A^TA\Vert$. The norm is the spectral norm. 
                \item The smallest value of $F(x_k)$, for all algorithms, over all tries are taken as an estimate of $F^+$. 
            \end{enumerate}
        \end{frame}
        \begin{frame}{Results plotted}
            With $A\in \RR^{N\times N}$ fixed, for each $x_0 \sim \mathcal N(I, \mathbf 0)$ realized, 30 experiments are preformed and the medium, minimum and maximum of $\delta_k$ are recorded for M-FISTA, Free R-WAPG, and V-FISTA. 
            \begin{figure}[H]
                \begin{subfigure}[b]{0.47\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{assets/lasso_batched_statistics_64-256.png}
                    \caption{LASSO experiment with $M = 64, N = 256$. Plots of minimum, maximum, and median $\delta_k$ with estimated $F^*$. }
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.47\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{assets/lasso_batched_statistics_64-128.png}
                    \caption{LASSO experiment with $M = 64, N = 128$. Plots of minimum, maximum, and median $\delta_k$ with estimated $F^*$. }
                \end{subfigure}
                \caption{LASSO experiments statistics for test algorithms. }
                \label{fig:batched-lasso}
            \end{figure}
        \end{frame}
        \begin{frame}{Estimates of strong convexity constant}
            The parameters in this specific trial are are $\mu = 7.432363627613958\times 10^{-18}$ and $L = 2321.737206983643$. 
            \begin{figure}[H]
                \begin{subfigure}[b]{0.47\textwidth}
                    \includegraphics[width=\textwidth]{assets/lasso_loss_256.png}
                    \caption
                    {A single run of LASSO experiment displaying $F(x_k) - F^*$ for several test algorithms.
                    }
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.47\textwidth}
                    \includegraphics[width=\textwidth]{assets/lasso_sc_estimates_256.png}
                    \caption{The $\mu_k$ estimated by FR-WAPG for one LASSO experiment. }
                \end{subfigure}
                \caption{A single LASSO experiment results, with $M = 64, N=256$. f}
                \label{fig:single-lass-mu-estimates}
            \end{figure}
        \end{frame}

\section{Our paper is on arxiv and references}
    \begin{frame}{Thanks for the participations. Merci pour les participations.}
        Our paper can be found on arxiv using the following QR code: 
        \begin{figure}
            \centering
            \includegraphics[width=17em]{assets/paper-qrcode.png}
        \end{figure}
    \end{frame}
    \begin{frame}{References}        
        \bibliography{references/R-WAPG.bib}
    \end{frame}

\end{document}