### 2025-08-14 Notebook LaTEX/Formalized Notes/Notes on QSCNVX/main.tex
>w<, nontrivial changes

\documentclass[12pt]{article}

% \input{presets/wang_full.tex}
\usepackage{ifthen}\newboolean{draftmode}\setboolean{draftmode}{true}
\input{presets/wang/input.tex}
\input{presets/misc.tex}
% \input{presets/julia_lstlisting.tex}

\begin{document}

\newcommand{\dist}{\ensuremath{\operatorname{dist}}}

\title{{\fontfamily{ptm}\selectfont Linear Convergence of Accelerated Gradient without Restart}}

\author{
    Author 1 Name, Author 2 Name
    \thanks{
        Subject type, Some Department of Some University, Location of the University,
        Country. E-mail: \texttt{author.name@university.edu}.
    }
}

\date{March 2, 2020}

\maketitle
\todoinline{This paper is currently in draft mode. Check source to change options. \cite{bauschke_convex_2017}}
% \vskip 8mm

\begin{abstract} 
    \noindent
    This is still a note for a draft so no abstract. 
\end{abstract}

\noindent{\bfseries 2010 Mathematics Subject Classification:}
Primary 47H05, 52A41, 90C25; Secondary 15A09, 26A51, 26B25, 26E60, 47H09, 47A63.
\noindent{\bfseries Keywords:}

% ==============================================================================
\section{Introduction}
    \textbf{Notations.}
    Unless specified, our ambient space is $\RR^n$ with Euclidean norm $\Vert \cdot\Vert$.
    Let $C\subseteq \RR^n$, $\Pi_C(\cdot)$ denotes the projection onto the set $C$, i.e: the closest point in $C$ to another point in $\RR^n$. 
    For a function of $F = f + g$, and a $B\ge 0$ where $f$ is $\mathcal C^1$ differentiable, and $g$ is l.s.c, we consider the proximal gradient operator: 
    \begin{align*}
        T_B(x) &= \argmin_{z} \left\lbrace
            g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{B}{2}\Vert x - z\Vert^2
        \right\rbrace
        \\
        &= \hprox_{B^{-1}g}(x - B^{-1}\nabla f(x)). 
    \end{align*}
    We also define the gradient mapping operator

\section{Precursors materials for our proofs of linear convergence}
    \begin{definition}[similar triangle form of accelerated proximal gradient]\label{def:st-apg}

    \end{definition}

\section{deriving the convergence rate}
    \subsection{preparations}
        \begin{lemma}[convergence preparations part I]

        \end{lemma}
        \begin{lemma}[convergence preparations part II]

        \end{lemma}
        \begin{lemma}[convergence preparations part III]

        \end{lemma}

\bibliographystyle{siam}

\bibliography{references/refs.bib}

\end{document}