### 2025-08-21 Notebook LaTEX/Formalized Notes/Notes on QSCNVX/main.tex
>w<, new stuff

\documentclass[12pt]{article}

% \input{presets/wang_full.tex}
\usepackage{ifthen}\newboolean{draftmode}\setboolean{draftmode}{true}
\input{presets/wang/input.tex}
\input{presets/misc.tex}
% \input{presets/julia_lstlisting.tex}

\begin{document}

\newcommand{\dist}{\ensuremath{\operatorname{dist}}}

\title{{\fontfamily{ptm}\selectfont Linear Convergence of Accelerated Gradient without Restart}}

\author{
    Author 1 Name, Author 2 Name
    \thanks{
        Subject type, Some Department of Some University, Location of the University,
        Country. E-mail: \texttt{author.name@university.edu}.
    }
}

\date{March 2, 2020}

\maketitle
\todoinline{This paper is currently in draft mode. Check source to change options. \cite{bauschke_convex_2017}}
% \vskip 8mm

\begin{abstract} 
    \noindent
    This is still a note for a draft so no abstract. 
\end{abstract}

\noindent{\bfseries 2010 Mathematics Subject Classification:}
Primary 47H05, 52A41, 90C25; Secondary 15A09, 26A51, 26B25, 26E60, 47H09, 47A63.
\noindent{\bfseries Keywords:}

\section{Introduction}
    \textbf{Notations.}
    Unless specified, our ambient space is $\RR^n$ with Euclidean norm $\Vert \cdot\Vert$.
    Let $C\subseteq \RR^n$, $\Pi_C(\cdot)$ denotes the projection onto the set $C$, i.e: the closest point in $C$ to another point in $\RR^n$. 
    For a function of $F = f + g$, and a $B\ge 0$ where $f$ is $\mathcal C^1$ differentiable, and $g$ is l.s.c, we consider the proximal gradient operator: 
    \begin{align*}
        T_B(x) &= \argmin_{z} \left\lbrace
            g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{B}{2}\Vert x - z\Vert^2
        \right\rbrace
        \\
        &= \hprox_{B^{-1}g}(x - B^{-1}\nabla f(x)). 
    \end{align*}
    We also define the gradient mapping operator $\mathcal G_B(x) = B^{-1}(x - T_B(x))$.

\section{Precursors materials for our proofs of linear convergence}
    The following two definitions defines the accelerated proximal gradient algorithm. 
    % DEFINITION =======================================================================================================
    \begin{definition}[similar triangle form of accelerated proximal gradient]\;\label{def:st-apg}\\
        The definition is about $((\alpha_k)_{k \ge 0}, (q_k)_{k \ge 0}, (B_k)_{k \ge 0}, (y_k)_{k \ge 0}, (x_k)_{k \ge -1}, (v_k)_{k \ge -1})$. 
        These sequences satisfy:
        \begin{enumerate}[nosep]
            \item $x_{-1}, y_{- 1}\in \RR^n$ are arbitrary initial condition of the algorithm;
            \item $(q_k)_{k \ge 1}$ be a sequence such that $q_k \in [0, 1)$ for all $k \ge 1$;
            \item $(\alpha_k)_{k \ge 1}$ be a sequence such that $\alpha_0 \in (0, 1]$, and for all $k \ge 1$ it has $\alpha_k \in (q_k, 1)$;
            \item $(B_k)_{k \ge 0}$ has $B_k \ge 0$. 
        \end{enumerate}
        Then an algorithm satisfies the similar triangle form of Nesterov's accelerated gradient if it generates iterates $(y_k, x_k, v_k)_{k \ge 1}$ such that for all $k\ge 0$: 
        \begin{align*}
            y_k &= \left(\frac{\alpha_k - q_k}{1 - q_k}\right)v_{k - 1} 
            + \left(\frac{1 - \alpha_k}{1 - q_k}\right) x_{k - 1},
            \\
            x_k &= T_{L_k}(y_k), D_f(x_k, y_k) \le \frac{B_k}{2}\Vert x_k - y_k\Vert^2, 
            \\
            v_k &= x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
        \end{align*}
    \end{definition}
    % DEFINITION =======================================================================================================
    \begin{definition}[relaxed momentum sequence]\label{def:rlx-momentum-seq}
        The following definition is about sequences $((\alpha_k)_{k \ge 0}, (q_k)_{k \ge 0}, (\rho_k)_{k \ge 0})$. 
        Let 
        \begin{enumerate}[nosep]
            \item $(q_k)_{k \ge 0}$ is a sequence such that $q_k \in [0, 1)$ for all $k \ge 0$;
            \item $(\alpha_k)_{k \ge 0}$ be such that $\alpha_0 \in (0, 1]$, and for all $k \ge 1$ it has $\alpha_k \in (q_k, 1)$;
            \item $(\rho_k)_{k \ge 0}$ is a strictly positive sequence for all $k \ge 1$. 
        \end{enumerate}
        The sequences $q_k, \alpha_k$ are considered relaxed momentum sequence if for all $k \ge 1$ it satisfies the relation that: 
        \begin{align*}
            \rho_{k - 1} &= \frac{\alpha_k(\alpha_k - q_k)}{(1 - \alpha_{k})\alpha_{k - 1}^2}. 
        \end{align*}
    \end{definition}
    % DEFINITION =======================================================================================================
    \begin{definition}[proximal gradient gap]\label{def:pg-gap}
        Let $F = f + g$ where $f$ is $L$ Lipschitz smooth and $g$ is convex. 
        Then the proximal gradient mapping $T_B(x) = \hprox_{B^{-1}g}(x -  B^{-1}\nabla f(x))$ is a singleton, which as domain on $\RR^n$. 
        Let $\mu, B$ be parameters such that $B > \mu \ge 0$. 
        We define the proximal gradient gap $\mathcal E(z, y, \mu)$ is a $\RR^n \times \RR^n \rightarrow \RR^n$ mapping: 
        \begin{align*}
            \mathcal E(z, y, \mu, B) &:= 
            F(z) - F(T_B(y)) 
            - \left\langle B(y - T_B(y)), z - y\right\rangle
            - \frac{\mu}{2}\Vert z - y\Vert^2
            - \frac{B}{2}\Vert y - T_B(y)\Vert^2. 
        \end{align*}
    \end{definition}
    \begin{remark}
        This expression is the same as the proximal gradient inequality. 
    \end{remark}

\section{Deriving the convergence rate}
    To derive the convergence rate of algorithm satisfying Definition \ref{def:st-apg}, \ref{def:rlx-momentum-seq}, we leverage Definition \ref{def:pg-gap}. 
    % ASSUMPTION =======================================================================================================
    \begin{assumption}[generic assumptions for convergence]\label{ass:for-cnvg}
        The following assumption is about $(F, f, g, \mathcal E, \mu, L)$, it is the configuration needed to derive the convergence rate of algorithms that satisfy Definition \ref{def:st-apg}. 
        There exists $B > \mu \ge 0$ such that the following are true. 
        \begin{enumerate}[nosep]
            \item Let $F = f + g$ where $f$ is $L$ Lipschitz smooth and, $g$ is closed convex and proper.
            \item $\forall y \in \RR^n\; \exists \bar y$ such that $\mathcal E(\bar y, y, \mu, B) \ge 0$. 
            \item For all $z, y \in \RR^n$, it has $\mathcal E(z, y, \mu, B) + \frac{\mu}{2}\Vert z - y\Vert^2 \ge 0$.   
        \end{enumerate}
        Note that, if the function is convex, all conditions are satisfies for $\mu = 0$, and for all $\bar y \in \RR^n$. 
    \end{assumption}
    % LEMMA ========================================================================================================
    \begin{lemma}[equivalent representations of the iterates part I]\label{lemma:st-iterates-alt-form-part1}
        Suppose that the sequences of $\alpha_k, q_k, y_k, v_k, x_k, B_k$ satisfy the similar triangle form, then for all $k \ge 0$ the iterates $v_k$ admits the following equivalent representations: 
        \begin{align*}
            v_k &= x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1})
            \\
            &= v_{k - 1} + \alpha_k^{-1}q_k(y_k - v_{k - 1}) - \alpha_{k}^{-1}B_k^{-1}\mathcal G_{B_k}(y_k). 
        \end{align*}
    \end{lemma}
    \begin{proof}
        Consider all $k \ge 1$. 
        The relations is direct, immediately from the update rule in Definition \ref{def:st-apg} of $y_k$ we have
        \begin{enumerate}[nosep]
            \item[(a)] $(\alpha_k - 1)x_{k - 1} = (\alpha_k - q_k)v_{k - 1} - (1 - q_k)y_k$. 
            \item[(b)] $x_k = y_k - B_k^{-1}\mathcal G_{B_k}(y_k)$. 
        \end{enumerate}
        \begin{align*}
            v_k &= x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1})
            \\
            &= (1 - \alpha_k^{-1})x_{k - 1} + \alpha_k^{-1}x_k
            \\
            &= \alpha_k^{-1}(\alpha_k - 1)x_{k - 1} + \alpha_k^{-1}x_k
            \\
            &\underset{\text{(a)}}{=} \alpha_k^{-1}(\alpha_k - q_k)v_{k - 1} - \alpha_{k}^{-1}(1 - q_k)y_k 
            + \alpha_k^{-1}x_k
            \\
            &\underset{\text{(b)}}{=} (1 - \alpha_k^{-1}q_k) v_{k - 1} - (\alpha_k^{-1} - \alpha_k^{-1}q_k)y_k
            + \alpha_k^{-1}(y_k - B_k^{-1}\mathcal G_{B_k}(y_k)). 
            \\
            &= 
            (1 - \alpha_k^{-1}q_k) v_{k - 1} + \alpha_k^{-1}q_ky_k
            - \alpha_k^{-1}B_k^{-1}\mathcal G_{B_k}(y_k)
            \\
            &= v_{k - 1} + \alpha_k^{-1}q_k(y_k - v_{k - 1}) - \alpha_{k}^{-1}B_k^{-1}\mathcal G_{B_k}(y_k). 
        \end{align*}
    \end{proof}
    \begin{lemma}[equivalent representations of the iterates part II]\;\label{lemma:st-iterates-alt-form-part2}\\
        Suppose the sequences of $\alpha_k, q_k, y_k, v_k, x_k, B_k$ satisfy the similar triangle form, then for all $k \ge 0$ the iterates $v_k$ admits the following equivalent representations: 
        \begin{align*}
            y_k &= 
            x_{k -1} + (1 - q_k)^{-1}(\alpha_{k - 1}^{-1} - 1)(\alpha_k - q_k)(x_{k - 1} - x_{k - 2})
            \\
            &= \left(\frac{\alpha_k - q_k}{1 - q_k}\right)v_{k - 1} + \left(\frac{1 - \alpha_k}{1 - q_k}\right) x_{k - 1}. 
        \end{align*}
    \end{lemma}
    \begin{proof}
        For all $k\ge 1$, from the update rules in Definition \ref{def:st-apg}: 
        \begin{align*}
            (1 - q_k)^{-1}y_k &= 
            (\alpha_k - q_k) v_{k - 1} 
            + (1 - \alpha_k)x_{k - 1}
            \\
            &= 
            (\alpha_k - q_k)\left(
                x_{k - 2} + \alpha_{k - 1}^{-1}(x_{k - 1} - x_{k - 2})
            \right) 
            + (1 - \alpha_k)x_{k - 1}
            \\
            &= 
            (\alpha_k - q_k)x_{k - 2} 
            + \alpha_{k - 1}^{-1}(x_{k - 1} - x_{k - 2}) + (1 - \alpha_k)x_{k - 1}
            \\
            &= (\alpha_k - q_k)(1 - \alpha_{k - 1}^{-1})x_{k - 2}
            + \left(
                \frac{\alpha_k - q_k}{\alpha_{k - 1}} + 1 - \alpha_k
            \right)x_{k - 1}. 
        \end{align*}
        Multiply $(1 - q_k)$ on both sides yield: 
        \begin{align*}
            y_k &= 
            \frac{(\alpha_k - q_k)(1 - \alpha_{k - 1}^{-1})}{1 - q_k}x_{k - 2} 
            + \left(
                \frac{\alpha_k - q_k}{\alpha_{k - 1}(1 - q_k)} + \frac{1 - \alpha_k}{1 - q_k}
            \right)x_{k - 1}
            \\
            &= 
            \frac{(\alpha_k - q_k)(1 - \alpha_{k - 1}^{-1})}{1 - q_k}x_{k - 2} 
            + \left(
                \frac{
                    (\alpha_{k - 1}^{-1} -1)(\alpha_k - q_k)
                    + \alpha_k - q_k + 1 - \alpha_k
                }{1 - q_k}
            \right)x_{k - 1}
            \\
            &= \frac{(\alpha_k - q_k)(1 - \alpha_{k - 1}^{-1})}{1 - q_k}x_{k - 2} 
            + \left(
                \frac{
                    (\alpha_{k - 1}^{-1} -1)(\alpha_k - q_k)g 
                }{1 - q_k} + 1
            \right)x_{k - 1}
            \\
            &= x_{k -1} + (1 - q_k)^{-1}\left(\alpha_{k - 1}^{-1} - 1\right)(\alpha_k - q_k)(x_{k - 1} - x_{k - 2}). 
        \end{align*}
    \end{proof}

    \subsection{Preparations for the convergence rate proof}
        The following lemma summarize important results that give a swift exposition for the proofs show up at the end for the convergence rate. 
        % LEMMA ========================================================================================================
        \begin{lemma}[convergence preparations part I]\label{lemma:cnvg-prep-part1}
            Let $(F, f, g,  \mathcal E, \mu, L)$ satisfies Assumption \ref{ass:for-cnvg}. 
            Suppose that 
            \begin{enumerate}[nosep]
                \item The sequence $(y_k, v_k, x_k)_{k \ge 0}$ satisfies Definition \ref{def:st-apg} where $T_B$ is defined on $F = f + g$. 
                \item The sequences $(\alpha_k)_{k\ge 0}, (\rho_{k})_{k \ge 0}, (q_k)_{k \ge 0}$ satisfies the definition of relaxed momentum sequence.
                \item We choose the parameters $q_k$ has $q_k = \mu/B_k$, with $B_k > \mu$, for all $k \ge 0$. 
            \end{enumerate}
            Then, for all $\bar x \in \RR^n$, $k \ge 0$: 
            \begin{align*}
                & \frac{B_k\alpha_k^2}{2}\Vert \bar x - v_k\Vert^2 
                -\frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x - v_{k - 1}\Vert^2
                \\
                &=
                \frac{\alpha_k \mu}{2}
                \Vert \bar x - v_{k - 1}\Vert^2
                + \frac{q_k\mu}{2}\Vert y_k - v_{k - 1} \Vert
                + \frac{q_k}{2\mu}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                \\ &\quad 
                - \langle
                    q_k(y_k - v_{k - 1}) + \alpha_k(v_{k - 1} - \bar x), \mathcal G_{B_k}(y_k)
                \rangle
                    + \alpha_k \mu
                    \left\langle 
                        v_{k - 1} - \bar x, 
                        y_k - v_{k - 1}
                    \right\rangle. 
            \end{align*}
        \end{lemma}
        \begin{proof}
            Consider any $\bar x \in \RR^n$. 
            \begin{align*}
                & \frac{B_k\alpha_k^2}{2}\Vert \bar x - v_k\Vert^2 
                \\
                &\underset{(1)}{=} \frac{B_k\alpha_k^2}{2}\left\Vert
                    \bar x - 
                    \left(
                        v_{k - 1} 
                        + \alpha_k^{-1}q_k(y_k - v_{k - 1})
                        - \alpha_k^{-1} B_k^{-1} \mathcal G_{B_k}(y_k)
                    \right)
                \right\Vert^2
                \\
                &= \frac{B_k\alpha_k^2}{2}\left\Vert
                    (\bar x - v_{k - 1})
                    - \alpha_k^{-1}\left(
                        q_k(y_k - v_{k - 1})
                        -  B_k^{-1} \mathcal G_{B_k}(y_k)
                    \right)
                \right\Vert^2
                \\
                &= \frac{B_k\alpha_k^2}{2}
                \Vert \bar x - v_{k - 1}\Vert^2
                + \frac{B_k}{2}\left\Vert 
                    q_k(y_k - v_{k - 1})
                    -  B_k^{-1} \mathcal G_{B_k}(y_k)
                \right\Vert^2
                    \\ &\quad 
                    - \alpha_k B_k
                    \left\langle 
                        \bar x - v_{k - 1}, 
                        q_k(y_k - v_{k - 1}) - B_k^{-1}\mathcal G_{B_k}(y_k)
                    \right\rangle
                \\
                &= 
                \frac{B_k\alpha_k^2}{2}
                \Vert \bar x - v_{k - 1}\Vert^2
                + \frac{B_kq_k^2}{2}\Vert y_k - v_{k - 1} \Vert
                + \frac{1}{2B_k}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                - q_k\langle 
                    y_k - v_{k - 1}, \mathcal G_{B_k}(y_k)
                \rangle
                    \\ &\quad 
                    + B_k\alpha_k
                    \left\langle 
                        v_{k - 1} - \bar x, 
                        q_k(y_k - v_{k - 1}) - B_k^{-1}\mathcal G_{B_k}(y_k)
                    \right\rangle
                \\
                &= 
                \frac{B_k\alpha_k^2}{2}
                \Vert \bar x - v_{k - 1}\Vert^2
                + \frac{B_kq_k^2}{2}\Vert y_k - v_{k - 1} \Vert
                + \frac{1}{2B_k}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                - q_k\langle 
                    y_k - v_{k - 1}, \mathcal G_{B_k}(y_k)
                \rangle
                    \\ &\quad 
                    + 
                    B_k\alpha_k\left\langle 
                        v_{k - 1} - \bar x, 
                        q_k(y_k - v_{k - 1}) 
                    \right\rangle
                    - \alpha_k \left\langle 
                        v_{k - 1} - \bar x,
                        \mathcal G_{B_k}(y_k)
                    \right\rangle
                \\
                &= 
                \frac{\alpha_k^2B_k}{2}
                \Vert \bar x - v_{k - 1}\Vert^2
                + \frac{q_k^2B_k}{2}\Vert y_k - v_{k - 1} \Vert
                + \frac{1}{2B_k}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                    \\ &\quad 
                    - \langle
                        q_k(y_k - v_{k - 1}) + \alpha_k(v_{k - 1} - \bar x), \mathcal G_{B_k}(y_k)
                    \rangle
                    \\ &\quad 
                    + \alpha_k q_k B_k
                    \left\langle 
                        v_{k - 1} - \bar x, 
                        y_k - v_{k - 1}
                    \right\rangle. 
            \end{align*}
            At (1) we used Lemma \ref{lemma:st-iterates-alt-form-part1}. 
            Subtracting $-\frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x - v_{k - 1}\Vert^2$ from both sides, the coefficient for $\Vert \bar x - v_{k - 1}\Vert^2$ comes out to be: 
            \begin{align*}
                & \frac{\alpha_k^2B_k}{2}
                - \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2} 
                = \frac{B_k}{2}(
                    \alpha_k^2 + (1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2
                )
                \underset{(1)}{=} \frac{B_k\alpha_kq_k}{2}. 
            \end{align*}
            At (1), we used the relation $(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2 = \alpha_k(\alpha_k - q_k)$ as in Definition \ref{def:rlx-momentum-seq}, for all $k \ge 1$. 
            Therefore, we have the equality: 
            \begin{align*}
                &
                \frac{B_k\alpha_k^2}{2}\Vert \bar x - v_k\Vert^2 
                -\frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x - v_{k - 1}\Vert^2
                \\
                &= 
                \frac{\alpha_k q_k B_k}{2}\Vert \bar x - v_{k - 1}\Vert^2
                + \frac{q_k^2B_k}{2}\Vert y_k - v_{k - 1} \Vert
                + \frac{1}{2B_k}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                    \\ &\quad 
                    - \langle
                        q_k(y_k - v_{k - 1}) + \alpha_k(v_{k - 1} - \bar x), \mathcal G_{B_k}(y_k)
                    \rangle
                    \\ &\quad 
                    + \alpha_k q_k B_k
                    \left\langle 
                        v_{k - 1} - \bar x, 
                        y_k - v_{k - 1}
                    \right\rangle. 
                \\
                &\underset{(1)}{=} 
                \frac{\alpha_k \mu}{2}
                \Vert \bar x - v_{k - 1}\Vert^2
                + \frac{q_k\mu}{2}\Vert y_k - v_{k - 1} \Vert
                + \frac{q_k}{2\mu}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                    \\ &\quad 
                - \langle
                        q_k(y_k - v_{k - 1}) + \alpha_k(v_{k - 1} - \bar x), \mathcal G_{B_k}(y_k)
                    \rangle
                    + \alpha_k \mu
                    \left\langle 
                        v_{k - 1} - \bar x, 
                        y_k - v_{k - 1}
                    \right\rangle. 
            \end{align*}
            At (1), we used the relation that $B_k = \mu/q_k$, for all $k \ge 0$. 
        \end{proof}
        % LEMMA ========================================================================================================
        \begin{lemma}[convergence preparations part II]\label{lemma:cnvg-prep-part2}
            The iterates $(y_k, x_k, v_k)_{k \ge 0}$ satisfies Definition \ref{def:st-apg} then, for all $k \ge 0, \bar x \in \RR^n$ the following identities: 
            \begin{align*}
                \alpha_k(v_{k - 1} - \bar x) 
                + q_k(y_k - v_{k - 1})
                + x_{k - 1} - y_k
                &= 
                \alpha_k(x_{k - 1} - \bar x). 
            \end{align*}
        \end{lemma}
        \begin{proof}
            We first establish two intermediate results. 
            From Definition \ref{def:st-apg}, it has for all $k \ge 0$: 
            \begin{align*}
                y_k &= 
                \left(\frac{\alpha_k - q_k}{1 - q_k}\right)v_{k - 1} 
                + \left(\frac{1 - \alpha_k}{1 - q_k}\right) x_{k - 1}
                \\
                &= 
                \left(
                    1 - \frac{1 - \alpha_k}{1 - q_k}
                \right)v_{k - 1} 
                + \left(\frac{1 - \alpha_k}{1 - q_k}\right) x_{k - 1}
                \\
                \iff 
                y_k - v_{k - 1}
                &= \left(
                    \frac{1 - \alpha_k}{1 - q_k}
                \right)(x_{k - 1} - v_{k - 1}). 
            \end{align*}
            Similarly: 
            \begin{align*}
                y_k &= 
                \left(\frac{\alpha_k - q_k}{1 - q_k}\right)v_{k - 1} 
                + \left(\frac{1 - \alpha_k}{1 - q_k}\right) x_{k - 1}
                \\
                &= 
                \left(\frac{\alpha_k - q_k}{1 - q_k}\right)v_{k - 1} 
                + \left(1 - \frac{\alpha_k - q_k}{1 - q_k}\right) x_{k - 1}
                \\
                \iff 
                y_k - x_{k - 1} &= 
                \left(
                    \frac{\alpha_k - q_k}{1 - q_k}
                \right)(v_{k - 1} - x_{k - 1}). 
            \end{align*}
            Now, we use the above two results and it derives 
            \begin{align*}
                & \alpha_k(v_{k - 1} - \bar x) 
                + q_k(y_k - v_{k - 1})
                + x_{k - 1} - y_k
                \\
                &=
                \alpha_k(v_{k - 1} - \bar x)
                + q_k\left(
                    \frac{1 - \alpha_k}{1 - q_k}
                \right)(x_{k - 1} - v_{k - 1})
                - \left(
                    \frac{\alpha_k - q_k}{1 - q_k}    
                \right)(v_{k - 1} - x_{k - 1}). 
                \\
                &= 
                \alpha_k(v_{k - 1} - \bar x)
                + (1 - q_k)^{-1}\left(
                    q_k - q_k\alpha_k +(\alpha_k - q_k)
                \right)
                (x_{k - 1} - v_{k - 1})
                \\
                &= 
                \alpha_k(v_{k - 1} - \bar x)
                + \alpha_k(x_{k - 1} - v_{k - 1})
                \\
                &= \alpha_k(x_{k - 1} - \bar x). 
            \end{align*}

        \end{proof}
        % LEMMA ========================================================================================================
        \begin{lemma}[convergence preparations part III]\;\label{lemma:cnvg-prep-part3}\\
            Suppose that all of the following are satisfied
            \begin{enumerate}[nosep]
                \item $(F, f, g, \mathcal E, \mu, L)$ satisfies Assumption \ref{ass:for-cnvg}. 
                \item The sequences $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}, (q_k)_{k \ge 0}$ satisfies Definition \ref{def:rlx-momentum-seq}. 
                \item We choose $(q_k)_{k \ge 0}$ is given by $q_k = \mu/B_k$ for all $k \ge 0$. 
                \item The sequence $(y_k, v_k, x_k)_{k \ge 0}$ satisfies Definition \ref{def:st-apg}. 
            \end{enumerate}
            Then, $\forall k \ge 1$, there exists $\bar x_k \in \RR^n$, such that: 
            \begin{align*}
                & F(x_k) - F(\bar x_k)
                + \frac{B_k\alpha_k^2}{2}\Vert \bar x - v_k\Vert^2 
                \\
                &\le 
                (1 - \alpha_k)(F(x_{k - 1}) - F(\bar x_k))
                + \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x_k - v_{k - 1}\Vert^2. 
            \end{align*}
        \end{lemma}
        \begin{proof}
            Recall Definition \ref{def:pg-gap}, consider: 
            \begin{align*}
                &\mathcal E(x_{k - 1}, y_k, \mu, B_k) 
                \\
                &=
                F(x_{k - 1}) - F(x_k)
                - \langle 
                    B_k(y_k - z_k), 
                    x_{k - 1} - y_k
                \rangle
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                - \frac{B_k}{2}\Vert y_k - x_k\Vert^2
                \\
                &= 
                F(x_{k - 1}) - F(x_k)
                - \langle 
                    \mathcal G_{B_k}(y_k), 
                    x_{k - 1} - y_k
                \rangle
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                - \frac{q_k}{2\mu}\Vert \mathcal G_{B_k}(y_k)\Vert^2. 
            \end{align*}
            On the second equality above, we used $B^{-1}_k(y_k - x_k)= \mathcal G_{B_k}(y_k)$, and $B_k = \mu/q_k$.
            For all $k \ge 0$, we define $\Xi_k$ and simplify using the above result: 
            \begin{align*}
                \Xi_k &\hspace{-0.3em}\defeq 
                \mathcal E(x_{k - 1}, y_k, \mu, B_k)
                + F(x_k) - F(\bar x_k)
                - (1 - \alpha_k)(F(x_{k - 1}) - F(\bar x_k))
                \\
                &= 
                \mathcal E(x_{k - 1}, y_k, \mu, B_k)
                + F(x_k) - \alpha_kF(\bar x_k) - (1 - \alpha_k)F(x_{k - 1})
                \\
                &=
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \langle \mathcal G_{B_k}(y_k), x_{k - 1} - y_k\rangle
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                - \frac{q_k}{2\mu}\Vert \mathcal G_{B_k}(y_k)\Vert^2. 
            \end{align*}
            Now consider the new term $\Xi_k'$ which we defined and simplify below: 
            {\allowdisplaybreaks
            \begin{align*}
                \Xi_k'
                &\hspace{-0.3em}:=
                \Xi_k + \frac{B_k\alpha_k^2}{2}\Vert \bar x_k - v_k\Vert^2 
                - \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x_k - v_{k - 1}\Vert^2
                \\
                &\underset{(1)}{=} 
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \langle \mathcal G_{B_k}(y_k), x_{k - 1} - y_k\rangle
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                - \frac{q_k}{2\mu}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                    \\&\quad 
                    + \frac{\alpha_k \mu}{2}
                    \Vert \bar x_k - v_{k - 1}\Vert^2
                    + \frac{q_k\mu}{2}\Vert y_k - v_{k - 1} \Vert^2
                    + \frac{q_k}{2\mu}\Vert \mathcal G_{B_k}(y_k)\Vert^2
                    \\ &\quad 
                    - \langle
                        q_k(y_k - v_{k - 1}) + \alpha_k(v_{k - 1} - \bar x_k), \mathcal G_{B_k}(y_k)
                    \rangle
                    + \alpha_k \mu
                    \left\langle 
                        v_{k - 1} - \bar x_k, 
                        y_k - v_{k - 1}
                    \right\rangle
                \\
                &= 
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \langle \mathcal G_{B_k}(y_k), x_{k - 1} - y_k\rangle
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                    \\&\quad 
                    + \frac{\alpha_k \mu}{2}
                    \Vert \bar x_k - v_{k - 1}\Vert^2
                    + \frac{q_k\mu}{2}\Vert y_k - v_{k - 1} \Vert^2
                    \\ &\quad 
                    - \langle
                        q_k(y_k - v_{k - 1}) + \alpha_k(v_{k - 1} - \bar x_k), \mathcal G_{B_k}(y_k)
                    \rangle
                    + \alpha_k \mu
                    \left\langle 
                        v_{k - 1} - \bar x_k, 
                        y_k - v_{k - 1}
                    \right\rangle
                \\
                &= 
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                + \frac{\alpha_k \mu}{2}
                \Vert \bar x_k - v_{k - 1}\Vert^2
                + \frac{q_k\mu}{2}\Vert y_k - v_{k - 1} \Vert^2
                    \\ &\quad 
                    - \langle
                        x_{k - 1} - y_k + q_k(y_k - v_{k - 1}) + \alpha_k(v_{k - 1} - \bar x_k), 
                        \mathcal G_{B_k}(y_k)
                    \rangle
                    \\&\quad 
                    + \alpha_k \mu
                    \left\langle 
                        v_{k - 1} - \bar x_k, 
                        q_k(y_k - v_{k - 1}) 
                    \right\rangle
                \\
                &\underset{(2)}{=} 
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                    \\&\quad 
                    + \frac{\alpha_k \mu}{2}
                    \Vert \bar x_k - v_{k - 1}\Vert^2
                    + \frac{q_k\mu}{2}\Vert y_k - v_{k - 1} \Vert^2
                    \\ &\quad 
                    - \alpha_k\langle
                        x_{k - 1} - \bar x_k, 
                        \mathcal G_{B_k}(y_k)
                    \rangle
                    + \alpha_k \mu
                    \left\langle 
                        v_{k - 1} - \bar x_k, 
                        q_k(y_k - v_{k - 1}) 
                    \right\rangle
                \\
                &=
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                    \\&\quad 
                    + \frac{\alpha_k \mu}{2}
                    \Vert \bar x_k - v_{k - 1}\Vert^2
                    + \frac{\alpha_k\mu}{2}\Vert y_k - v_{k - 1}\Vert^2
                    + \alpha_k \mu
                    \left\langle 
                        v_{k - 1} - \bar x_k, 
                        q_k(y_k - v_{k - 1}) 
                    \right\rangle
                    \\ &\quad 
                    - \alpha_k\langle
                        x_{k - 1} - \bar x_k, 
                        \mathcal G_{B_k}(y_k)
                    \rangle
                    + \frac{q_k\mu - \mu\alpha_k}{2}\Vert y_k - v_{k - 1} \Vert^2
                \\
                &= 
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                    \\&\quad 
                    \frac{\alpha_k\mu}{2}\Vert y_k - \bar x_k\Vert^2
                    - \alpha_k\langle
                        x_{k - 1} - \bar x_k, 
                        \mathcal G_{B_k}(y_k)
                    \rangle
                    + \frac{q_k\mu - \mu\alpha_k}{2}\Vert y_k - v_{k - 1} \Vert^2
                \\
                &\underset{(3)}{\le} 
                \alpha_kF(x_{k - 1}) - \alpha_k F(\bar x_k)
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2 
                    \\&\quad 
                    \frac{\alpha_k\mu}{2}\Vert y_k - \bar x_k\Vert^2
                    - \alpha_k\langle
                        x_{k - 1} - \bar x_k, 
                        \mathcal G_{B_k}(y_k)
                    \rangle. 
                \\
                &= \alpha_k\left(
                    F(x_{k  - 1}) - F(\bar x_k) 
                    - \langle x_{k - 1} - \bar x_k, \mathcal G_{B_k}(y_k)\rangle
                    +  \frac{\mu}{2}\Vert y_k - \bar x_k\Vert^2
                \right)
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2. 
            \end{align*}
            }
            At (1), we used Lemma \ref{lemma:cnvg-prep-part3}, and substituted $\Xi_k$. 
            At (2), we used Lemma \ref{lemma:cnvg-prep-part2} to simplify the inner product. 
            At (3), we used the $\alpha_k > q_k$ as in Definition \ref{def:rlx-momentum-seq}, hence it makes the coefficient $q_k \mu - \mu\alpha_k \le 0$, which gives us the inequality. 
            Now, subtracting $\mathcal E(x_{k - 1}, y_k, \mu, B_k)$ from both sides of the inequality will yield: 
            {\allowdisplaybreaks\small
            \begin{align*}
                & \Xi_k' - \mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \\
                &\le 
                \alpha_k\left(
                    F(x_{k  - 1}) - F(\bar x_k) 
                    - \langle x_{k - 1} - \bar x_k, \mathcal G_{B_k}(y_k)\rangle
                    +  \frac{\mu}{2}\Vert y_k - \bar x_k\Vert^2
                    - \mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \right)
                    \\ &\quad 
                    - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                    - (1 - \alpha_k)\mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \\
                &\underset{(1)}{=} 
                \alpha_k\left(
                    F(x_k) - F(\bar x_k) 
                    - \langle \mathcal G_{B_k}(y_k), x_{k - 1} - \bar x_k - (x_{k - 1} - y_k)\rangle
                \right)
                    \\&\quad 
                    + \alpha_k\left(
                        \frac{\mu}{2}\Vert y_k - \bar x_k\Vert^2
                        + \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                        + \frac{B_k}{2} \Vert y_k - x_k\Vert^2
                    \right)
                    \\ &\quad 
                    - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                    - (1 - \alpha_k)\mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \\
                &= \alpha_k\left(
                    F(x_k) - F(\bar x_k) 
                    - \langle \mathcal G_{B_k}(y_k), y_k - \bar x_k\rangle
                    + \frac{\mu}{2}\Vert y_k - \bar x_k\Vert^2
                    + \frac{B_k}{2} \Vert y_k - x_k\Vert^2
                    + \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                \right)
                    \\ &\quad 
                    - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                    - (1 - \alpha_k)\mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \\
                &=
                \alpha_k\left(
                    - \mathcal E(\bar x_k, y_k, \mu, B_k)
                    + \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                \right)
                    \\ &\quad 
                    - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                    - (1 - \alpha_k)\mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \\
                &\underset{(2)}{\le} 
                \frac{\alpha_k \mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                - \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                - (1 - \alpha_k)\mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \\
                &= -(1 - \alpha_k)\left(
                    \mathcal E(x_{k - 1}, y_k, \mu, B_k)
                    + \frac{\mu}{2}\Vert x_{k - 1} - y_k\Vert^2
                \right) 
                \\
                & \underset{(3)}{\le} 0. 
            \end{align*}
            }
            At (1), we substituted $\mathcal E(x_{k - 1}, y_k, \mu, B_k)$. 
            At (2), we used the $\mathcal E(\bar x_k, y_k, \mu, B_k)\le 0$ by chosing $\bar x_k = \bar y$ in Assumption \ref{ass:for-cnvg} (iii) to make the inequality. 
            At (3), we used Assumption \ref{ass:for-cnvg} (iv). 
            At this point, we had proved what we wanted becaues using the definitions of $\Xi_k, \Xi_k'$ it has:
            \begin{align*}
                & \Xi_k' - \mathcal E(x_{k - 1}, y_k, \mu, B_k) 
                \\
                &= 
                \Xi_k 
                + \frac{B_k\alpha_k^2}{2}\Vert \bar x_k - v_k\Vert^2 
                - \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x_k - v_{k - 1}\Vert^2
                - \mathcal E(x_{k - 1}, y_k, \mu, B_k) 
                \\
                &= \mathcal E(x_{k - 1}, y_k, \mu, B_k)
                + F(x_k) - F(\bar x_k)
                - (1 - \alpha_k)(F(x_{k - 1}) - F(\bar x_k))
                    \\ &\quad 
                    + \frac{B_k\alpha_k^2}{2}\Vert \bar x_k - v_k\Vert^2 
                    - \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x_k - v_{k - 1}\Vert^2
                    - \mathcal E(x_{k - 1}, y_k, \mu, B_k)
                \\
                &=  
                F(x_k) - F(\bar x_k)
                - (1 - \alpha_k)(F(x_{k - 1}) - F(\bar x_k))
                    \\ &\quad 
                    + \frac{B_k\alpha_k^2}{2}\Vert \bar x_k - v_k\Vert^2 
                    - \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x_k - v_{k - 1}\Vert^2
                \\
                &\le 0. 
            \end{align*}
        \end{proof}
    \subsection{Proving the convergence rate}
        To finally find the convergence rate, we will strengthen Assumption \ref{ass:for-cnvg}. 
        \begin{assumption}[Assumptions for linear convergence rate]

        \end{assumption}
        \begin{assumption}[Assumption for sublinear convergence rate]

        \end{assumption}

\bibliographystyle{siam}

\bibliography{references/refs.bib}

\end{document}