### 2025-09-18 Notebook LaTEX/R-WAPG/presentation-shorter.tex
>w<, changes

\documentclass[11pt]{beamer}
\usetheme{Madrid}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
% \DeclareMathOperator{\argmin}{argmin}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage{mathtools}

% THEMES AND BEAMER SETTINGS ===================================================
% \usetheme{Madrid}
\graphicspath{{.}}
\input{presets/wang/custom_commands.tex} % IMPORT WANG'S LATEX CUSTOM COMMANDS. 
\setbeamertemplate{theorems}[numbered] % ADD NUMBERING TO ALL AMS THEOREMS. 
\setbeamertemplate{footline}[frame number] % ADD PAGE NUMBERS ON BOTTOM. 
\setbeamertemplate{blocks}[rounded][shadow=true]
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize subitem}{$\circ$}
% \setbeamercolor{block title}{bg=cyan,fg=black}  % CHANGE THE BLOCK STYLE IN BEAMER. 
% \setbeamercolor{block body}{bg=lime,fg=black} % CHANGE THE BLOCK STYLE IN BEAMER.

% BIB STYLES SETTINGS ----------------------------------------------------------
\setbeamertemplate{bibliography item}{\insertbiblabel}
\setbeamerfont{bibliography item}{size=\footnotesize}
\setbeamerfont{bibliography entry author}{size=\footnotesize}
\setbeamerfont{bibliography entry title}{size=\footnotesize}
\setbeamerfont{bibliography entry location}{size=\footnotesize}
\setbeamerfont{bibliography entry note}{size=\footnotesize}
% \setbeamercovered{transparent}  % GREY OUT PAUSED FUTURE ITEMS IN SLID. 
\bibliographystyle{siam}

% --- Show a title page at the start of each section. 
% \AtBeginSection[]{
%   \begin{frame}
%   \vfill
%   \centering
%   \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%     \usebeamerfont{title}\insertsectionhead\par%
%   \end{beamercolorbox}
%   \vfill
%   \end{frame}
% }

% SLIDE INFORMATION ============================================================

\author[Hongda Li]{Hongda Li}

\title{
    Relaxed Weak Accelerated Proximal Gradient Method: a Unified Framework for Nesterov's Accelerations
}
% \newcommand{\email}{lalala@lala.la}
\institute[UBCO]{
    University of British Columbia Okanagan
}
\date[\today]{\today \\ \vspace{1cm} \tiny{Joint work with Shawn/Xianfu Wang}}
\subject{Nesterov's acceleration and its applications}

% SLIDES ELEMENTS CUSTOMIZATIONS ===============================================
\theoremstyle{definition}
\newtheorem{remark}{Remark}[section]
\newtheorem{assumption}{Assumption}[section]
\newtheorem{proposition}{Proposition}[section]

% DOCUMENT STARTS ==============================================================
\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\begin{frame}{ToC}
    \tableofcontents
\end{frame}

\section{Introduction and background}
    \begin{frame}{Nesterov's accelerated gradient in 1983}
        \begin{block}{Nesterov's accelerated gradient in 1983}
            Initialized $x_0 = y_0 \in \RR^n$. 
            The original formulation by Nesterov in 1983 \cite{nesterov_method_1983} has
            {\small
            \begin{align*}
                & x_{k + 1} := y_k - L^{-1}\nabla F(y_k),
                \\
                & t_{k + 1} := 1/2\left(1 + \sqrt{1 + 4t_{k}^2}\right),
                \\
                & \theta_{k + 1} := (t_{k} - 1)/t_{k + 1}, \label{eqn:example-algorithm}
                \\
                & y_{k + 1} := x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k).
            \end{align*}    
            }
        \end{block}
        When $F$ is convex and, $L$ Lipschitz smooth. 
        If minimizer $x^+$ exists, then: 
        Then it has
        \begin{itemize}
            \item a convergence rate of $\mathcal O(1/k^2)$ for $(F(x_k) - F(x^+))_{k \ge 1}$;
            \item this convergence rate is optimal in the sense as proposed by Nesterov \cite{nesterov_lectures_2018}.
        \end{itemize}
    \end{frame}
    \begin{frame}{Choices of momentum parameters $t_k, \theta_k$}
        \begin{block}{Here are prominent results in the literature}
            \begin{itemize}
                \item Beck, Teboulle 2009 \cite{beck_fast_2009}, the FISTA paper: $t_k (t_k - 1) \le t_{k - 1}^2$. 
                    \begin{itemize}
                        \item $(F(x_k) - F(x^+))_{k \ge 0}$ converges at rate $\mathcal O(1/k^2)$ when $t_k(t_k - 1) = t_{k - 1}^2$. 
                        \item $\theta_k$ admits no closed form. 
                    \end{itemize}
                \item Chambolle and Dossal 2015 \cite{chambolle_convergence_2015}: $t_k = (k + a -1)/a$ with $a > 2$. 
                    \begin{itemize}
                        \item Retains convergence rate $\mathcal O(1/k^2)$ with $\theta_{k} = \frac{k - 2}{k + a - 1}$. 
                        \item Iterates has weak convergence in Hilbert space. 
                    \end{itemize}
                \item Nesterov \cite{nesterov_lectures_2018} and, Beck \cite{beck_first-order_2017}
                : $\theta_k = \frac{\sqrt{L/\mu} - 1}{\sqrt{L/\mu} + 1}$.
                    \begin{itemize}
                        \item Converges at rate $\mathcal O\left(\left(1 - \sqrt{\mu/L}\right)^k\right)$, if $f$ is also $\mu$ strongly convex with $\mu > 0$. 
                    \end{itemize}
                \item Apidopoulos et al. 2018 \cite{apidopoulos_convergence_2018}: $\theta_k = \frac{n}{n + b}$ for $b \in (0, 3)$. 
                \begin{itemize}
                    \item Has convergence rate $\mathcal O(1/k^{2b/3})$. 
                \end{itemize}
            \end{itemize}    
        \end{block}
    \end{frame}
    \begin{frame}{A major question}
        The following question lead our investigation: 
        \begin{tcolorbox}
            How much can we relax the choice of $\theta_k, t_k$ and still get an upper bound for the convergence rate? 
        \end{tcolorbox}
        We proposed a framework called Relaxed, Weak Accelerated Proximal Gradient (R-WAPG) and, these are what we did: 
        \begin{itemize}
            \item An upper bound of function value with the weakest assumption on momentum parameter.
            \item A unified analysis with strong convexity $\mu \ge 0$. 
            \item Various equivalent representation of R-WAPG are given and, convergence results varified for examples in the literature. 
            \item Numerical experiments are conducted to verify the theory. 
        \end{itemize}
    \end{frame}

\section{The R-WAPG Framework}
    \begin{frame}{The assumptions we make throughout}
        \begin{assumption}[Convex smooth plus nonsmooth]
            Let the ambient space be $\RR^n$, equipped with Euclidean inner product and norm. 
            Define $F := f + g$.
            \begin{enumerate}
                \item $f: \RR^n \rightarrow \RR$ is $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex.
                \item $g: \RR^n \rightarrow \overline \RR$ is proper, closed and convex. The extended real is defined as $\overline \RR := \RR \cup \{\infty\}$.
                \item A minimizer exists for the optimization problem: $F^+ = \min_x \left\lbrace f(x) + g(x)\right\rbrace$.
            \end{enumerate}
        \end{assumption}
        \textbf{This assumption the full scope of the theoretical and practical discussion, assume it throughout.}
    \end{frame}
    \begin{frame}{Proximal gradient, gradient mapping}
        \begin{definition}[The proximal gradient operator]
            Define for all $x\in \RR^n$: 
            \begin{align*}
                T_L(y) 
                &:= \argmin_{x \in \RR^n} \left\lbrace
                    g(x) + \langle \nabla f(y), x\rangle + L/2\Vert x - y\Vert^2
                \right\rbrace 
                \\
                &= \left(I + L^{-1}\partial g\right)^{-1}
                \left(I - L^{-1}\nabla f\right)(y),
                \\
                \mathcal G_L(y)
                &:= L(I - T_L)y.
            \end{align*}
        \end{definition}
        Note, the $I$ here is the identity operator in $\RR^n$. 
        $L > 0$ is the Lipschitz smoothness parameter for $f$. 
    \end{frame}
    \begin{frame}{The R-WAPG sequences}
        In this section, we highlight key theoretical results that allowed us to unify convergence rate for various variants of FISTA in the literature.
        \begin{definition}[R-WAPG sequences]\label{def:rwapg-seq}
            Assume $0 \le \mu < L$. 
            Let $(\alpha_k)_{k \ge 0}$ be such that: 
            \begin{align*}
                & \alpha_0 \in (0, 1], \\
                & \alpha_k \in (\mu/L, 1) \quad (\forall k \ge 1). 
            \end{align*}    
            Define $(\rho_k)_{k \ge 0}$
            \begin{align*}
                \rho_k := \frac{\alpha_{k + 1}^2 - (\mu/L) \alpha_{k + 1}}{
                    (1 - \alpha_{k + 1}) \alpha_k^2.
                }
            \end{align*}
            We call the sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge0}$ the R-WAPG sequences. 
        \end{definition}
    \end{frame}
    \begin{frame}{R-WAPG}
        \begin{definition}[R-WAPG]\label{def:wapg}
            Choose any $x_1 \in \RR^n, v_1 \in \RR^n$.
            Let $(\alpha_k)_{k \ge0}, (\rho_k)_{k \ge 0}$ be given by Definition \ref{def:rwapg-seq}.
            The algorithm generates a sequence of vector $(y_k, x_{k + 1}, v_{k + 1})_{k \ge 1}$ for all $k\ge 1$ by such that:
            \begin{tcolorbox}\vspace{-1em}
                \begin{align*}
                    \gamma_k &\defeq \rho_{k -1}L\alpha_{k - 1}^2,
                    \\
                    L\alpha_k^2 &= (1 - \alpha_k)\gamma_k + \mu \alpha_k, 
                    \\
                    \hat \gamma_{k + 1} & \defeq L\alpha_k^2,
                    \\
                    y_k &=
                    (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k),
                    \\
                    g_k &= \mathcal G_L (y_k),
                    \\
                    v_{k + 1} &=
                    \hat\gamma^{-1}_{k + 1}
                    (\gamma_k(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k),
                    \\
                    x_{k + 1} &= T_L y_k.
                \end{align*}
            \end{tcolorbox}
        \end{definition}
    \end{frame}

\section{Our Theoretical Results}
    \begin{frame}{Iterates by R-WAPG has other representations}
        \begin{proposition}[Alternative representations of the iterates]\label{prop:wapg-first-equivalent-repr}
            If the sequence $(y_k, v_k, x_k)_{k \ge 1}$ is produced by R-WAPG (Definition \ref{def:wapg}),
            Then they satisfy for all $k\ge 1$: 
            {\footnotesize
            \begin{align}
                & \begin{aligned}
                    y_{k} &=
                    \left(
                        1 + \frac{L - L\alpha_{k}}{L\alpha_{k} - \mu}
                    \right)^{-1}
                    \left(
                        v_{k} +
                        \left(\frac{L - L\alpha_{k}}{L\alpha_{k} - \mu} \right) x_{k}
                    \right)
                    \\
                    &= x_{k + 1} +
                    \frac{\rho_k\alpha_k(1 - \alpha_k)}
                    {\rho_k\alpha_k^2 + \alpha_{k + 1}}(x_{k + 1} - x_k), 
                \end{aligned}
                % \label{eqn:rwapg-first-equiv-form-eqn-1}
                \\
                & x_{k + 1} =
                y_k - L^{-1} \mathcal G_L (y_k),
                \\
                & \begin{aligned}
                    v_{k + 1} 
                    &=
                    \left(
                        1 + \frac{\mu}{L \alpha_k - \mu}
                    \right)^{-1}
                    \left(
                        v_k +
                        \left(\frac{\mu}{L \alpha_k - \mu}\right) y_k
                    \right) - \frac{1}{L\alpha_{k}}\mathcal G_L (y_k)
                    \\
                    &= 
                    x_{k + 1} + (\alpha_k^{-1} - 1)(x_{k + 1} - x_k).
                \end{aligned}
            \end{align}
            }
        \end{proposition}
        If $\mu = 0$, then it simplifies:
        {\small
        \begin{align*}
            (\forall k \ge 1) \quad
            \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}
            & = \alpha_{k + 1}(\alpha_k^{-1} - 1).
        \end{align*}
        }
    \end{frame}
    \begin{frame}{Convergence of R-WAPG}
        After six pages of math in the paper, we deduced the following theorem: 
        \begin{proposition}[R-WAPG convergence]\label{prop:wapg-convergence}
            Fix any arbitrary $x^* \in \RR^n, N \in \mathbb N$.
            Let $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ be R-WAPG sequences.
            Let vector sequences $(y_k, v_{k}, x_{k})_{k \ge 1}$ given by Definition \ref{def:wapg}. 
            Then for all $k = 1, 2, \ldots, N$:
            {\small
            \begin{align*}
                & F(x_{k + 1}) - F(x^*) + \frac{L \alpha_k^2}{2}\Vert v_{k + 1} - x^*\Vert^2
                \\
                &\le
                \left(
                    \prod_{i = 0}^{k - 1} \max(1, \rho_{i})
                \right)
                \left(
                    \prod_{i = 1}^{k} \left(1  - \alpha_i\right)
                \right)
                \left(
                    F(x_1) - F(x^*) + \frac{L\alpha_0^2}{2}\Vert v_1 - x^*\Vert^2
                \right).
            \end{align*}
            }
        \end{proposition}
        To use this theorem for the convergence of existing variants of Accelerated Proximal gradient method, we need alternative representations of R-WAPG to fit what commonly appears in the literature. 
    \end{frame}
    \begin{frame}{Unifying variants of FISTA in the literatures}
        Let's label 
        \begin{enumerate}
            \item[(i)] R-WAPG with $\mu \ge 0$, %\ref{def:wapg} 
            \item[(ii)] Chambolle, Dossal 2015 \cite{chambolle_convergence_2015} with $\mu \ge 0$,
            \item[(iii)] V-FISTA Beck (10.7.7) \cite{beck_first-order_2017}, with $\mu > 0$,
            \item[(iv)] R-WAPG with $\mu > 0$. %\ref{def:wapg}
        \end{enumerate}
        \begin{table}[H]
            \centering
            {\scriptsize
            \begin{tabular}{|l|l|l|l|l|}
            \hline
                Algorithm 
                & 
                $\alpha_k$ 
                & 
                $\rho_k$ 
                & 
                $F(x_k) - F^+ \le \mathcal O(\cdot)$ 
            \\ \hline
                (i) &
                $\alpha_k \in(\mu/L, 1)$ &
                $\rho_k > 0$ &
                \begin{tabular}{l}
                    $\prod_{i = 0}^{k - 1}\max(1, \rho_i)(1 - \alpha_{i + 1})$
                    \\
                    % (Proposition \ref{prop:wapg-convergence})
                \end{tabular}
            \\ \hline
                (ii) &
                $ 0 < \alpha_k^{-2} \le \alpha_{k + 1}^{-1} - \alpha_{k + 1}^{-2}$ &
                $\rho_k \ge 1$ &
                \begin{tabular}{l}
                    $\alpha_k^{2}$ \\ 
                    % (Theorem \ref{thm:r-wapg-on-cham-doss})
                \end{tabular}
            \\ \hline
                (iii) &
                $\alpha_k = \sqrt{\mu/L}$ &
                $\rho_k = 1$ &
                \begin{tabular}{l}
                    $(1 - \sqrt{\mu/L})^k$,
                    \\
                    % (Theorem \ref{thm:fixed-momentum-fista}, remark)
                \end{tabular}
            \\ \hline
                (iv) &
                $\alpha_k = \alpha \in (\mu/L, 1)$ &
                $\rho_k = \rho > 0$ &
                \begin{tabular}{l}
                    $\left(1 - \min\left(\mu/(\alpha L), \alpha\right)\right)^{k}$
                    \\
                    % (Theorem \ref{thm:fixed-momentum-fista})
                \end{tabular}
            \\ \hline
            \end{tabular}
            }
        \end{table}
        Note that in Chambolle and Dossal \cite{chambolle_convergence_2015} and their sequence $t_k = \alpha_k^{-1}$ and $\alpha_k = a/(k + 1)$ satisfies $\alpha_k^{-1} \le \alpha_{k + 1}^{-1} - \alpha_{k + 1}^{-2}$. 
    \end{frame}

\section{Adaptive momentum sequence and numerical experiments}
    \begin{frame}{Introducing numerical experiments}
        All convex functions are strongly convex with strong convexity constant $\mu = 0$. 
        With $\mu = 0$, it has from Definition \ref{def:rwapg-seq} that for all $k \ge 0$: 
        \begin{enumerate}
            \item It allows $\alpha_k \in (0, 1)$. The sequence $(\alpha_k)_{k \ge 1}$ is as loose as possible. 
            \item It has $\rho_{k - 1} = \alpha_k/((1 - \alpha_k)\alpha_{k - 1}^2)$, hence $(1 - \alpha_k)\rho_{k - 1} = \alpha_k^2/\alpha_{k - 1}^2$. 
        \end{enumerate}
        This simplifies our convergence claim into: 
        \begin{align*}
            \prod_{i = 0}^{k - 1}
            \max(1, \rho_i)(1 - \alpha_{i + 1})
            &= 
            \prod_{i = 0}^{k - 1}
            \max(1 - \alpha_{i + 1}, \rho_i(1 - \alpha_{i + 1}))
            \\
            &= 
            \prod_{i = 0}^{k - 1}
            \max\left(
            1 - \alpha_{i + 1}, \frac{\alpha_{i + 1}^2}{\alpha_i^2}\right). 
        \end{align*}
        This insipired our numerical experiments. 
    \end{frame}
    \begin{frame}{Free R-WAPG}
        \begin{algorithm}[H]
            \begin{algorithmic}
            {\footnotesize
            \STATE{\textbf{Input: } $f, g, L > \mu \ge 0, x_0 \in \RR^n, N \in \N$}
            \STATE{\textbf{Initialize: }$y_0 := x_0;L_0 := 1; \mu_0 := 1/2; \alpha_0 = 1$;}
            \STATE{\textbf{Compute: } $f(y_k)$; }
            \FOR{$k = 0, 1, 2, \cdots, N$}
                \STATE{\textbf{Compute: }$\nabla f(y_k); x^+:= [I + L_k^{-1}\partial g]^{-1}(y_k - L_k^{-1}\nabla f(y_k))$;}
                \WHILE{$L_k/2\Vert x^+ - y_k\Vert^2 < D_f(x^+, y_k)$}
                    \STATE{$L_k:= 2L_k$;}
                    \STATE{$x^+ = [I + L_k^{-1}\partial g]^{-1}(y_k - L_k^{-1}\nabla f(y_k))$; }
                \ENDWHILE
                \STATE{$x_{k + 1} := x^+$;}
                \STATE{
                    \textcolor{red}{
                        $\alpha_{k + 1} := (1/2)\left(\mu_k/L_k - \alpha_{k}^2 + \sqrt{(\mu_k/L_k - \alpha_{k}^2)^2 + 4\alpha_{k}^2}\right)$;
                    }
                }
                \STATE{$\theta_{k + 1} := \alpha_k(1 - \alpha_k)/(\alpha_k^2 + \alpha_{k + 1})$;}
                \STATE{$y_{k + 1}:= x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k)$; }
                \STATE{\textbf{Compute: } $f(y_{k + 1})$}
                \STATE{
                    \textcolor{red}{
                        $\mu_{k + 1} := D_f(y_{k + 1}, y_{k})/\Vert y_{k + 1} - y_k\Vert^2 + (1/2)\mu_k$;
                    }
                }
            \ENDFOR
            }
            \end{algorithmic}
            \caption{Free R-WAPG}
            \label{alg:free-rwapg}
        \end{algorithm}
    \end{frame}
    \begin{frame}{A basic experiment on convex quadratic functions}
        Consider $\min_{x}\{F(x) := f(x) + 0\}$ with $f(x) = (1/2)\langle x, Ax\rangle$. 
        We are measuring: 
        \begin{align*}
            \delta_k := \log_2\left(
                \frac{F(x) - F^+}{F(x_0) - F^+}\right).  
        \end{align*}
        Here are the parameters. 
        \begin{enumerate}
            \item $A \in \RR^{N \times N}$ square diagonal defined by 
            \begin{align*}
                (\forall i = 1, \ldots, N)\; A_{i, i} = \begin{cases}
                    0 & i = 1
                    \\
                    \mu + \frac{(i - 1)(L - \mu)}{N - 1} & i \ge 2
                \end{cases}
            \end{align*}
            \item $L = 1, \mu = 10^{-5}$ are known in prior. 
            \item All algorithm terminates after $\Vert \mathcal G_L(y_k) \Vert \le 10^{-10}$. 
            \item Initial conditions $x_0 \sim \mathcal N(I, \mathbf 0)$, i.i.d. 
        \end{enumerate}
    \end{frame}
    \begin{frame}{Experiment results 1}
        A realization of $x_0 \sim \mathcal N(I, \mathbf 0)$ is used for V-FISTA, M-FISTA (Monotone restarted FISTA) and Free R-WAPG. 
        Below are plots of the medium, minimum, and maximum of $\delta_k$: 
        \begin{figure}[H]
            \begin{subfigure}[b]{0.47\textwidth}
                \centering
                \includegraphics[width=\textwidth]{assets/simple_regression_batched-256.png}
                \caption{$N = 256$, simple convex quadratic.}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.47\textwidth}
                \centering
                \includegraphics[width=\textwidth]{assets/simple_regression_batched-1024.png}
                \caption{$N = 1024$, simple convex quadratic. }
            \end{subfigure}
            \caption{
                Statistics for experiments with simple convex quadratic for V-FISTA, M-FISTA, and R-WAPG.
            }
            \label{fig:simple-quadratic-NOG}
        \end{figure}
    \end{frame}
    \begin{frame}{$\mu$ estimations}
        Free R-WAPG estimates $\mu_k$ each iteration. 
        The value of $\mu_k$ had been recorded for one trial and this is a plot of the estimations: 
        \begin{figure}[H]
            \centering
            \begin{subfigure}[b]{0.47\textwidth}
                \centering
                \includegraphics[width=\textwidth]{assets/simple_regression_loss_sc_estimates_1024.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.47\textwidth}
                \centering
                \includegraphics[width=\textwidth]{assets/simple_regression_loss_1024.png}
            \end{subfigure}
            \caption{
                $N = 1024$, the $\mu$ estimates produced by Algorithm \ref{alg:free-rwapg} (R-WAPG) is recorded.
            }
            \label{fig:simple-quadratic-r-wapg-mu-estimates}
        \end{figure}
    \end{frame}
    \begin{frame}{Real time R-WAPG upper bound}
        By collecting $(\alpha_k)_{k \ge 0}$ while the algorithm is running we made the following plot for the simple quadratic experiment. 
        \begin{figure}[H]
            \centering
            \begin{subfigure}[b]{0.75\textwidth}
                \centering
                \includegraphics[width=\textwidth]{
                    assets/simple_regression_rwapg_upperbnd_1024.png
                }
            \end{subfigure}
            \caption{
                $N = 1024$, the upper bound estimates in real time from the collected $(\alpha_k)_{k \ge 0}$ sequence. 
            }
            \label{fig:simple-quadratic-r-wapg-rwapg-upperbnd}
        \end{figure}
    \end{frame}

\section{References and draft on arxiv}
    \begin{frame}{Thanks for the participations. Merci pour les participations.}
        Our paper can be found on arxiv using the following QR code: 
        \begin{figure}
            \centering
            \includegraphics[width=15em]{assets/paper-qrcode.png}
        \end{figure}
        Questions?
    \end{frame}
    \begin{frame}{References}        
        \bibliography{references/R-WAPG.bib}
    \end{frame}

\end{document}